### 2023-11-28 methods.qmd
Testing data sources

---
title: "Methods"
author: "Jacqueline Grout"
format: html
editor: visual
---

## GP Survey

Technical annex:

Q55 base: Smoking habits - Total responses

Q55_1pct: Smoking habits - % Never smoked

Q55_2pct: Smoking habits - % Former smoker

Q55_3pct: Smoking habits - % Occasional smoker

Q55_4pct: Smoking habits - % Regular smoker

<https://gp-patient.co.uk/downloads/2023/GPPS_2023_Technical_Annex_PUBLIC.pdf>

### 2023-11-28 _targets.R

@@ -40,9 +40,6 @@ tar_option_set(
 # in {targets}, and its configuration for your machine is below.
 options(clustermq.scheduler = "multiprocess")
 
-# tar_make_future() is an older (pre-{crew}) way to do distributed computing
-# in {targets}, and its configuration for your machine is below.
-# Install packages {{future}}, {{future.callr}}, and {{future.batchtools}} to allow use_targets() to configure tar_make_future() options.
 
 # Run the R scripts in the R/ folder with your custom functions:
 tar_source()

### 2023-11-28 test_targets.R

@@ -1,0 +1,8 @@
+# load packages
+library(targets)
+
+
+source("R/functions.R")
+
+
+use_targets()

### 2023-11-28 tests.R

@@ -28,3 +28,10 @@ metric25b <- tar_read(metric25b) |> as_tibble()
 
 metric32 <- tar_read(metric32) |> as_tibble()
 
+indid <- 92304
+smok <- fingertips_data(IndicatorID = indid,AreaTypeID ="All")
+Metric05 <- Metric05 |> filter(AreaType=="GPs")
+Metric05 <- Metric05 %>% select(AreaCode,Value, Count)
+colnames(Metric05)[colnames(Metric05) == 'Value'] <- 'Metric05Prev'
+colnames(Metric05)[colnames(Metric05) == 'Count'] <- 'Metric05Count'
+#View(Metric05)

### 2023-12-01 methods.qmd
Added code to test a process to rectify ethnicity being for lsoa 2021 and gp list sizes for 2011 lsoas

@@
 <https://gp-patient.co.uk/downloads/2023/GPPS_2023_Technical_Annex_PUBLIC.pdf>
+
+## CHD synthetic prevalence estimates
+
+CHD synthetic prevalence estimates data has been removed from Fingertips by OHID.
+
+OHID have sent the archived data. This is for 7564 practices from 2015, of which 6297 are current practices.
+
+Need to identify practices with missing data and work out how to get a prevalence for them.
+
+Also have file from previous project - checked data is the same but need to check practices included - maybe there are more?
+
+## Census 2021 Ethnicity by LSOA
+
+Some LSOAs that have gp list population assigned to them do not feature in the census data of ethnicity by lsoa. Need to check if these are the changes from 2011 to 2021 census.
+
+Possible solution would be to work out the average proportions for all the other LSOAs for the practice and distribute the gp list population for the lsoas that have no match proportionally across the ethnic groups.
+
+Another option would be where lsoas in 2021 have a parent lsoa in 2011 then apply the average ethnicity % for the child lsoas to the parent (this is when one 2011 lsoa has been split into multiple 2021 lsoas.) Where the lsoas in 2021 have a child lsoa in 2011 then apply the ethnicity % for the 2021 parent to the 2011 children (this is when several lsoas have been combined into one new lsoa)
+
+join lsoa_lookup to census 21eth using the 2021 lsoa code in the lookup dataset.
+
+## Which date should be used for the gp list size dataset?
+
+Data mainly 2022/23 could take April 22 or April 23 file. Will try both. April 2023 is still using census 2011 lsoas.
+
+<https://digital.nhs.uk/data-and-information/publications/statistical/patients-registered-at-a-gp-practice/october-2022>
+
+Have decided to use October 2022 and also downloaded the GP to icb mapping file from the same period

### 2023-12-01 R/functions.R

@@ -86,7 +86,8 @@ data |>
 join_gp_and_eth <- function(eth,gp){
   gp |> 
     left_join(eth) |>
-    mutate(est_all_asian = (all_asian/sum_all)*number_of_patients,
+    mutate(est_all = (sum_all/sum_all)*number_of_patients,
+           est_all_asian = (all_asian/sum_all)*number_of_patients,
            est_bangladeshi = (bangladeshi/sum_all)*number_of_patients,
            est_chinese = (chinese/sum_all)*number_of_patients,          
            est_indian = (indian/sum_all)*number_of_patients,   
@@ -112,5 +113,4 @@ join_gp_and_eth <- function(eth,gp){
            est_other_other = (other_other/sum_all)*number_of_patients,
            est_white_other = (white_other/sum_all)*number_of_patients
            ) 
-#  |>  na.omit()
 }

### 2023-12-01 TestScript.R

@@ -101,3 +101,263 @@ Test_Eth_by_gp2 <-Test_Eth_by_gp |>
 gp_lsoa2 <- gp_lsoa |>
   group_by(practice_code) |>
   summarise(number_of_patients=sum(number_of_patients))
+
+A81001 <- Test_Eth_by_gp2 |>
+  filter(practice_code=="A81001")
+
+A81001gp <- gp_lsoa |>
+  filter(practice_code=="A81001")
+
+lsoa_eth_sum_A81001 <- lsoa_eth_sum |>
+filter(lsoa_code == "E01011959"|
+       lsoa_code == "E01012012"|
+       lsoa_code == "E01012023"|
+       lsoa_code == "E01012031"|
+       lsoa_code == "E01012038"|
+       lsoa_code == "E01012041"|
+       lsoa_code == "E01012044"|
+         lsoa_code == "E01012049"|
+         lsoa_code == "E01012081"|
+         lsoa_code == "E01012088"|
+         lsoa_code == "E01012132"|
+         lsoa_code == "E01012172"|
+         lsoa_code == "E01012187"|
+         lsoa_code == "E01012188"|
+         lsoa_code == "E01012189"|
+         lsoa_code == "E01012190"|
+         lsoa_code == "E01012191"|
+         lsoa_code == "E01012192"|
+         lsoa_code == "E01012194"|
+         lsoa_code == "E01012196"|
+         lsoa_code == "E01012201"|
+         lsoa_code == "E01012202"|
+         lsoa_code == "E01012204"|
+         lsoa_code == "E01012205"|
+         lsoa_code == "E01012206"|
+         lsoa_code == "E01012207"|
+         lsoa_code == "E01012208"|
+         lsoa_code == "E01012209"|
+         lsoa_code == "E01012210"|
+         lsoa_code == "E01012211"|
+         lsoa_code == "E01012212"|
+         lsoa_code == "E01012213"|
+         lsoa_code == "E01012215"|
+         lsoa_code == "E01012216"|
+         lsoa_code == "E01012217"|
+         lsoa_code == "E01012218"|
+         lsoa_code == "E01012219"|
+         lsoa_code == "E01012223"|
+         lsoa_code == "E01012224"|
+         lsoa_code == "E01012225"|
+         lsoa_code == "E01012226"|
+         lsoa_code == "E01012230"|
+         lsoa_code == "E01012232"|
+         lsoa_code == "E01012235"|
[…]

### 2023-12-01 _targets.R

@@ -69,14 +69,18 @@ list(
   tar_target(metric28,read_csv_file(data_path10)),
   tar_target(data_path11, "data/metric29.csv", format = "file"), # SUS data metric 29
   tar_target(metric29,read_csv_file(data_path11)),
-  tar_target(data_path12, "data/April22GPListLSOA.csv", format = "file"), #GP list size
-  tar_target(gp_list_lsoa,read_csv_file(data_path12)),
+  # tar_target(data_path12, "data/April22GPListLSOA.csv", format = "file"), #GP list size
+  # tar_target(gp_list_lsoa,read_csv_file(data_path12)),
   tar_target(data_path13, "data/EthnicityByLSOACensus21.csv", format = "file"), #Ethnicity by LSOA Census 2021
   tar_target(eth_lsoa_census,read_csv_file(data_path13)),
-  tar_target(data_path14, "data/gp-reg-pat-prac-lsoa-all.csv", format = "file"), #GP Reg Pat Prac LSOA
+  tar_target(data_path14, "data/gp-reg-pat-prac-lsoa-all.csv", format = "file"), #GP Reg Pat Prac LSOA April 2023
   tar_target(gp_reg_pat_prac_lsoa,read_csv_file(data_path14)),
   tar_target(data_path15, "data/LSOA_(2011)_to_LSOA_(2021)_to_Local_Authority_District_(2022)_Lookup_for_England_and_Wales.csv", format = "file"), #lookup
   tar_target(lsoa_lookup,read_csv_file(data_path15)),
+  tar_target(data_path17,"data/gp-reg-pat-prac-map.csv", format = "file"),
+  tar_target(gp_icb_mapping,read_csv_file(data_path17)|> 
+               select(practice_code,practice_name,sub_icb_location_code, sub_icb_location_name,
+                      icb_code,icb_name,comm_region_code,comm_region_name)),  
   tar_target(metric13,get_my_fingertips_gp_data(273,"2021/22")), # metric 13
   tar_target(metric6,get_my_fingertips_gp_data(93088,"2021/22")), # metric 6
   tar_target(metric7,get_my_fingertips_gp_data(241,"2021/22")), # metric 7
@@ -110,8 +114,13 @@ list(
                rename(value=prevalence_percent_15)), # metric 13b 22/23
   #process ethnicity data
[…]

### 2023-12-01 tests.R

@@ -1,6 +1,6 @@
 # load packages
 library(targets)
-
+library(dplyr)
 
 source("R/functions.R")
 
@@ -24,6 +24,10 @@ gp_lsoa <- tar_read(gp_lsoa) |> as_tibble()
 
 lsoa_eth_sum <- tar_read(lsoa_eth_sum) |> as_tibble()
 
+lsoa_lookup <- tar_read(lsoa_lookup) |> as_tibble()
+
+
+
 metric13b <- tar_read(metric13b) |> as_tibble()
 
 metric25b <- tar_read(metric25b) |> as_tibble()
@@ -37,3 +41,6 @@ Metric05 <- Metric05 %>% select(AreaCode,Value, Count)
 colnames(Metric05)[colnames(Metric05) == 'Value'] <- 'Metric05Prev'
 colnames(Metric05)[colnames(Metric05) == 'Count'] <- 'Metric05Count'
 #View(Metric05)
+
+
+

### 2023-12-01 methods.qmd
Incorporated code to rectify 2021 lsoa to 2011 lsoa into the targets pathway

@@
 
+Have opted for the latter solution.
+
 ## Which date should be used for the gp list size dataset?
@@
 
-Have decided to use October 2022 and also downloaded the GP to icb mapping file from the same period
+Have decided to use October 2022 and also downloaded the GP to icb mapping file from the same period and joined to the gp list size to easily create analysis by icb/sub-icb.
+
+## Clustering
+
+### Attempt 1

### 2023-12-01 R/functions.R

@@ -114,3 +114,86 @@ join_gp_and_eth <- function(eth,gp){
            est_white_other = (white_other/sum_all)*number_of_patients
            ) 
 }
+
+sum_eth_by_lsoa11 <- function(lookup,eth){
+  #join the lookup to the ethnicity data
+
+  lookup |> 
+  left_join(eth, join_by(lsoa21cd==lsoa_code)) |>
+  #sum the ethnicity in the dataframe with 2011 lsoa group by
+  group_by(f_lsoa11cd) |> 
+  summarise(sum_all = sum(sum_all),
+            all_asian = sum(all_asian),
+            bangladeshi = sum(bangladeshi),
+            chinese = sum(chinese),
+            indian = sum(indian),
+            pakistani = sum(pakistani),
+            other_asian = sum(other_asian),
+            all_black = sum(all_black),
+            blk_african = sum(blk_african),
+            blk_caribbean = sum(blk_caribbean),
+            other_blk = sum(other_blk),
+            all_mixed = sum(all_mixed),
+            mixed_other = sum(mixed_other),
+            mixed_white_asian = sum(mixed_white_asian),
+            mixed_white_blk_african = sum(mixed_white_blk_african),
+            mixed_white_blk_caribbean = sum(mixed_white_blk_caribbean),
+            all_white = sum(all_white),
+            white_british = sum(white_british),
+            white_irish = sum(white_irish),
+            white_gypsy_or_irish_traveller = sum(white_gypsy_or_irish_traveller),
+            white_roma = sum(white_roma),
+            white_other_white = sum(white_other_white),
+            all_other = sum(all_other),
+            other_arab = sum(other_arab),
+            other_other = sum(other_other),
+            white_other = sum(white_other)
+  ) |>  # have tested this works!
+    rename(lsoa_code=f_lsoa11cd)
+
+}
+
+#######
+
+sum_ethnicities <- function(gp_eth){
+
+# add up and check the total matches the number of patients
+
+  gp_eth |>
+  group_by(practice_code) |> 
+  summarise(gp_sum_est_all_asian = sum(est_all_asian),
+            gp_sum_est_bangladeshi = sum(est_bangladeshi),
+            gp_sum_est_chinese = sum(est_chinese),
+            gp_sum_est_indian = sum(est_indian),
[…]

### 2023-12-01 TestScript.R

@@ -265,99 +265,5 @@ lsoa_lookup |>
   filter(sum_n>1)
 
 
-######## THE PLAN ###################################################
-#join the lookup to the ethnicity data
-Test_join21 <-  lsoa_lookup |> 
-  left_join(lsoa_eth_sum, join_by(lsoa21cd==lsoa_code))
-#sum the ethnicity in the dataframe with 2011 lsoa group by
-Test_join11 <- Test_join21 |> 
-  group_by(f_lsoa11cd) |> 
-  summarise(sum_all = sum(sum_all),
-            all_asian = sum(all_asian),
-            bangladeshi = sum(bangladeshi),
-            chinese = sum(chinese),
-            indian = sum(indian),
-            pakistani = sum(pakistani),
-            other_asian = sum(other_asian),
-            all_black = sum(all_black),
-            blk_african = sum(blk_african),
-            blk_caribbean = sum(blk_caribbean),
-            other_blk = sum(other_blk),
-            all_mixed = sum(all_mixed),
-            mixed_other = sum(mixed_other),
-            mixed_white_asian = sum(mixed_white_asian),
-            mixed_white_blk_african = sum(mixed_white_blk_african),
-            mixed_white_blk_caribbean = sum(mixed_white_blk_caribbean),
-            all_white = sum(all_white),
-            white_british = sum(white_british),
-            white_irish = sum(white_irish),
-            white_gypsy_or_irish_traveller = sum(white_gypsy_or_irish_traveller),
-            white_roma = sum(white_roma),
-            white_other_white = sum(white_other_white),
-            all_other = sum(all_other),
-            other_arab = sum(other_arab),
-            other_other = sum(other_other),
-            white_other = sum(white_other)
-            )  # have tested this works!
-
-#join the gp list to this via the 2011 lsoa column
-#and calculate the number for each ethnicity based on ethnicity and list size
-Test_join11 <- Test_join11 |>
-  rename(lsoa_code=f_lsoa11cd)
-
-gp_lsoa_with_eth <- join_gp_and_eth(Test_join11, gp_lsoa)
 
 
-# add up and check the total matches the number of patients
-
-gp_lsoa_with_eth2 <-gp_lsoa_with_eth |>
[…]

### 2023-12-01 _targets.R

@@ -120,9 +120,17 @@ list(
   
   tar_target(joined,join_gp_and_eth(lsoa_eth_sum,gp_lsoa)),
   tar_target(lsoa_lookup_eng,lsoa_lookup |>
-               filter(str_detect(lsoa21cd, '^E')))
+               filter(str_detect(lsoa21cd, '^E'))),
+  tar_target(eth_joined,sum_eth_by_lsoa11(lsoa_lookup_eng,lsoa_eth_sum)),
+  #join the gp list to this via the 2011 lsoa column
+  #and calculate the number for each ethnicity based on ethnicity and list size
+  tar_target(gp_lsoa_with_eth,join_gp_and_eth(eth_joined,gp_lsoa)),
+  tar_target(gp_lsoa_with_eth_sum,sum_ethnicities(gp_lsoa_with_eth))
   )
   
+
+
+
 
 
 #-----------------------------------------------------------------------------

### 2023-12-01 R/functions.R
Test clustering methods

@@ -81,8 +81,7 @@ data |>
 }
 
 
-
-# join census ethnicity data to gp lists by lsoa
+# Join census ethnicity data to gp lists by lsoa
 join_gp_and_eth <- function(eth,gp){
   gp |> 
     left_join(eth) |>
@@ -155,11 +154,11 @@ sum_eth_by_lsoa11 <- function(lookup,eth){
 
 #######
 
-sum_ethnicities <- function(gp_eth){
+sum_ethnicities <- function(data){
 
 # add up and check the total matches the number of patients
 
-  gp_eth |>
+  data |>
   group_by(practice_code) |> 
   summarise(gp_sum_est_all_asian = sum(est_all_asian),
             gp_sum_est_bangladeshi = sum(est_bangladeshi),

### 2023-12-01 TestScript.R

@@ -265,5 +265,55 @@ lsoa_lookup |>
   filter(sum_n>1)
 
 
+##### Clustering tests#####
+library(factoextra)
+library(cluster)
+gp_reg_pat_prac <- tar_read(gp_reg_pat_prac_lsoa) |> as_tibble() |>
+  group_by(practice_code) |>
+  summarise(list_tot= sum(number_of_patients))
+
+gp_lsoa_with_eth_sum <- tar_read(gp_lsoa_with_eth_sum) |> as_tibble()
+
+full_cats <- gp_lsoa_with_eth_sum |>
+  select(practice_code,gp_sum_est_bangladeshi,gp_sum_est_chinese,gp_sum_est_indian,
+           gp_sum_est_pakistani,gp_sum_est_other_asian,gp_sum_est_blk_african,gp_sum_est_blk_caribbean,
+           gp_sum_est_other_blk,gp_sum_est_all_mixed,gp_sum_est_white_british,gp_sum_est_white_irish,
+           gp_sum_est_white_other,gp_sum_est_other_arab,gp_sum_est_other_other,gp_sum_total)
+  
+five_cats <- gp_lsoa_with_eth_sum |>
+  select(practice_code,gp_sum_est_all_asian,gp_sum_est_all_black, gp_sum_est_all_mixed, gp_sum_est_all_white,gp_sum_est_all_other,gp_sum_total2)
+
+full_cats_numeric <- full_cats |>
+  mutate(gp_sum_est_bangladeshi=as.numeric(gp_sum_est_bangladeshi),
+         gp_sum_est_chinese=as.numeric(gp_sum_est_chinese),
+         gp_sum_est_indian=as.numeric(gp_sum_est_indian),
+         gp_sum_est_pakistani=as.numeric(gp_sum_est_pakistani),
+         gp_sum_est_other_asian=as.numeric(gp_sum_est_other_asian),
+         gp_sum_est_blk_african=as.numeric(gp_sum_est_blk_african),
+         gp_sum_est_blk_caribbean=as.numeric(gp_sum_est_blk_caribbean),
+         gp_sum_est_other_blk=as.numeric(gp_sum_est_other_blk),
+         gp_sum_est_all_mixed=as.numeric(gp_sum_est_all_mixed),
+         gp_sum_est_white_british=as.numeric(gp_sum_est_white_british),
+         gp_sum_est_white_irish=as.numeric(gp_sum_est_white_irish),
[…]

### 2023-12-05 clustering_analysis.qmd
Clustering so far written up

---
title: "Cluster Analysis Experiments"
lang: en-GB
author: "Jacqueline Grout"
date: last-modified
date-format: "YYYY-MM-DD"
format:
  html:
    grid:
      sidebar-width: 250px
      body-width: 800px
      margin-width: 250px
      gutter-width: 1.5rem
    embed-resources: true
    smooth-scroll: true
    theme: cosmo
    fontcolor: black
    toc: true
    toc-location: left
    toc-title: Contents
    toc-depth: 3
editor: visual
execute:
  echo: false
  message: false
  warning: false
  freeze: auto
editor_options: 
  chunk_output_type: console
---

##

```{r}
#| echo: false
# Read in the analysis RData file ----
library(factoextra)
library(cluster)
library(dplyr)
library(targets)
load("cluster_analysis_v1.Rdata")
```

## Looking for the "elbow"

### All Categories

```{r}
#| echo: false
full_cats_plot
```

### Five Categories

```{r}
#| echo: false
five_cats_plot
```

## All Categories Gap Plot

Plot to show number of clusters vs. gap statistic for 15 clusters maximum after 50 iterations

```{r}
#| echo: false
full_cats_gap_plot
```

## Five Categories Gap Plot

Plot to show number of clusters vs. gap statistic for 15 clusters maximum after 50 iterations

```{r}
#| echo: false
five_cats_gap_plot
```

## Results of k-medoids model

All categories with 15 clusters

```{r}
#| echo: false
#plot results of final k-medoids model
fviz_cluster(pams_full_cats, data = full_cats)

```

Visualisation of the ethnic groups of the 15 medoid practices :

![](secret/Picture2.png){fig-align="center"}

Five categories with 4 clusters

```{r}
#| echo: false
#plot results of final k-medoids model
fviz_cluster(pams_five_cats, data = five_cats)
```

Visualisation of the ethnic groups of the 4 medoid practices :

![](secret/Picture1.png){fig-align="center"}

Looking at the 15 clusters visualisation for all categories by eye I wondered if 8 clusters made more sense:

Visualisation of the ethnic groups of the 8 medoid practices :

![](secret/Picture3.png){fig-align="center"}

## Next steps?

I wondered about including total list size as a variable?

Should I break the practices up some how by taking out white?

Other clustering methods?

Other ideas?

### 2023-12-05 TestScript.R

@@ -328,17 +328,26 @@ five_cats_plot <- fviz_nbclust(scale_five_cats, pam, method = "wss")
 gap_stat_full_cats <- clusGap(scale_full_cats,
                     FUN = pam,
                     K.max = 15, #max clusters to consider
-                    B = 500) #total bootstrapped iterations
+                    B = 50) #total bootstrapped iterations
 
 gap_stat_five_cats <- clusGap(scale_five_cats,
                               FUN = pam,
                               K.max = 15, #max clusters to consider
-                              B = 500) #total bootstrapped iterations
+                              B = 50) #total bootstrapped iterations
 
 #plot number of clusters vs. gap statistic
 full_cats_gap_plot <- fviz_gap_stat(gap_stat_full_cats)
 five_cats_gap_plot <- fviz_gap_stat(gap_stat_five_cats)
 
+#FULL CATS WITH 15 CLUSTERS
+#make this example reproducible
+set.seed(1)
+#perform k-medoids clustering with k = 15 clusters
+pams_full_cats <- pam(scale_full_cats, 15, metric = 'euclidean', stand = FALSE)
+#view results
+pams_full_cats
+
+#FIVE CATS WITH 4 CLUSTERS
 #make this example reproducible
 set.seed(1)
 #perform k-medoids clustering with k = 4 clusters
@@ -347,21 +356,49 @@ pams_five_cats <- pam(scale_five_cats, 4, metric = 'euclidean', stand = FALSE)
 pams_five_cats
 
 #plot results of final k-medoids model
+fviz_cluster(pams_full_cats, data = full_cats)
 fviz_cluster(pams_five_cats, data = five_cats)
 
 #add cluster assignment to original data
-final_data <- cbind(five_cats, cluster = pams_five_cats$cluster)
+final_data_full_cats <- cbind(full_cats, cluster = pams_full_cats$cluster)
+final_data_five_cats <- cbind(five_cats, cluster = pams_five_cats$cluster)
 
 #view final data
-head(final_data)
+head(final_data_full_cats)
+head(final_data_five_cats)
 
 
-N81115 
-D82048 
-K81026 
-M85058
+#These are the 15 cluster medoids for full cats
+# N81115 
+# D82048 
+# K81026 
+# M85058
+
+plot_data <- full_cats |>   rownames_to_column(var = 'practice_code') |>
+  filter(practice_code %in% c("C81022 ","J82145","G81059","J83053","C81036 ","E85735",
[…]

### 2023-12-05 TestScript.R
Transformed data to % ethnicity and re-examined options for cluster number

@@ -264,141 +264,3 @@ lsoa_lookup |>
   count(lsoa21cd) |>
   summarise(sum_n=sum(n)) |>
   filter(sum_n>1)
-
-
-##### Clustering tests#####
-library(factoextra)
-library(cluster)
-library(dplyr)
-library(targets)
-gp_reg_pat_prac <- tar_read(gp_reg_pat_prac_lsoa) |> as_tibble() |>
-  group_by(practice_code) |>
-  summarise(list_tot= sum(number_of_patients))
-
-gp_lsoa_with_eth_sum <- tar_read(gp_lsoa_with_eth_sum) |> as_tibble()
-
-full_cats <- gp_lsoa_with_eth_sum |>
-  select(practice_code,gp_sum_est_bangladeshi,gp_sum_est_chinese,gp_sum_est_indian,
-           gp_sum_est_pakistani,gp_sum_est_other_asian,gp_sum_est_blk_african,gp_sum_est_blk_caribbean,
-           gp_sum_est_other_blk,gp_sum_est_all_mixed,gp_sum_est_white_british,gp_sum_est_white_irish,
-           gp_sum_est_white_other,gp_sum_est_other_arab,gp_sum_est_other_other)
-  
-five_cats <- gp_lsoa_with_eth_sum |>
-  select(practice_code,gp_sum_est_all_asian,gp_sum_est_all_black, gp_sum_est_all_mixed, gp_sum_est_all_white,gp_sum_est_all_other)
-
-
-full_cats <-
-full_cats |>
-  remove_rownames() |>
-  column_to_rownames(var = 'practice_code')
-
-five_cats <-
-  five_cats |>
-  remove_rownames() |>
-  column_to_rownames(var = 'practice_code')
-
-
-#remove rows with missing values
-full_cats <- na.omit(full_cats)
-five_cats <- na.omit(five_cats)
-
-#scale each variable to have a mean of 0 and sd of 1
-# full_cats <- scale(full_cats)
-# five_cats <- scale(five_cats)
-
-scale_full_cats <- scale(full_cats)
-scale_five_cats <- scale(five_cats)
-
-# v <- prcomp(scale_full_cats)$x[,1]
-# scale_full_cats <- scale_full_cats[order(v),]
-# 
-# v <- prcomp(scale_five_cats)$x[,1]
-# scale_five_cats <- scale_five_cats[order(v),]
-
-head(scale_full_cats)
-head(scale_five_cats)
-
[…]

### 2023-12-20 clustering_analysis.qmd
Transformed data to % ethnicity and re-examined options for cluster number
Matching practices

@@
 library(targets)
+library(GGally)
+library(gt)
+library(tidyverse)
+
 load("cluster_analysis_v1.Rdata")
@@
 
+::: panel-tabset
+## Actual Numbers
+
+```{r}
+#| echo: false
+full_cats_plot
+```
+
+## Percent
+
+```{r}
+#| echo: false
+full_cats_percents_plot
+```
+:::
+
 ### Five Categories
@@
 ```
+
+::: panel-tabset
+## Actual numbers
+
+```{r}
+#| echo: false
+five_cats_plot
+```
+
+## Percent
+
+```{r}
+#| echo: false
+five_cats_percents_plot
+```
+:::
 
@@
 ```
+
+## Correlation Matrix
+
+```{r}
+#Correlation Matrix
+ggcorr(full_cats,
+       method = "pairwise",
+       nbreaks = 10,
+       hjust = 0.9,
+       label = TRUE,
+       label_size = 3,
+       layout.exp = 3)+
+  theme(legend.position = c(0.25,0.65))
+```
+
+## Scree Plots
+
+::: panel-tabset
+## All categories
+
+```{r}
+pca <- princomp(scale_full_cats)
+
+screeplot(pca, npcs=14)
+```
+
+## Excluding White British
+
+```{r}
+pca2 <- princomp(scale_full_cats_percents_ex_whitebritish)
+
+screeplot(pca2, npcs=14)
+```
+:::
 
@@
 
+### Alternative K-Medoid models
+
+#### Experiment with removal of White British completely but leave all practices in the dataset
+
+```{r}
+pams_full_cats_percents_ex_whitebritish8_clusters_plot
+```
+
+```{r}
+final_data_full_cats_percents_ex_whitebritish8_clusters |>
+  rownames_to_column(var = 'practice_code') |>
+  group_by(cluster)|>
+  count("practice_code")|>
+  select(cluster,n) |>
+  gt()
+```
+
+![](secret/Picture4.png){fig-align="center"}
+
+#### Experiment with removal of practice with \> 85% White British manually into their own group
+
+```{r}
+pams_full_cats_percents_ex_85whitebritish7_clusters_plot
+```
+
+![](secret/Picture5.png){fig-align="center"}
+
 ## Next steps?
@@
 
-Should I break the practices up some how by taking out white?
-
 Other clustering methods?

### 2023-12-20 methods.qmd
Test clustering methods
Matching practices

@@
 ### Attempt 1
+
+<https://www.statology.org/k-medoids-in-r/>
+
+https://www.rdocumentation.org/packages/cluster/versions/2.1.6
+
+look at 18:50 in this video: https://www.youtube.com/watch?v=dKXTNxNDpMA
+
+## Mapping new GP practices to old
+
+### To obtain CHD prevalence for as many new practices as possible
+
+```{r}
+library(stringr)
+gp_lsoa_with_eth_sum <- tar_read(gp_lsoa_with_eth_sum) |> as_tibble()
+
+metric1 <- tar_read(metric1) |> as_tibble()
+
+gp_history <- tar_read(gp_history) |> as_tibble()
+
+match_gps <- gp_lsoa_with_eth_sum |> left_join(metric1, join_by(practice_code==AreaCode)) |>
+  mutate(chdprev=case_when(Value>0 ~ Value))
+
+#select just gps with missing data
+missing_data_gps <- match_gps |> 
+  filter(Value<0|is.na(Value)==TRUE) |>
+  select(practice_code)
+
+#get the postcodes
+missing_data_gps <- missing_data_gps |> 
+  left_join(gp_history, join_by(practice_code==org_code))|>
+  select(practice_code,postcode)|>
+  mutate(part_practice_code=substr(postcode,1,str_length(postcode)-1))
+
+#remove active practices from the gp_history and attach the CHD prevalence(metric1)
+gp_history <- gp_history |> 
+  mutate(part_practice_code=substr(postcode,1,str_length(postcode)-1)) |>
+  filter(status_code!="A") |>
+  inner_join(metric1, join_by(org_code==AreaCode))
+
+test <- missing_data_gps |> left_join(gp_history, join_by(part_practice_code==part_practice_code))
+
+```

### 2023-12-20 R/rii_calculations.R

@@ -1,0 +1,186 @@
+
+# Load libraries ----------------------------------------------------------
+library(dplyr)
+library(tidyr)
+library(janitor)
+library(stringr)
+library(purrr)
+library(broom)
+library(glue)
+
+# Load data ---------------------------------------------------------------
+# This data has a row for each IMD decile. The first column is the IMD decile
+# number, the second column is the total GP list size for that decile.
+# Each of the other columns is the total activity for that metric and decile.
+# In this example there are 32 metrics.
+#
+# This test dataset has been created using a random number generator with data
+# items in a pattern, range and with an average similar to the original dataset.
+#
+
+activity_type_decile <-
+  readRDS("data/activity_by_type_decile_stg1_testdata.rds") |>
+  clean_names()
+
+# Metrics 2 - 9 listed as use GP list size 16+ as denominator
+# else uses CHD prevalence
+by_list_total <- c(
+  "02", "05",
+  "06", "07",
+  "08", "09"
+)
+
+# Produces a string to be used in an if(else) in later code to match against
+# the metric numbers listed in by_list_total
+by_list_regex <- glue("^metric({paste(by_list_total, collapse = '|')})_total$")
+
+# Data manipulation code --------------------------------------------------
+activity_long <- activity_type_decile |>
+  pivot_longer(
+    cols = c(starts_with("metric"), -metric01_total),
+    names_to = "metric_name",
+    values_to = "metric_total"
+  ) |>
+  mutate(
+    total_column = ifelse(
+      str_detect(metric_name, by_list_regex),
+      list_size_total,
+      metric01_total
+    ),
+    total_ratio = metric_total / total_column * 1000
+  ) |>
+  select(
+    quantile = gp_im_dquantile,
+    metric01_total,
+    metric_name,
+    metric_total,
+    total_ratio
+  ) |>
+  mutate(activity = (total_ratio * metric01_total / 1000)) |>
+  group_by(metric_name) |>
+  # arrange() required for first() and last() functions to calculate
+  # most_depr_value and least_depr_value
+  arrange(
+    metric_name,
+    quantile
+  ) |>
+  mutate(
+    total_activity = sum(activity),
+    prevalence = sum(metric01_total),
+    overall_value = total_activity / prevalence * 1000,
[…]

### 2023-12-20 _targets.R

@@ -98,7 +98,10 @@ list(
                pivot_wider(names_from=c(ind_code,measure),
                            names_sep="_",
                            values_from = value)), 
-  
+  tar_target(data_path24,"data/20231201_92847.xlsx", format="file"), #CHD synthetic prevalence metric 1
+  tar_target(metric1,read_excel_file(data_path24)), 
+  tar_target(data_path25,"data/epraccur_gp.csv", format="file"), #All practices
+  tar_target(gp_history,read_csv_file(data_path25)), 
   tar_target(metric13,get_my_fingertips_gp_data(273,"2021/22")), # metric 13
   tar_target(metric6,get_my_fingertips_gp_data(93088,"2021/22")), # metric 6
   tar_target(metric7,get_my_fingertips_gp_data(241,"2021/22")), # metric 7

### 2023-12-20 clustering_generate_data.R

@@ -258,6 +258,7 @@ pams_five_cats_percents5_clusters_plot <- fviz_cluster(pams_five_cats_percents, 
 
 #add cluster assignment to original data
 final_data_full_cats_percent_6_clusters <- cbind(full_cats, cluster = pams_full_cats_percents$cluster)
+final_data_five_cats_percent_5_clusters <- cbind(five_cats, cluster = pams_five_cats_percents$cluster)
 
 final_data_full_cats_percent_6_clusters |>
   rownames_to_column(var = 'practice_code') |>
@@ -267,6 +268,10 @@ final_data_full_cats_percent_6_clusters |>
 #These are the 4 cluster medoids for full cats with percents
 full_cats |>   rownames_to_column(var = 'practice_code') |>
   filter(practice_code %in% c("J81626","G81059","H81048","F86018"))
+
+#These are the 5 cluster medoids for full cats with percents
+output_data <- five_cats |>   rownames_to_column(var = 'practice_code') |>
+  filter(practice_code %in% c("P83007","D82034","H82088","K81089","P84669"))
 
 #These are the 6 cluster medoids for full cats with percents
 full_cats |>   rownames_to_column(var = 'practice_code') |>
@@ -284,6 +289,28 @@ ggcorr(full_cats,
        layout.exp = 3)+
   theme(legend.position = c(0.25,0.65))
 
+pca <- princomp(scale_full_cats)
+pca$loadings
+
+loadings <- pca$loadings
+ggplot(data=full_cats,aes(x=gp_sum_est_blk_caribbean, y=gp_sum_est_blk_african)) +
+  geom_point(apha=.3) +
+  stat_ellipse(type='norm',level=.99)+
+  geom_abline(intercept=0,slope=loadings[2,1]/loadings[1,1]) +
+  geom_abline(intercept=0,slope=loadings[2,2]/loadings[1,2])
+
+screeplot(pca, npcs=14)
+
+pca2 <- princomp(scale_full_cats_percents_ex_whitebritish)
+
+pca2$loadings
+
+screeplot(pca2, npcs=14)
+
+
+
+################################################################################
+  
 # Experiment with removal of White British completely but leave all practices in the dataset
 full_cats_percents_ex_whitebritish <- gp_lsoa_with_eth_sum |>
[…]

### 2023-12-20 tests.R

@@ -26,7 +26,9 @@ lsoa_eth_sum <- tar_read(lsoa_eth_sum) |> as_tibble()
 
 lsoa_lookup <- tar_read(lsoa_lookup) |> as_tibble()
 
+gp_lsoa_with_eth_sum <- tar_read(gp_lsoa_with_eth_sum) |> as_tibble()
 
+metric1 <- tar_read(metric1) |> as_tibble()
 
 metric40 <- tar_read(metric40) |> as_tibble()
 

### 2024-01-25 postcode.R
Cleaned up code added comments

@@ -108,51 +108,3 @@ metric1_updated <- joined |>
 still_missing_prev <- full_list |>
   filter(is.na(chd_pre_to_use)==TRUE)
 
-
-
-get_missing_chd_prevalence <- function(metric1,gp_history_short,joined_gp_history_and_chd_prev,gp_list_summary,gp_geocoded){
-  
-  
-  joined <- gp_list_summary |>
-    left_join(joined_gp_history_and_chd_prev) |>
-    #where there is no matching value from orig in the gp list summary then flag these 
-    mutate(has_orig_prev=case_when(Value>=0 ~ 1, .default =  0))
-  
-  orig_with_geocode_and_id <- gp_geocoded |>
-    select(org_code,practice_name,postcode,has_orig_prev,geometry) |>
-    mutate(id=row_number())
-  
-  no_prev <- orig_with_geocode_and_id |>
-    filter(has_orig_prev == 0)
-  
-  with_prev <- orig_with_geocode_and_id |>
-    filter(has_orig_prev == 1)
-  
-  no_prev |>
-    select(no_prev_postcode = postcode) |>
-    st_join(with_prev, st_nearest_feature)
-  
-  no_prev_with_neighbours <-  st_sf(no_prev) |>
-    select(no_prev_postcode = postcode,no_prev_org_code=org_code) |>
-    # finds all the gp's in 1.5 km
-    st_join(st_sf(with_prev), st_is_within_distance, 1500)
-  
-  no_prev_with_neighbours_and_neighbours_prev <- no_prev_with_neighbours |>
-    left_join(joined, join_by(org_code)) |>
-    group_by(no_prev_org_code) |>
-    summarise(prev_est=mean(Value)) |>
-    st_drop_geometry(select(no_prev_org_code,prev_est))
-  
-  #xxxx estimates out of xxx so just xx practices missing *******
-  
-
-  metric1_updated <- joined |>
-    left_join(no_prev_with_neighbours_and_neighbours_prev,join_by(org_code==no_prev_org_code)) |>
-    mutate(chd_prev_to_use = case_when(Value>=0 ~ Value,
-                                       .default = prev_est)
-    ) |>
-    select(org_code,chd_prev_to_use)
-  
-  return(metric1_updated)
-}
-

### 2024-01-27 methods.qmd
Cleaned up code added comments
Adding over 45 flag, more clustering, impute chd prev

--- 2023-12-01 methods.qmd
@@
 
+nearest neighbours to impute chd prevalence
+
 ## Census 2021 Ethnicity by LSOA
@@
 
+Imputing data in Census 2021:
+
+<https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/methodologies/itemeditingandimputationprocessforcensus2021englandandwales>
+
 ## Which date should be used for the gp list size dataset?
@@
 
+Used October 2022 gp list size by age and gender to calculate % of list size aged 45+
+
 ## Clustering
@@
 ### Attempt 1
+
+<https://www.statology.org/k-medoids-in-r/>
+
+https://www.rdocumentation.org/packages/cluster/versions/2.1.6
+
+look at 18:50 in this video: https://www.youtube.com/watch?v=dKXTNxNDpMA
+
+## Mapping new GP practices to old
+
+### To obtain CHD prevalence for as many new practices as possible
+
+link to to the blog
+
+<https://the-strategy-unit.github.io/data_science/blogs/posts/2024-01-17_nearest_neighbour.html>
+
+151 missing and this finds all but 26 using a 1.5km radius - more or less?

### 2024-01-27 R/functions.R

@@ -282,3 +282,46 @@ get_missing_chd_prevalence <- function(metric1,gp_history_short,joined_gp_histor
     select(org_code,chd_prev_to_use)
   return(metric1_updated)
 }
+
+
+# get_full_cats <- function(gp_lsoa_with_eth_sum){
+# full_cats <-gp_lsoa_with_eth_sum |>
+#     select(practice_code,gp_sum_est_bangladeshi,gp_sum_est_chinese,gp_sum_est_indian,
+#            gp_sum_est_pakistani,gp_sum_est_other_asian,gp_sum_est_blk_african,gp_sum_est_blk_caribbean,
+#            gp_sum_est_other_blk,gp_sum_est_all_mixed,gp_sum_est_white_british,gp_sum_est_white_irish,
+#            gp_sum_est_white_other,gp_sum_est_other_arab,gp_sum_est_other_other)
+# return(full_cats)
+# }
+
+get_over45_perc <- function(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)
+{gp_over45_perc <- gp_reg_pat_prac_sing_age_male |>
+    rbind(gp_reg_pat_prac_sing_age_female)|>
+    filter(age != "ALL") |>
+    group_by(org_code,age) |>
+    summarise(number_of_patients=sum(number_of_patients)) |>
+    dplyr::mutate(age=as.numeric(case_when(age=="95+" ~ 95,
+                                           .default = as.numeric(age))
+    )) |>
+    mutate(over_45=case_when(age>= 45 ~1,
+                             .default = 0)) |>
+    group_by(org_code,over_45)|>
+    summarise(number_of_patients=sum(number_of_patients))|>
+    group_by(org_code) |>
+    mutate(total_patients=sum(number_of_patients))|>
+    ungroup() |>
+    mutate(perc=(number_of_patients/total_patients)*100)|>
+    filter(over_45==1) |>
+    select(org_code,perc)|>
+    rename(practice_code=org_code,perc_over45=perc)
+
+
+return(gp_over45_perc)
+
+}
+
+
+join_over45_to_gp_lsoa_with_eth_sum <- function(gp_lsoa_with_eth_sum,gp_over45_perc){
+  gp_lsoa_with_eth_sum_over45perc <- gp_lsoa_with_eth_sum |>
+    left_join(gp_over45_perc)
+  return(gp_lsoa_with_eth_sum_over45perc)
+}

### 2024-01-27 _targets.R

@@ -53,7 +53,15 @@ list(
   tar_target(eth_lsoa_census,read_csv_file(data_path13)),
   tar_target(data_path14, "data/gp-reg-pat-prac-lsoa-all.csv", format = "file"), #GP Reg Pat Prac LSOA October 2022
   tar_target(gp_reg_pat_prac_lsoa,read_csv_file(data_path14)),
-  tar_target(data_path15, "data/LSOA_(2011)_to_LSOA_(2021)_to_Local_Authority_District_(2022)_Lookup_for_England_and_Wales.csv", format = "file"), #lookup
+  tar_target(data_path26, "data/gp-reg-pat-prac-sing-age-female.csv", format = "file"), #GP Reg Pat Prac Females October 2022
+  tar_target(gp_reg_pat_prac_sing_age_female, read_csv_file(data_path26)|>
+                                                select(org_code,age,number_of_patients)), 
+  tar_target(data_path27, "data/gp-reg-pat-prac-sing-age-male.csv", format = "file"), #GP Reg Pat Prac Males October 2022
+  tar_target(gp_reg_pat_prac_sing_age_male,read_csv_file(data_path27)|>
+               select(org_code,age,number_of_patients)), 
+  tar_target(gp_over45_perc,get_over45_perc(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)),
+ 
+   tar_target(data_path15, "data/LSOA_(2011)_to_LSOA_(2021)_to_Local_Authority_District_(2022)_Lookup_for_England_and_Wales.csv", format = "file"), #lookup
   tar_target(lsoa_lookup,read_csv_file(data_path15)),
   tar_target(data_path17,"data/gp-reg-pat-prac-map.csv", format = "file"),
   tar_target(gp_icb_mapping,read_csv_file(data_path17)|> 
@@ -128,6 +136,8 @@ list(
   #and calculate the number for each ethnicity based on ethnicity and list size
   tar_target(gp_lsoa_with_eth,join_gp_and_eth(eth_joined,gp_lsoa)),
   tar_target(gp_lsoa_with_eth_sum,sum_ethnicities(gp_lsoa_with_eth)),
+  #add the perc over 45 on the gp list
+  tar_target(gp_lsoa_with_eth_sum_over45perc,join_over45_to_gp_lsoa_with_eth_sum(gp_lsoa_with_eth_sum,gp_over45_perc)),
   
[…]

### 2024-01-27 clustering_generate_data.R

@@ -5,6 +5,7 @@ library(cluster)
 library(dplyr)
 library(targets)
 library(tidyverse)
+
 gp_reg_pat_prac <- tar_read(gp_reg_pat_prac_lsoa) |> as_tibble() |>
   group_by(practice_code) |>
   summarise(list_tot= sum(number_of_patients))
@@ -571,3 +572,73 @@ hcl$labels
 
 cut <- cutree(hcl, k = 10, h = NULL)
 df_cut <- as.data.frame(cut)
+
+##############################################################################
+
+gp_lsoa_with_eth_sum_over45perc <- tar_read(gp_lsoa_with_eth_sum_over45perc) |> as_tibble()
+
+full_cats <- gp_lsoa_with_eth_sum_over45perc |>
+  select(practice_code,gp_sum_est_bangladeshi,gp_sum_est_chinese,gp_sum_est_indian,
+         gp_sum_est_pakistani,gp_sum_est_other_asian,gp_sum_est_blk_african,gp_sum_est_blk_caribbean,
+         gp_sum_est_other_blk,gp_sum_est_all_mixed,gp_sum_est_white_british,gp_sum_est_white_irish,
+         gp_sum_est_white_other,gp_sum_est_other_arab,gp_sum_est_other_other,perc_over45)
+
+five_cats <- gp_lsoa_with_eth_sum_over45perc |>
+  select(practice_code,gp_sum_est_all_asian,gp_sum_est_all_black, gp_sum_est_all_mixed, 
+         gp_sum_est_all_white,gp_sum_est_all_other,perc_over45)
+
+#Move the practice code into row names
+full_cats <-
+  full_cats |>
+  remove_rownames() |>
+  column_to_rownames(var = 'practice_code')
+
+#Move the practice code into row names
+five_cats <-
+  five_cats |>
+  remove_rownames() |>
+  column_to_rownames(var = 'practice_code')
+
+
+#remove rows with missing values
+full_cats <- na.omit(full_cats)
+five_cats <- na.omit(five_cats)
+
+#scale each variable to have a mean of 0 and sd of 1
+# full_cats <- scale(full_cats)
+# five_cats <- scale(five_cats)
+
+scale_full_cats <- scale(full_cats)
+scale_five_cats <- scale(five_cats)
+
+head(scale_full_cats)
+head(scale_five_cats)
+
[…]

### 2024-01-27 postcode_nearest_neighbours.R

@@ -1,51 +1,0 @@
-library(tidyverse)
-library(sf)
-library(tidygeocoder)
-library(leaflet)
-
-postcodes <- tribble(
-  ~postcode, ~type,
-  "WS7 3AE", "home",
-  "WS7 3XH", "gp",
-  "WS7 3XE", "gp",
-  "WS11 9SE", "gp",
-  "WS7 1AQ", "gp",
-  "WS7 2EY", "gp"
-) |>
-  mutate(id = row_number()) |>
-  geocode(postalcode = postcode) |>
-  st_as_sf(coords = c("long", "lat"), crs = 4326)
-
-# nothing more than a check that the points look valid
-leaflet(postcodes) |>
-  addTiles() |>
-  addMarkers()
-
-
-homes <- postcodes |>
-  filter(type == "home")
-
-gps <- postcodes |>
-  filter(type == "gp")
-
-homes |>
-  select(home_postcode = postcode) |>
-  st_join(gps, st_nearest_feature)
-
-homes |>
-  select(home_postcode = postcode) |>
-  # finds all the gp's in 1.5 km
-  st_join(gps, st_is_within_distance, 1500)
-
-# one thing to note, you will see in both of these examples we end up keeping
-# geometry from home
-
-# something like this can get back the gp geometry
-
-homes |>
-  select(home_postcode = postcode) |>
-  # finds all the gp's in 1.5 km
-  st_join(gps, st_is_within_distance, 1500) |>
-  st_drop_geometry() |>
-  select(id, home_postcode) |>
-  inner_join(x = gps, by = join_by("id")) # note, we flip the order of the join

### 2024-02-02 methods.qmd
Re-organise metrics and remove metrics

@@
 
+Variation in age groups for different ethnicities:
+
+<https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/>
+
+<https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/#download-the-data>
+
 ## Which date should be used for the gp list size dataset?
@@
 151 missing and this finds all but 26 using a 1.5km radius - more or less?
+
+## Disparity
+
+<https://www.health.state.mn.us/data/mchs/pubs/raceethn/rankingbyratio20032007.pdf>

### 2024-02-02 R/functions.R

@@ -10,11 +10,18 @@ read_csv_file <- function(file){
 }
 
 
-#Get Fingertips data
+#Get Fingertips data Value
 get_my_fingertips_gp_data <- function(ind,year){
   Metric <- fingertips_data(IndicatorID = ind,AreaTypeID = "All") |>
     filter(AreaType=="GPs", Timeperiod==year) |>
     select(AreaCode,Value)
+}
+
+#Get Fingertips data count
+get_my_fingertips_gp_data_count <- function(ind,year){
+  Metric <- fingertips_data(IndicatorID = ind,AreaTypeID = "All") |>
+    filter(AreaType=="GPs", Timeperiod==year) |>
+    select(AreaCode,Count)
 }
 
 #Get QOF via Excel file
@@ -310,6 +317,26 @@ return(gp_over45_perc)
 
 }
 
+get_gp_16andover_pop <- function(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)
+{gp_16andover_pop <- gp_reg_pat_prac_sing_age_male |>
+  rbind(gp_reg_pat_prac_sing_age_female)|>
+  filter(age != "ALL") |>
+  group_by(org_code,age) |>
+  summarise(number_of_patients=sum(number_of_patients)) |>
+  dplyr::mutate(age=as.numeric(case_when(age=="95+" ~ 95,
+                                         .default = as.numeric(age))
+  )) |>
+  filter(as.numeric(age) >= 16) |>
+  group_by(org_code) |>
+  summarise(number_of_patients16andover=sum(number_of_patients)) |>
+  ungroup() |>
+  select(org_code,number_of_patients16andover)|>
+  rename(practice_code=org_code)
+
+return(gp_16andover_pop)
+
+}
+
 
 join_over45_to_gp_lsoa_with_eth_sum <- function(gp_lsoa_with_eth_sum,gp_over45_perc){
   gp_lsoa_with_eth_sum_over45perc <- gp_lsoa_with_eth_sum |>
@@ -397,14 +424,15 @@ return(metric23_updated)
 
 
 
-add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,gp_lsoa_with_eth_sum,
+add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,gp_16andover_pop,
                             metric1_updated,metric6,metric7,metric8,metric9,metric11,metric13,metric13b,
                             metric14,metric15,metric16,metric16b,metric17,metric18,metric19,metric20,
[…]

### 2024-02-02 _targets.R

@@ -9,7 +9,7 @@ library(targets)
 tar_option_set(
   packages = c("tibble","fingertipsR","readxl","tidyverse","utils","janitor",
                "readr","visNetwork","odbc","stringr","MLID","sf","tidygeocoder",
-               "cluster","factoextra"), # packages that your targets need to run
+               "cluster","factoextra","purrr","broom","glue"), # packages that your targets need to run
   format = "rds"
 
   # Set other options as needed.
@@ -77,28 +77,23 @@ list(
   tar_target(gp_reg_pat_prac_sing_age_male,read_csv_file(data_path27)|>
                select(org_code,age,number_of_patients)), 
   tar_target(gp_over45_perc,get_over45_perc(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)),
+  tar_target(gp_16andover_pop,get_gp_16andover_pop(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)),
  
-   tar_target(data_path15, "data/LSOA_(2011)_to_LSOA_(2021)_to_Local_Authority_District_(2022)_Lookup_for_England_and_Wales.csv", format = "file"), #lookup
+  tar_target(data_path15, "data/LSOA_(2011)_to_LSOA_(2021)_to_Local_Authority_District_(2022)_Lookup_for_England_and_Wales.csv", format = "file"), #lookup
   tar_target(lsoa_lookup,read_csv_file(data_path15)),
   tar_target(data_path17,"data/gp-reg-pat-prac-map.csv", format = "file"),
   tar_target(gp_icb_mapping,read_csv_file(data_path17)|> 
                select(practice_code,practice_name,sub_icb_location_code, sub_icb_location_name,
                       icb_code,icb_name,comm_region_code,comm_region_name)),  
   
-  tar_target(data_path18,"data/CVDP003.csv", format = "file"),  
-  tar_target(cvdp003,read_csv_file(data_path18)),
-  
-  tar_target(data_path19,"data/CVDP006.csv", format = "file"), 
-  tar_target(cvdp006,read_csv_file(data_path19)), 
-  
-  tar_target(data_path20,"data/CVDP007.csv", format = "file"), 
-  tar_target(cvdp007,read_csv_file(data_path20)),  
   
   tar_target(data_path21,"data/CVDP008.csv", format = "file"), 
-  tar_target(cvdp008,read_csv_file(data_path21)), 
[…]

### 2024-02-02 index_disparity_calcs.R

@@ -1,0 +1,74 @@
+library(glue)
+library(dplyr)
+library(tidyverse)
+
+get_disparity_ratio <- function(activity_by_type_clusters_stg1) {
+  options(scipen=999) # disable scientific notation
+  #options(scipen=0)
+  
+  activity_type_decile <- activity_by_type_clusters_stg1
+  
+  # Metrics 2 - 9 listed as use GP list size 16+ as denominator
+  # else uses CHD prevalence
+  by_list_total <- c(
+    "2", "5",
+    "6", "7",
+    "8", "9"
+  )
+  
+  # Produces a string to be used in an if(else) in later code to match against
+  # the metric numbers listed in by_list_total
+  by_list_regex <- glue("^metric({paste(by_list_total, collapse = '|')})_total$")
+  
+  # Data manipulation code --------------------------------------------------
+  activity_long <- activity_type_decile |>
+    pivot_longer(
+      cols = c(starts_with("metric"), -metric1_total),
+      names_to = "metric_name",
+      values_to = "metric_total"
+    ) |>
+    mutate(
+      total_column = ifelse(
+        str_detect(metric_name, by_list_regex),
+        list_size_total,
+        metric1_total
+      ),
+      total_rate = metric_total / total_column * 1000
+    ) |>
+    select(
+      quantile = cluster,
+      metric1_total,
+      metric_name,
+      metric_total,
+      total_rate
+    ) |>
+    mutate(activity = (total_rate * metric1_total / 1000)) |>
+    group_by(metric_name) |>
+    # arrange() required for min() and max() functions to calculate
+    # largest_value and smallest_value
+    arrange(
+      metric_name,
+      quantile
+    ) |>
+    mutate(
+      total_activity = sum(activity),
+      prevalence = sum(metric1_total),
+      overall_value = total_activity / prevalence * 1000,
+      population = metric1_total,
+      largest_value = max(total_rate),
+      smallest_value = min(total_rate),
+      total_pop = sum(population),
+      proportion_pop = population / total_pop,
+      abs_range = largest_value - smallest_value, # not used in this particular code
+      rel_range = largest_value / smallest_value,
+      disparity_ratio = total_rate / largest_value,
+      abs_diff = abs(total_rate - largest_value)
+    ) |>
+    ungroup() 
+ 
+  
+  return(activity_long)
+}
+
+table_data <- activity_long |>
+  select(cluster=quantile, metric_name, disparity_ratio, largest_value)

### 2024-02-02 to_do_thurs.R

@@ -1,40 +1,0 @@
-library(tidyverse)
-library(targets)
-
-
-# Look at this data
-cvdp003 <- tar_read(cvdp003) |> as_tibble()
-# cvdp006, cvdp007,cvdp008,cvdp009
-
-ncdes_data <- tar_read(ncdes_data) |> as_tibble()
-
-#check each metric for format / value to use etc
-
-
-############################################################################
-
-
-
-final_data_full_cats_percent_over45_5_clusters <- tar_read(final_data_full_cats_percent_over45_5_clusters) |> as_tibble()
-gp_lsoa_with_eth_sum <- tar_read(gp_lsoa_with_eth_sum) |> as_tibble()
-clustered_gp_and_metrics <- tar_read(clustered_gp_and_metrics)|> as_tibble()
-activity_by_type_clusters_stg1 <- tar_read(activity_by_type_clusters_stg1)|> as_tibble()
-
-
-93088
-92588
-
-  Metric <- fingertips_data(IndicatorID = 273,AreaTypeID = "All") 
-  
-  #Get Fingertips data
-  get_my_fingertips_gp_data <- function(ind,year){
-    Metric <- fingertips_data(IndicatorID = ind,AreaTypeID = "All") |>
-      filter(AreaType=="GPs", Timeperiod==year) 
-  }
-  
- metric_ft <- get_my_fingertips_gp_data(241,"2022/23")
-  
-  Metric <- Metric |>
-    filter(AreaType=="GPs") 
-  
-  unique(Metric$AreaType)

### 2024-02-02 clustering_analysis.qmd
Added cluster charts to targets pipeline

@@
 library(tidyverse)
+library(targets)
 
@@
 
-## Next steps?
+## Five clusters using 14 ethnicities
 
-I wondered about including total list size as a variable?
+```{r}
+tar_read(elbow_plot)
+```
 
-Other clustering methods?
+```{r}
+tar_read(cluster_plot)
+```
 
-Other ideas?
+```{r}
+tar_read(cluster2_chart)
+```
+
+```{r}
+tar_read(cluster2_map)
+```
+
+## Five clusters using 14 ethnicities with % pop 45+
+
+```{r}
+tar_read(elbow_plot_over45)
+```
+
+```{r}
+tar_read(cluster_plot_over45)
+```
+
+```{r}
+tar_read(cluster1_chart)
+```
+
+```{r}
+tar_read(cluster1_map)
+```

### 2024-02-02 R/functions.R

@@ -445,20 +445,41 @@ get_scale_full_cats_percents<- function(full_cats_percents){
   return(scale_full_cats_percents)
 }
 
-get_clusters <- function(scale_full_cats_percents,full_cats_percents){
-  #FULL CATS WITH 5 CLUSTERS PERCENT BASED
-  #make this example reproducible
-  set.seed(99)
-  #perform k-medoids clustering with k = 5 clusters
-  pams_full_cats_percents <- pam(scale_full_cats_percents, 5, metric = 'euclidean', stand = FALSE)
+get_clusters <- function(pams_full_cats_percents,full_cats_percents){
   
-  final_data_full_cats_percent_5_clusters <- 
+  
+  final_data <- 
     cbind(full_cats_percents, cluster = pams_full_cats_percents$cluster)|>
-    rename(gp_practice_code=practice_code)
+    rename(gp_practice_code=practice_code)|>
+    mutate(cluster_renum=case_when(cluster==2~1,
+                                    cluster==1~2,
+                                    cluster==3~3,
+                                    cluster==5~4,
+                                    cluster==4~5))|>
+    select(-cluster)|>
+    rename(cluster=cluster_renum)
   
-  return(final_data_full_cats_percent_5_clusters)
+  return(final_data)
 }
 
+get_pams <-  function(scale_full_cats_percents){
+  set.seed(99)
+#perform k-medoids clustering with k = 5 clusters
+pams_full_cats_percents <- pam(scale_full_cats_percents, 5, metric = 'euclidean', stand = FALSE)
+return(pams_full_cats_percents)
+}
+
+get_elbow_plot <- function(scaled_data){
+  
+elbow_plot <- fviz_nbclust(scaled_data, pam, method = "wss")
+return(elbow_plot)
+}
+
+get_cluster_plot <- function(pams_data){
+  #plot results of final k-medoids model
+  cluster_plot <- fviz_cluster(pams_data, data = full_cats)
+  return(cluster_plot)
+}
 
 
 #proccess metric 23 to work out whether los of cabg and pci is below trimpoint
@@ -493,7 +514,7 @@ add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final
                             metric39,metric40){
   clustered_gp_and_metrics <-
[…]

### 2024-02-02 _targets.R

@@ -9,9 +9,10 @@ library(targets)
 tar_option_set(
   packages = c("tibble","fingertipsR","readxl","tidyverse","utils","janitor",
                "readr","visNetwork","odbc","stringr","MLID","sf","tidygeocoder",
-               "cluster","factoextra","purrr","broom","glue"), # packages that your targets need to run
+               "cluster","factoextra","purrr","broom","glue","RColorBrewer","leaflet"), # packages that your targets need to run
   format = "rds"
 
+ 
   # Set other options as needed.
 )
 
@@ -222,11 +223,17 @@ tar_target(metric40, ncdes_data|>
   #cluster the practices 1
   tar_target(full_cats_percents_over45,get_full_cats_percents_over45(gp_lsoa_with_eth_sum_over45perc)),
   tar_target(scale_full_cats_percents_over45,get_scale_full_cats_percents_over45(full_cats_percents_over45)),
-  tar_target(final_data_full_cats_percent_over45_5_clusters,get_clusters(scale_full_cats_percents_over45,full_cats_percents_over45)),
+  tar_target(pams_full_cats_percents_over45,get_pams(scale_full_cats_percents_over45)),
+  tar_target(final_data_full_cats_percent_over45_5_clusters,get_clusters(pams_full_cats_percents_over45,full_cats_percents_over45)),
+  tar_target(elbow_plot_over45,get_elbow_plot(scale_full_cats_percents_over45)),
+  tar_target(cluster_plot_over45,get_cluster_plot(pams_full_cats_percents_over45)),
   #cluster the practices 2
   tar_target(full_cats_percents,get_full_cats_percents(gp_lsoa_with_eth_sum)),
   tar_target(scale_full_cats_percents,get_scale_full_cats_percents(full_cats_percents)),
-  tar_target(final_data_full_cats_percent_5_clusters,get_clusters(scale_full_cats_percents,full_cats_percents)),
+  tar_target(pams_full_cats_percents,get_pams(scale_full_cats_percents)),
+  tar_target(final_data_full_cats_percent_5_clusters,get_clusters(pams_full_cats_percents,full_cats_percents)),
+  tar_target(elbow_plot,get_elbow_plot(scale_full_cats_percents)),
+  tar_target(cluster_plot,get_cluster_plot(pams_full_cats_percents)),
 
 
 #add in the missing metrics*******
@@ -243,6 +250,11 @@ tar_target(clustered_gp_and_metrics,
[…]

### 2024-02-02 clustering_analysis.qmd
Disparity calculations

@@
 ```{r}
+
+final_data_full_cats_percent_5_clusters <- tar_read(final_data_full_cats_percent_5_clusters) |> as_tibble()
+
+final_data_full_cats_percent_5_clusters |>
+  group_by(cluster)|>
+  count("practice_code")|>
+  gt()
+```
+
+```{r}
 tar_read(cluster2_chart)
@@
 ```{r}
+final_data_full_cats_percent_over45_5_clusters <- tar_read(final_data_full_cats_percent_over45_5_clusters) |> as_tibble()
+
+final_data_full_cats_percent_over45_5_clusters |>
+  group_by(cluster)|>
+  count("practice_code")|>
+  gt()
+```
+
+```{r}
 tar_read(cluster1_chart)

### 2024-02-02 R/functions.R

@@ -482,7 +482,7 @@ get_cluster_plot <- function(pams_data){
 }
 
 
-#proccess metric 23 to work out whether los of cabg and pci is below trimpoint
+#process metric 23 to work out whether los of cabg and pci is below trimpoint
 metric23_below_trimpoint <- function(metric23,metric23trimpoints){
   #Join the PCI and CABG spells to the trim point data by HRG, flag those with a spell duration less than relevant trim point
   #(take average spell duration for the HRG where spell duration is missing and remove data where still null spell duration
@@ -574,7 +574,7 @@ add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final
 process_metrics <-function(clustered_gp_and_metrics){
 
 activity_by_type_clusters_stg1<-clustered_gp_and_metrics |>
-  filter(metric1 != "NA") |>
+  filter(metric1 != "NA") |> # removes 26 practices (need to check details of these....)
   mutate(list_size_total = replace_na(list_size, 0)) |>
   mutate(metric1_total = replace_na(metric1, 0)) |>
   #mutate(metric2_total = replace_na(metric2, 0)) |>

### 2024-02-02 backgroundjob.R

@@ -1,4 +1,0 @@
-load("cluster_analysis_v1.Rdata")
-
-
-

### 2024-02-02 index_disparity_calcs.R

@@ -36,13 +36,12 @@ get_disparity_ratio <- function(activity_by_type_clusters_stg1) {
       total_rate = metric_total / total_column * 1000
     ) |>
     select(
-      quantile = cluster,
+      quantile = cluster2,
       metric1_total,
       metric_name,
       metric_total,
       total_rate
     ) |>
-    mutate(activity = (total_rate * metric1_total / 1000)) |>
     group_by(metric_name) |>
     # arrange() required for min() and max() functions to calculate
     # largest_value and smallest_value
@@ -51,14 +50,8 @@ get_disparity_ratio <- function(activity_by_type_clusters_stg1) {
       quantile
     ) |>
     mutate(
-      total_activity = sum(activity),
-      prevalence = sum(metric1_total),
-      overall_value = total_activity / prevalence * 1000,
-      population = metric1_total,
       largest_value = max(total_rate),
       smallest_value = min(total_rate),
-      total_pop = sum(population),
-      proportion_pop = population / total_pop,
       abs_range = largest_value - smallest_value, # not used in this particular code
       rel_range = largest_value / smallest_value,
       disparity_ratio = total_rate / largest_value,
@@ -71,4 +64,4 @@ get_disparity_ratio <- function(activity_by_type_clusters_stg1) {
 }
 
 table_data <- activity_long |>
-  select(cluster=quantile, metric_name, disparity_ratio, largest_value)
+  select(cluster=quantile, metric_name, total_rate,disparity_ratio, largest_value)

### 2024-02-09 R/data_collection_functions.R
Organising functions into sensible groups in a variety of scripts

@@ -1,0 +1,58 @@
+################################################################################
+
+# Functions for various data collection methods
+
+################################################################################
+
+
+#Read an excel data file
+read_excel_file <- function(file){
+  excel_data <- read_excel(file) 
+}
+
+#Read a csv data file
+read_csv_file <- function(file){
+  csv_data <- read_csv(file) |>
+    clean_names()
+}
+
+
+
+#Get Fingertips data Value
+get_my_fingertips_gp_data <- function(ind,year){
+  Metric <- fingertips_data(IndicatorID = ind,AreaTypeID = "All") |>
+    filter(AreaType=="GPs", Timeperiod==year) |>
+    select(AreaCode,Value)
+}
+
+#Get Fingertips data count
+get_my_fingertips_gp_data_count <- function(ind,year){
+  Metric <- fingertips_data(IndicatorID = ind,AreaTypeID = "All") |>
+    filter(AreaType=="GPs", Timeperiod==year) |>
+    select(AreaCode,Count)
+}
+
+#Get QOF via Excel file
+read_qof_excel_file <- function(file){
+  excel_qof_data <- read_xlsx(file,skip=11,.name_repair = "unique_quiet") 
+}
+
+
+
+#Connect to Server and get data
+get_data_via_server <- function(){
+  StrategicWorking_DB <- dbConnect(odbc(), 
+                                   driver = "SQL Server",
+                                   server="MLCSU-BI-SQL",
+                                   database="StrategicWorking")
+  server_metric<- dbFetch(dbSendQuery(StrategicWorking_DB,"SELECT [PRACTICE_CODE]
+,[VALUE]
+FROM [StrategicWorking].[adhoc].[QOF_ACHIEVEMENT_2021_v2]
+where
+indicator_code = 'CHD007'
+and
+measure = 'PCAS'
+order by PRACTICE_CODE")) |>
+    clean_names()
+  #colnames(server_metric)[colnames(server_metric) == 'VALUE'] <- 'Metric32'
+}

### 2024-02-09 R/functions.R

@@ -1,224 +1,8 @@
-#Read an excel data file
-read_excel_file <- function(file){
-  excel_data <- read_excel(file) 
-}
+################################################################################
 
-#Read a csv data file
-read_csv_file <- function(file){
-  csv_data <- read_csv(file) |>
-    clean_names()
-}
+# Other Functions
 
-
-#Get Fingertips data Value
-get_my_fingertips_gp_data <- function(ind,year){
-  Metric <- fingertips_data(IndicatorID = ind,AreaTypeID = "All") |>
-    filter(AreaType=="GPs", Timeperiod==year) |>
-    select(AreaCode,Value)
-}
-
-#Get Fingertips data count
-get_my_fingertips_gp_data_count <- function(ind,year){
-  Metric <- fingertips_data(IndicatorID = ind,AreaTypeID = "All") |>
-    filter(AreaType=="GPs", Timeperiod==year) |>
-    select(AreaCode,Count)
-}
-
-#Get QOF via Excel file
-read_qof_excel_file <- function(file){
-  excel_qof_data <- read_xlsx(file,skip=11,.name_repair = "unique_quiet") 
-}
-
-#Connect to Server and get data
-get_data_via_server <- function(){
-StrategicWorking_DB <- dbConnect(odbc(), 
-                                 driver = "SQL Server",
-                                 server="MLCSU-BI-SQL",
-                                 database="StrategicWorking")
-server_metric<- dbFetch(dbSendQuery(StrategicWorking_DB,"SELECT [PRACTICE_CODE]
-,[VALUE]
-FROM [StrategicWorking].[adhoc].[QOF_ACHIEVEMENT_2021_v2]
-where
-indicator_code = 'CHD007'
-and
-measure = 'PCAS'
-order by PRACTICE_CODE")) |>
-  clean_names()
-#colnames(server_metric)[colnames(server_metric) == 'VALUE'] <- 'Metric32'
-}
-
-# Process Census 2021 ethnicity data to remove wales and organise ethnicities
-process_census21data <- function(data){
-  data |>
-    filter(str_detect(lsoa_code, '^E'))|> 
-    group_by(lsoa_code) |> 
-    summarise(sum_all = sum(total_all_usual_residents),
-              all_asian = sum(asian_asian_british_or_asian_welsh),
-              bangladeshi = sum(asian_asian_british_or_asian_welsh_bangladeshi),
-              chinese = sum(asian_asian_british_or_asian_welsh_chinese),
[…]

### 2024-02-09 visualise_8_clusters.R

@@ -1,0 +1,129 @@
+library(factoextra)
+library(cluster)
+library(dplyr)
+library(targets)
+library(GGally)
+library(gt)
+library(tidyverse)
+library(targets)
+
+load("cluster_analysis_v1.Rdata")
+
+
+View(final_data_full_cats_percents_ex_85whitebritish7_clusters)
+View(final_data_full_cats_percents_ex_whitebritish8_clusters)
+
+pams_full_cats_percents_ex_85whitebritish7_clusters_plot
+
+final_data_full_cats_percents_ex_85whitebritish7_clusters|>
+ # rownames_to_column(var = 'practice_code') |>
+  filter(practice_code=="P85016")
+
+wb <-full_cats_percents |>
+  filter(gp_perc_est_white_british>= 85)
+
+final_data_full_cats_percents_ex_85whitebritish7_clusters
+  
+eight_clusters <- wb |>
+  rownames_to_column(var = 'practice_code') |>
+  select(practice_code)|>
+  mutate(cluster=8)|>
+  rbind(final_data_full_cats_percents_ex_85whitebritish7_clusters|>select(practice_code,cluster))
+
+gp_lsoa_with_eth_sum <- tar_read(gp_lsoa_with_eth_sum)|> as_tibble()
+
+eight_clusters <- eight_clusters |> left_join(gp_lsoa_with_eth_sum)
+eight_full_cats_percents <- eight_full_cats_percents |> left_join(eight_clusters)
+
+eight_full_cats_percents <- get_full_cats_percents(eight_clusters)
+
+final_data_full_cats_percent_8_clusters <- final_data_full_cats_percent_8_clusters |> rename(cluster2=cluster)
+head(eight_clusters)
+View(eight_full_cats_percents)
+
+final_data_full_cats_percent_8_clusters <- get_clusters(eight_clusters,eight_full_cats_percents)
+gp_geocoded <- tar_read(gp_geocoded) |> as_tibble()
+
+eight_full_cats_percents <- eight_full_cats_percents |> rename(gp_practice_code=practice_code)
+
+
+cluster_eight_map <- get_cluster2_map(eight_full_cats_percents,gp_geocoded)
+cluster_eight_chart <- get_cluster2_chart(eight_full_cats_percents)
+
+get_cluster2_chart <- function(clustered_gp_and_metrics){
+  
+  
+  titles <-clustered_gp_and_metrics |> 
+  #  rownames_to_column(var = 'practice_code') |>
+    pivot_longer(
[…]

### 2024-02-20 R/data_collection_functions.R
Adding more metrics to the data collection

@@ -29,7 +29,7 @@ get_my_fingertips_gp_data <- function(ind,year){
 get_my_fingertips_gp_data_count <- function(ind,year){
   Metric <- fingertips_data(IndicatorID = ind,AreaTypeID = "All") |>
     filter(AreaType=="GPs", Timeperiod==year) |>
-    select(AreaCode,Count)
+    select(AreaCode,Count,Denominator)
 }
 
 #Get QOF via Excel file

### 2024-02-20 R/metric_processing_functions.R

@@ -72,22 +72,23 @@ metric23_below_trimpoint <- function(metric23,metric23trimpoints){
 
 add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final_data_full_cats_percent_5_clusters,
                             gp_16andover_pop,
-                            metric1_updated,metric6,metric7,metric8,metric9,metric11,metric13,metric13b,
+                            metric1_updated,metric2,metric5,metric6,metric7,metric8,metric9,metric11,metric12,metric13,metric13b,
                             metric14,metric15,metric16,metric16b,metric17,metric18,metric19,metric20,
                             metric21,metric22,metric23_updated,metric25b,metric27,metric28,
-                            metric29,metric31,metric32,metric34,metric38,
+                            metric29,metric31,metric32,metric33,metric34,metric38,
                             metric39,metric40){
   clustered_gp_and_metrics <-
     final_data_full_cats_percent_over45_5_clusters |> rename(cluster1=cluster)|>
     left_join(final_data_full_cats_percent_5_clusters|>select(gp_practice_code,cluster2=cluster)) |>
     left_join(gp_16andover_pop|>select(gp_practice_code=practice_code,list_size=number_of_patients16andover)) |>
     left_join(metric1_updated)|>
+    mutate(metric1=(metric1/100)*list_size) |>
+    left_join(metric2)|>
     
-    #2 - to do
     #3 - deleted
     #4 - deleted
-    #5 - to do
     
+    left_join(metric5)|>
     left_join(metric6)|>
     mutate(metric6=(metric6/100)*list_size)|>
     left_join(metric7)|>
@@ -96,14 +97,13 @@ add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final
     mutate(metric8=(metric8/100)*list_size)|>
     left_join(metric9)|>    
     mutate(metric9=(metric9/100)*list_size)|>
+    
     #10 - might not be available
     
     left_join(metric11)|>
-    #12 - to do
+    left_join(metric12)|>
     left_join(metric13)|>
-    mutate(metric13=(metric13/100)*list_size)|>
     left_join(metric13b)|>
-    mutate(metric13b=(metric13b/100)*list_size)|>
     left_join(metric14)|>
     left_join(metric15)|>
     left_join(metric16)|>
[…]

### 2024-02-20 _targets.R

@@ -113,9 +113,25 @@ list(
                clean_names()),
   tar_target(metric23_updated,metric23_below_trimpoint(metric23,metric23trimpoints)), #process metric 23
   
+  tar_target(data_path30,"data/gp_survey_smoking_prevalence.xlsx", format="file"), #Smoking synthetic prevalence estimates metric 2
+  tar_target(metric2,read_excel_file(data_path30)|>
+               clean_names()|>
+               select(gp_practice_code=practice_code,metric2_denom=smoking_habits_unweighted_total_responses,percent_occasional_smoker,percent_regular_smoker)|>
+               mutate(metric2_num=(as.numeric(percent_occasional_smoker)+as.numeric(percent_regular_smoker))*metric2_denom)|>
+               select(-percent_occasional_smoker,-percent_regular_smoker)), #metric2
+  
+  tar_target(data_path31,"data/QOFPCASMOK2223.xlsx", format="file"), #Smoking synthetic prevalence estimates metric 12
+  tar_target(metric12,read_excel_file(data_path31)|>
+               clean_names()|>
+               select(gp_practice_code=practice_code,metric12_denom=denominator_plus_pc_as,metric12_num=pc_as)), #metric12
+  
   tar_target(metric13,get_my_fingertips_gp_data(273,"2022/23")|>
                rename(gp_practice_code=AreaCode,
                       metric13=Value)), # metric 13
+  tar_target(metric5,get_my_fingertips_gp_data_count(91280,"2022/23")|>
+               rename(gp_practice_code=AreaCode,
+                      metric5_num=Count,
+                      metric5_denom=Denominator)), # metric 5
   tar_target(metric6,get_my_fingertips_gp_data(92588,"2022/23")|>
                rename(gp_practice_code=AreaCode,
                       metric6=Value)), # metric 6
@@ -149,34 +165,44 @@ list(
 #  rename(gp_practice_code=AreaCode,
 #         metric10=Value)), # metric 10
 
-  tar_target(metric11,get_my_fingertips_gp_data(90619,"2022/23")|>
-  rename(gp_practice_code=AreaCode,
-         metric11=Value)), # metric 11
+  tar_target(metric11,get_my_fingertips_gp_data_count(90619,"2022/23")|>
+                      rename(gp_practice_code=AreaCode,
+                      metric11_num=Count,
+                      metric11_denom=Denominator)), # metric 11
 
[…]

### 2024-02-20 processing_dev.R

@@ -1,0 +1,40 @@
+#process metrics ready to calc
+library(tidyverse)
+library(fingertipsR)
+library(readxl)
+
+clustered_gp_and_metrics <- targets::tar_read(clustered_gp_and_metrics) |> dplyr::as_tibble()
+
+metric12 <- targets::tar_read(metric12) |> dplyr::as_tibble()
+
+#Get Fingertips data count
+get_my_fingertips_gp_data_count <- function(ind,year){
+  Metric <- fingertips_data(IndicatorID = ind,AreaTypeID = "All") |>
+    filter(AreaType=="GPs", Timeperiod==year)
+}
+
+metric_33 <- get_my_fingertips_gp_data_count(91262,"2022/23")
+
+#metric1 - ok
+#metric6 - check which is denom - list or prev - think its list - ok
+#metric7 - check which is denom - list or prev - think its list - ok
+#metric8 - check which is denom - list or prev - think its list - ok
+#metric9 - ok - DQ with practice F85002?
+
+#metric 11 - need to create a numerator and a denominator - currently a %
+#done
+
+#metric 13 - keep 13b and remove 13
+# should be a % then as a prop of list size
+
+#metric 14 - ok
+#metric 15 ok
+
+
+
+metric13_all <- fingertips_data(IndicatorID = 273,AreaTypeID = "All") |>
+    filter(AreaType=="GPs", Timeperiod=="2022/23") 
+
+qof_chd_2223 <- tar_read(qof_chd_2223) |> as_tibble()
+
+# check the metric 13b columns

### 2024-02-23 R/functions.R
Added treemaps to describe clusters and charts per metric showing rates by cluster

@@ -3,6 +3,27 @@
 # Other Functions
 
 ################################################################################
+#get SU colours
+# get_su_colours <- function(){
+#   su_yellow <- '#f9bf07'
+#   su_red <- '#ec6555'
+#   su_blue <- '#5881c1'
+#   su_grey <- '#686f73'
+#   su_black <- '#2c2825'
+#   
+#   su_light_yellow <- '#ffe699'
+#   su_light_red <- '#f4a39a'
+#   su_light_blue <- '#b4c6e6'
+#   su_light_grey <- '#d9d9d9'
+#   
+#   su_colour_list <- list(c(su_yellow,su_red,su_blue,su_grey,su_black,su_light_yellow,su_light_red,su_light_blue,su_light_grey),
+#                          c("su_yellow","su_red","su_blue","su_grey","su_black","su_light_yellow","su_light_red","su_light_blue",
+#                            "su_light_grey")
+#                          )
+#   
+#   return(su_colour_list)
+#   
+# }
 
 
 #join the chd prev to gp history
@@ -31,6 +52,27 @@ get_geocoded_data <- function(gp_list_summary,orig,joined_gp_history_and_chd_pre
 }
 
 
+get_45over_list <- function(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)
+{gp_list_45over <- gp_reg_pat_prac_sing_age_male |>
+  rbind(gp_reg_pat_prac_sing_age_female)|>
+  filter(age != "ALL") |>
+  group_by(org_code,age) |>
+  summarise(number_of_patients=sum(number_of_patients)) |>
+  dplyr::mutate(age=as.numeric(case_when(age=="95+" ~ 95,
+                                         .default = as.numeric(age))
+  )) |>
+  mutate(over_45=case_when(age>= 45 ~1,
+                           .default = 0)) |>
+  group_by(org_code,over_45)|>
+  summarise(number_of_patients=sum(number_of_patients))|>
+  ungroup()|>
+  filter(over_45==1)|>
+  select(practice_code=org_code,number_45over=number_of_patients)
+
+
+return(gp_list_45over)
+
+}
 
 
 

### 2024-02-23 R/metric_processing_functions.R

@@ -79,16 +79,17 @@ combine_metrics17_and_32 <- function(metric32,metric17){
 }
 
 add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final_data_full_cats_percent_5_clusters,
-                            gp_16andover_pop,
+                            gp_16andover_pop,gp_list_45over,
                             metric1_updated,metric2,metric5,metric6,metric7,metric8,metric9,metric11,metric12,metric13,metric13b,
                             metric14,metric15,metric16,metric16b,metric17,metric18,metric19,metric20,
-                            metric21,metric22,metric23_updated,metric25b,metric27,metric28,
-                            metric29,metric31,metric32_updated,metric33,metric34,metric38,
+                            metric21,metric22,metric23_updated,metric25b,metric27,metric28,metric28b,
+                            metric29,metric29b,metric31,metric32_updated,metric33,metric34,metric38,
                             metric39,metric40){
   clustered_gp_and_metrics <-
     final_data_full_cats_percent_over45_5_clusters |> rename(cluster1=cluster)|>
     left_join(final_data_full_cats_percent_5_clusters|>select(gp_practice_code,cluster2=cluster)) |>
     left_join(gp_16andover_pop|>select(gp_practice_code=practice_code,list_size=number_of_patients16andover)) |>
+    left_join(gp_list_45over|>select(gp_practice_code=practice_code,list_size_45=number_45over)) |>
     left_join(metric1_updated)|>
     mutate(metric1=(metric1/100)*list_size) |>
     left_join(metric2)|>
@@ -129,7 +130,9 @@ add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final
     #26 - to do
     left_join(metric27)|>
     left_join(metric28)|>
+    left_join(metric28b)|>
     left_join(metric29)|>
+    left_join(metric29b)|>
     #30 - not available
     left_join(metric31)|>
     left_join(metric32_updated)|>
@@ -151,6 +154,7 @@ process_metrics <-function(clustered_gp_and_metrics){
   activity_by_type_clusters_stg1<-clustered_gp_and_metrics |>
     filter(metric1 != "NA") |> # removes 26 practices (need to check details of these....)
     mutate(list_size_total = replace_na(list_size, 0)) |>
+    mutate(list_size_45_total = replace_na(list_size_45, 0)) |>
     mutate(metric1_total = replace_na(metric1, 0)) |>
[…]

### 2024-02-23 _targets.R

@@ -9,7 +9,7 @@ library(targets)
 tar_option_set(
   packages = c("tibble","fingertipsR","readxl","tidyverse","utils","janitor",
                "readr","visNetwork","odbc","stringr","MLID","sf","tidygeocoder",
-               "cluster","factoextra","purrr","broom","glue","RColorBrewer","leaflet"), # packages that your targets need to run
+               "cluster","factoextra","purrr","broom","glue","RColorBrewer","leaflet","treemapify"), # packages that your targets need to run
   format = "rds"
 
  
@@ -61,9 +61,15 @@ list(
   tar_target(data_path10, "data/metric28.csv", format = "file"), # SUS data metric 28
   tar_target(metric28,read_csv_file(data_path10)|> 
                rename(metric28=number_chd_deaths)),
+  tar_target(data_path32, "data/metric28b.csv", format = "file"), # SUS data metric 28b deaths <75yrs
+  tar_target(metric28b,read_csv_file(data_path32)|>
+               rename(metric28b=number_chd_deaths)),
   tar_target(data_path11, "data/metric29.csv", format = "file"), # SUS data metric 29
   tar_target(metric29,read_csv_file(data_path11)|> 
                rename(metric29=deaths)),
+  tar_target(data_path33, "data/metric29b.csv", format = "file"), # SUS data metric 29b deaths <75yrs
+  tar_target(metric29b,read_csv_file(data_path33)|>
+               rename(metric29b=deaths)),
   
   # tar_target(data_path12, "data/April22GPListLSOA.csv", format = "file"), #GP list size
   # tar_target(gp_list_lsoa,read_csv_file(data_path12)),
@@ -78,6 +84,7 @@ list(
   tar_target(gp_reg_pat_prac_sing_age_male,read_csv_file(data_path27)|>
                select(org_code,age,number_of_patients)), 
   tar_target(gp_over45_perc,get_over45_perc(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)),
+  tar_target(gp_list_45over,get_45over_list(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)),
   tar_target(gp_16andover_pop,get_gp_16andover_pop(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)),
  
[…]

### 2024-02-23 processing_dev.R

@@ -11,6 +11,10 @@ activity_by_type_clusters_stg4<- targets::tar_read(activity_by_type_clusters_stg
 activity_by_type_clusters_stg5<- targets::tar_read(activity_by_type_clusters_stg5) |> dplyr::as_tibble()
 activity_by_type_clusters_stg6<- targets::tar_read(activity_by_type_clusters_stg6) |> dplyr::as_tibble()
 ncdes_data <- tar_read(ncdes_data) |> as_tibble()
+gp_over45_perc <- tar_read(gp_over45_perc) |> as_tibble()
+gp_reg_pat_prac_sing_age_male <- tar_read(gp_reg_pat_prac_sing_age_male) |> as_tibble()
+gp_reg_pat_prac_sing_age_female <- tar_read(gp_reg_pat_prac_sing_age_female) |> as_tibble()
+
 
 
 activity_by_type_clusters_stg7 <- activity_by_type_clusters_stg6 |>
@@ -24,24 +28,48 @@ activity_by_type_clusters_stg7 <- activity_by_type_clusters_stg6 |>
 
 activity_by_type_clusters_stg7 |>
   mutate(metric=fct_rev(factor(metric,levels=c("metric2_rel_iod","metric5_rel_iod","metric6_rel_iod","metric7_rel_iod","metric8_rel_iod",
-                                   "metric9_rel_iod","metric34_rel_iod","metric38_rel_iod","metric11_rel_iod","metric12_rel_iod","metric13_rel_iod","metric33_rel_iod","metric14_rel_iod",
-                                   "metric15_rel_iod","metric16_rel_iod","metric31_rel_iod", "metric17_rel_iod","metric32_rel_iod", 
-                                   "metric39_rel_iod","metric40_rel_iod","metric18_rel_iod",
-                                   "metric19_rel_iod","metric20_rel_iod","metric21_rel_iod", "metric22_rel_iod","metric23_rel_iod",
-                                   "metric25_rel_iod","metric27_rel_iod","metric28_rel_iod", "metric29_rel_iod")))) |>
+                                               "metric33_rel_iod","metric9_rel_iod","metric34_rel_iod","metric38_rel_iod","metric11_rel_iod",
+                                               "metric12_rel_iod","metric13_rel_iod","metric14_rel_iod", "metric15_rel_iod","metric16_rel_iod",
+                                               "metric31_rel_iod", "metric17_rel_iod","metric32_rel_iod", "metric39_rel_iod",
[…]

### 2024-03-04 findings.qmd
Tweak rate charts

---
title: "Inequalities by ethnicity in the CHD pathway"
lang: en-GB
author: "Jacqueline Grout"
date: last-modified
date-format: "YYYY-MM-DD"
format:
  html:
    grid:
      sidebar-width: 250px
      body-width: 800px
      margin-width: 250px
      gutter-width: 1.5rem
    embed-resources: true
    smooth-scroll: true
    theme: cosmo
    fontcolor: black
    toc: true
    toc-location: left
    toc-title: Contents
    toc-depth: 3
editor: visual
execute:
  echo: false
  message: false
  warning: false
  freeze: auto
editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false

library(targets)

library(tidyverse)
```

## Clustered GP Practices

```{r}
tar_read(cluster2_map)
```

### Cluster descriptions

#### Cluster 1 - Least diverse

This cluster is the least diverse of the five and the median percentage of White patients is 97% (94% White British).

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_1)
```

## Fourteen categories

94% White British

```{r}
tar_read(cluster2_eth_chart_1)
```
:::

#### Cluster 2

In this cluster the median percentage of White patients is 93% (87% White British). Those patients whose ethnicity is White are more likely to be White Irish or other White ethnicities (0.7% and 4% respectively) when compared to cluster 1. The median percentage of patients with a mixed ethnicity in this cluster is 2%.

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_2)
```

## Fourteen categories

87% White British

```{r}
tar_read(cluster2_eth_chart_2)
```
:::

#### Cluster 3

In this cluster the median percentage of White patients is 78% (69% White British). The median percentage of patients with a mixed ethnicity in this cluster is 3.8%, which is higher than in cluster 2. Indian patients make up 3.3% of the patients, and 2.6% are patients whose ethnicity is Black African.

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_3)
```

## Fourteen categories

69% White British

```{r}
tar_read(cluster2_eth_chart_3)
```
:::

#### Cluster 4

In cluster 4 just over half the patients are White with a median percentage of 55% (35% White British). The ethnic group of White has fewer White British than clusters 1 to 3, with 17% of patients having an ethnicity of White Other The median percentage of patients whose ethnicity is Black African in this cluster is 8.3%, which is higher than in cluster 3 and the median percentage of Black Caribbean patients is 3.9%. The median percentage of Black and Asian is 29% in this cluster.

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_4)
```

## Fourteen categories

35% White British

```{r}
tar_read(cluster2_eth_chart_4)
```
:::

#### Cluster 5 - Most diverse

This cluster is the most diverse. Overall the median percentage of Asian patients in this cluster is 45%. There are 13.3% Indian, 9.7% Pakistani and 2.9% Bangladeshi. Mixed and other ethnic groups form a large proportion of the patients in this cluster.

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_5)
```

## Fourteen categories

22% White British

```{r}
tar_read(cluster2_eth_chart_5)
```
:::

## Relative Index of Disparity

```{r}
tar_read(rel_iod_chart)
```

## Rate

### Risk factors

```{r}
tar_read(rate_chart_risk_fact)
```

### Risk factor identification

```{r}
tar_read(rate_chart_risk_fact_ident)
```

### Primary prevention

```{r}
tar_read(rate_chart_prim_prevent)

```

### Disease identification

```{r}
tar_read(rate_chart_disease_ident)
```

### Secondary prevention

```{r}
tar_read(rate_chart_second_prevent)
```

### Tertiary prevention

```{r}
tar_read(rate_chart_tert_prevent)
```

### Intermediate outcome

```{r}
tar_read(rate_chart_int_out)
```

### Full outcomes

```{r}
tar_read(rate_chart_full_out)
```

### 2024-03-04 R/chart_functions.R

@@ -1,0 +1,211 @@
+
+get_rel_iod_data_for_chart <- function(activity_by_type_clusters_stg6){
+  chart_iod_data <- activity_by_type_clusters_stg6 |>
+  select(cluster2,list_size_total,ends_with("rel_iod"))|>
+  pivot_longer(cols=starts_with("metric"),
+               names_to="metric")|>
+  filter(cluster2==1)|>
+  select(-cluster2,-list_size_total)
+return(chart_iod_data)
+}
+
+
+get_rel_iod_chart <- function(chart_iod_data){
+
+  rel_iod_chart <- chart_iod_data |>
+  mutate(metric=fct_rev(factor(metric,levels=c("metric2_rel_iod","metric5_rel_iod","metric6_rel_iod","metric7_rel_iod","metric8_rel_iod",
+                                               "metric33_rel_iod","metric9_rel_iod","metric34_rel_iod","metric38_rel_iod","metric11_rel_iod",
+                                               "metric12_rel_iod","metric13_rel_iod","metric14_rel_iod", "metric15_rel_iod","metric16_rel_iod",
+                                               "metric31_rel_iod", "metric17_rel_iod","metric32_rel_iod", "metric39_rel_iod",
+                                               "metric40_rel_iod","metric18_rel_iod", "metric19_rel_iod","metric20_rel_iod","metric21_rel_iod",
+                                               "metric22_rel_iod","metric23_rel_iod", "metric25_rel_iod","metric26_rel_iod","metric27_rel_iod",
+                                               "metric28_rel_iod","metric28b_rel_iod", "metric29_rel_iod","metric29b_rel_iod")))) |>
+ 
+    mutate(metric_name=case_when(metric=="metric2_rel_iod"~"Smoking prev est",
+                               metric=="metric5_rel_iod"~"Smoking register",
+                               metric=="metric6_rel_iod"~"Obesity register",
+                               metric=="metric7_rel_iod"~"Diabetes register",
+                               metric=="metric8_rel_iod"~"Depression register",
+                               metric=="metric33_rel_iod"~"45+ BP check < 5 years",
+                               metric=="metric9_rel_iod"~"CVD risk register",
+                               metric=="metric34_rel_iod"~"CVD pat treated with LLT",
+                               metric=="metric38_rel_iod"~"Pat at risk treated with LLT",
+                               metric=="metric11_rel_iod"~"Smoking cessation support offered",
[…]

### 2024-03-04 R/metric_processing_functions.R

@@ -306,9 +306,74 @@ calc_iod_rate <-function(activity_by_type_clusters_stg1){
            metric34_rate=metric34_num_total/metric1_total,
            metric38_rate=metric38_num_total/metric1_total,
            metric39_rate=metric39_num_total/metric1_total,
-           metric40_rate=metric40_num_total/metric1_total
-           
-           
+           metric40_rate=metric40_num_total/metric1_total,
+           metric2_upper_ci=(metric2_num_total+(sqrt(metric2_num_total*1.96)))*(1/list_size_total),
+           metric2_lower_ci=(metric2_num_total-(sqrt(metric2_num_total*1.96)))*(1/list_size_total),
+           metric5_upper_ci=(metric5_num_total+(sqrt(metric5_num_total*1.96)))*(1/list_size_total),
+           metric5_lower_ci=(metric5_num_total-(sqrt(metric5_num_total*1.96)))*(1/list_size_total),
+           metric6_upper_ci=(metric6_total+(sqrt(metric6_total*1.96)))*(1/list_size_total),
+           metric6_lower_ci=(metric6_total-(sqrt(metric6_total*1.96)))*(1/list_size_total),
+           metric7_upper_ci=(metric7_total+(sqrt(metric7_total*1.96)))*(1/list_size_total),
+           metric7_lower_ci=(metric7_total-(sqrt(metric7_total*1.96)))*(1/list_size_total),
+           metric8_upper_ci=(metric8_total+(sqrt(metric8_total*1.96)))*(1/list_size_total),
+           metric8_lower_ci=(metric8_total-(sqrt(metric8_total*1.96)))*(1/list_size_total),
+           metric9_upper_ci=(metric9_total+(sqrt(metric9_total*1.96)))*(1/list_size_total),
+           metric9_lower_ci=(metric9_total-(sqrt(metric9_total*1.96)))*(1/list_size_total),
+           metric11_upper_ci=(metric11_num_total+(sqrt(metric11_num_total*1.96)))*(1/metric1_total),
+           metric11_lower_ci=(metric11_num_total-(sqrt(metric11_num_total*1.96)))*(1/metric1_total),
+           metric12_upper_ci=(metric12_num_total+(sqrt(metric12_num_total*1.96)))*(1/metric1_total),
+           metric12_lower_ci=(metric12_num_total-(sqrt(metric12_num_total*1.96)))*(1/metric1_total),
[…]

### 2024-03-04 _targets.R

@@ -272,13 +272,13 @@ tar_target(metric40, ncdes_data|>
   tar_target(clusters_for_nacr,get_clusters_for_nacr(clustered_gp_and_metrics,gp_icb_mapping)),
 
   #cluster the practices by region
-  tar_target(full_cats_percents_region,get_full_cats_percents(gp_lsoa_with_eth_sum_region)),
-  tar_target(scale_full_cats_percents_region,get_scale_full_cats_percents(full_cats_percents_region)),
-  tar_target(pams_full_cats_percents_region,get_pams(scale_full_cats_percents_region)),
-  tar_target(final_data_full_cats_percent_5_clusters_region,get_clusters(pams_full_cats_percents_region,full_cats_percents_region)|>
-                                                            rename(cluster2=cluster)),
-  tar_target(elbow_plot_region,get_elbow_plot(scale_full_cats_percents_region)),
-  tar_target(cluster_plot_region,get_cluster_plot(pams_full_cats_percents_region)),
+#  tar_target(full_cats_percents_region,get_full_cats_percents(gp_lsoa_with_eth_sum_region)),
+ # tar_target(scale_full_cats_percents_region,get_scale_full_cats_percents(full_cats_percents_region)),
+#  tar_target(pams_full_cats_percents_region,get_pams(scale_full_cats_percents_region)),
+ # tar_target(final_data_full_cats_percent_5_clusters_region,get_clusters(pams_full_cats_percents_region,full_cats_percents_region)|>
+  #                                                          rename(cluster2=cluster)),
+ # tar_target(elbow_plot_region,get_elbow_plot(scale_full_cats_percents_region)),
+ # tar_target(cluster_plot_region,get_cluster_plot(pams_full_cats_percents_region)),
 
 
 #add in the missing metrics*******
@@ -296,7 +296,6 @@ tar_target(clustered_gp_and_metrics,
                            metric32_updated,metric33,metric34,metric38,
                            metric39,metric40)),
 #plot clusters
-#tar_target(cluster2_map,get_cluster2_map(clustered_gp_and_metrics,gp_geocoded)),
 tar_target(cluster1_map,get_cluster1_map(clustered_gp_and_metrics,gp_geocoded)),
[…]

### 2024-03-04 arrows.R

@@ -1,0 +1,52 @@
+rate_chart_data <- tar_read(rate_chart_data)|>as_tibble()
+activity_by_type_clusters_stg6 <- tar_read(activity_by_type_clusters_stg6)|>as_tibble()
+rate_chart <- tar_read(rate_chart)
+
+
+
+geom_segment(
+  x = 1, y = 1,
+  xend = 4, yend = 7,
+  lineend = "round", # See available arrow types in example above
+  linejoin = "round",
+  size = 2, 
+  arrow = arrow(length = unit(0.3, "inches")),
+  colour = "#EC7014" # Also accepts "red", "blue' etc
+)
+
+palette <- brewer.pal(5, "Set1")
+intercept_data <- rate_chart_data |>
+  select(metric_name,global_rate,pathway)|>
+  unique()|>
+  filter(pathway=="Full outcomes")
+
+arrow_data <- rate_chart_data |> 
+
+   mutate(x_start=cluster2,
+         y_start=1.1*global_rate,
+         x_end=cluster2,
+         y_end=0.9*rate)|>
+  select(-rate,-global_rate,-upper_ci,-lower_ci)
+
+
+rate_chart <- rate_chart_data|>
+  filter(pathway=="Full outcomes")|>
+  ggplot(aes(x=as.factor(cluster2),y=rate, fill=as.factor(cluster2) )) +  
+  geom_bar(stat = "identity") +
+  geom_errorbar( aes(x=as.factor(cluster2), ymin=lower_ci, ymax=upper_ci), width=0.4, colour="grey", alpha=0.9, size=0.5) +
+  scale_fill_manual(values = palette)+
+  theme(legend.position="none")+
+  xlab("Cluster (1=Least diverse, 5=Most diverse)") +
+  ylab("Rate")+
+  facet_wrap(vars(metric_name),ncol=3,scales="free_y")+
+  geom_hline(aes(yintercept=global_rate),intercept_data)+
+  geom_segment(aes(arrow_data),x = x_start, y = y_start,
+               xend = x_end, yend = y_end,
+               lineend = "round", # See available arrow types in example above
+               linejoin = "round",
+               size = 2, 
+               arrow = arrow(length = unit(0.3, "inches")),
+               colour = "grey",
+               inherit.aes=FALSE
+               )
+  

### 2024-03-04 worked_example.R

@@ -1,0 +1,28 @@
+one_metric_example_data <- tar_read(activity_by_type_clusters_stg6) |>
+  as_tibble()|>
+  select(cluster2,metric1_total,starts_with("metric29b"))|>
+  rename(denominator=metric1_total,numerator=metric29b_total)
+
+
+
+
+# Function to compute index of disparity
+compute_id <- function(one_metric_example_data) {
+  # Compute index of disparity
+  data <- one_metric_example_data|>
+    mutate(upper_ci=(numerator+(sqrt(numerator)*1.96))*(1/denominator),
+           lower_ci=(numerator-(sqrt(numerator)*1.96))*(1/denominator),
+           se=(numerator-upper_ci)/1.96,
+           sd=(numerator-mean(numerator))
+           rate=numerator/denominator,
+           global_rate=sum(numerator)/sum(denominator),
+           diff = rate-global_rate,
+           abs_diff=abs(diff*denominator))
+  iod_abs <- sum(data$abs_diff)/2
+  iod_rel <- sum(data$abs_diff)/(2*sum(data$numerator))
+  return(iod_rel)
+}
+
+
+
+rnorm()

### 2024-03-06 findings.qmd
Finesse iod and ci chart

@@
 ```
+
+## Introduction
+
+In the earlier report ..........
+
+## CHD Metrics
+
+jhjkhhjhkj
 
@@
 
+```{r}
+tar_read(ci_iod_chart)
+```
+
 ## Rate
@@
 ```
+
+## Appendix
+
+### Definitions and data sources of pathway metrics
+
+### Methods explained
+
+#### CHD synthetic prevalence estimates
+
+CHD synthetic prevalence estimates data has been removed from Fingertips by OHID.
+
+OHID have sent the archived data. This is for 7564 practices from 2015, of which 6297 are current practices.
+
+Need to identify practices with missing data and work out how to get a prevalence for them.
+
+Also have file from previous project - checked data is the same but need to check practices included - maybe there are more?
+
+nearest neighbours to impute chd prevalence
+
+#### Census 2021 Ethnicity by LSOA
+
+#### Which date should be used for the gp list size dataset?

### 2024-03-06 R/chart_functions.R

@@ -119,24 +119,73 @@ get_ci_iod_chart <- function(iod_with_ci){
                                                                      "BP reading < 140/90 (<80 yrs CHD)","Readmit<30 days of PCI / CABG",
                                                                      "Emergency admissions for CHD","CHD Hospital Deaths","CHD Hospital Deaths <75", 
                                                                      "CHD Deaths","CHD Deaths <75")))) |>
+    mutate(pathway_level = case_when(metric_name == 'metric2' ~ 'Risk',
+                                 metric_name == 'metric5' ~ 'Risk identif.',
+                                 metric_name == 'metric6' ~ 'Risk identif.',
+                                 metric_name == 'metric7' ~ 'Risk identif.',
+                                 metric_name == 'metric8' ~ 'Risk identif.',
+                                 metric_name == 'metric33' ~ 'Risk identif.',  
+                                 metric_name == 'metric9' ~ 'Risk identif.',
+                                 metric_name == 'metric34' ~ '1°c prevent',
+                                 metric_name == 'metric38' ~ '1°c prevent',
+                                 metric_name == 'metric11' ~ '1°c prevent', 
+                                 metric_name == 'metric12' ~ '1°c prevent', 
+                                 metric_name == 'metric13' ~ 'Dis. ident',
+                                 metric_name == 'metric14' ~ 'Dis. ident',
+                                 metric_name == 'metric15' ~ 'Dis. ident',
+                                 metric_name == 'metric16' ~ '2°c prevent',
+                                 metric_name == 'metric31' ~ '2°c prevent',
+                                 metric_name == 'metric17' ~ '2°c prevent',
+                                 metric_name == 'metric32' ~ '2°c prevent',
+                                 metric_name == 'metric39' ~ '2°c prevent',
+                                 metric_name == 'metric40' ~ '2°c prevent',
+                                 metric_name == 'metric18' ~ '2°c prevent',
+                                 metric_name == 'metric19' ~ '2°c prevent',
+                                 metric_name == 'metric20' ~ '3°c prevent',
+                                 metric_name == 'metric21' ~ '3°c prevent',
+                                 metric_name == 'metric22' ~ '3°c prevent',                              
[…]

### 2024-03-06 _targets.R

@@ -485,14 +485,6 @@ tar_target(rate_chart_tert_prevent,get_rate_chart(rate_chart_data,"Tertiary prev
 tar_target(rate_chart_int_out,get_rate_chart(rate_chart_data,"Intermediate outcome")),
 tar_target(rate_chart_full_out,get_rate_chart(rate_chart_data,"Full outcomes"))
 
-
-
-
-#######################################################################################
-#calculate Disparity Ratio
-#tar_target(activity_long,get_disparity_ratio(activity_by_type_clusters_stg1))
-
-
 )
 
 

### 2024-03-07 findings.qmd
Add functions to create IoD calcs and plots by region and ICB

@@
     grid:
-      sidebar-width: 250px
-      body-width: 800px
-      margin-width: 250px
+      sidebar-width: 200px
+      body-width: 950px
+      margin-width: 150px
       gutter-width: 1.5rem
@@
 library(tidyverse)
+library(grid)
+library(gridExtra)
+library(cowplot)
 ```
@@
 ```{r}
-tar_read(rel_iod_chart)
-```
-
-```{r}
 tar_read(ci_iod_chart)
@@
 
+## Regions - Relative Index of Disparity
+
+### London
+
+```{r}
+regional_charts <- tar_read(regional_charts)
+regional_charts[1]
+```
+
+### South West
+
+```{r}
+
+regional_charts[2]
+```
+
+### South East
+
+```{r}
+
+regional_charts[3]
+```
+
+### Midlands
+
+```{r}
+regional_charts[4]
+```
+
+### East of England
+
+```{r}
+regional_charts[5]
+```
+
+### North West
+
+```{r}
+regional_charts[6]
+```
+
+### North East and Yorkshire
+
+```{r}
+regional_charts[7]
+```
+
+```{r}
+
+```
+
 ## Appendix
@@
 #### Which date should be used for the gp list size dataset?
+
+### ICBs- Relative Index of Disparity
+
+```{r}
+icb_charts <- tar_read(icb_charts)
+icb_charts[1]
+icb_charts[2]
+icb_charts[3]
+icb_charts[4]
+icb_charts[5]
+icb_charts[6]
+icb_charts[7]
+icb_charts[8]
+icb_charts[9]
+icb_charts[10]
+icb_charts[11]
+icb_charts[12]
+```

### 2024-03-07 R/chart_functions.R

@@ -4,8 +4,8 @@ get_rel_iod_data_for_chart <- function(activity_by_type_clusters_stg6){
   select(cluster2,list_size_total,ends_with("rel_iod"))|>
   pivot_longer(cols=starts_with("metric"),
                names_to="metric")|>
-  filter(cluster2==1)|>
-  select(-cluster2,-list_size_total)
+  select(-cluster2,-list_size_total)|>
+  dplyr::distinct(metric,value)
 return(chart_iod_data)
 }
 
@@ -191,6 +191,7 @@ get_ci_iod_chart <- function(iod_with_ci){
 
 
 get_rate_chart_data <- function(activity_by_type_clusters_stg6){
+  options(scipen=999)
   rate_data <- activity_by_type_clusters_stg6 |>
   select(cluster2,ends_with("_rate"),ends_with("ci"))|>
   select(-ends_with("global_rate"),-ends_with("diff_rate"))|>

### 2024-03-07 R/iod_calc_functions.R

@@ -96,7 +96,7 @@ get_rates_per_cluster <- function(one_metric_data){
 }
 
 get_ci_iod_together <- function(rates_per_cluster,met_name,metric_data){
-  iod_with_ci <- as_tibble(quantile(rates_per_cluster$iod_rel, c(.05, .95)) )|>
+  iod_with_ci <- as_tibble(quantile(rates_per_cluster$iod_rel, c(.025, .975)) )|>
     mutate(item_number=row_number(),
            metric_name=met_name,
            item_name=case_when(item_number==1~"lower_ci",
@@ -109,5 +109,4 @@ get_ci_iod_together <- function(rates_per_cluster,met_name,metric_data){
     pivot_wider(values_from="value",
                 names_from=item_name)
   return(iod_with_ci)
-  
 }

### 2024-03-07 _targets.R

@@ -19,6 +19,7 @@ tar_option_set(
 # tar_make_clustermq() is an older (pre-{crew}) way to do distributed computing
 # in {targets}, and its configuration for your machine is below.
 options(clustermq.scheduler = "multiprocess")
+options(scipen=999)
 
 
 # Run the R scripts in the R/ folder with your custom functions:
@@ -483,7 +484,36 @@ tar_target(rate_chart_disease_ident,get_rate_chart(rate_chart_data,"Disease iden
 tar_target(rate_chart_second_prevent,get_rate_chart(rate_chart_data,"Secondary prevention")),
 tar_target(rate_chart_tert_prevent,get_rate_chart(rate_chart_data,"Tertiary prevention")),
 tar_target(rate_chart_int_out,get_rate_chart(rate_chart_data,"Intermediate outcome")),
-tar_target(rate_chart_full_out,get_rate_chart(rate_chart_data,"Full outcomes"))
+tar_target(rate_chart_full_out,get_rate_chart(rate_chart_data,"Full outcomes")),
+
+
+#Regional versions
+tar_target(region_clustered_gp_and_metrics,clustered_gp_and_metrics|>
+              left_join(gp_icb_mapping, join_by(gp_practice_code==practice_code))),
+tar_target(regional_charts,region_clustered_gp_and_metrics |>
+             split(region_clustered_gp_and_metrics$comm_region_code) |>
+             map(\(df) process_metrics(df) )|>
+             map(\(df) calc_iod_rate(df) )|>
+             map(\(df) calc_iod_global_rate(df) )|>
+             map(\(df) calc_iod_diff_rate(df) )|>
+             map(\(df) calc_iod_diff(df) )|>
+             map(\(df) calc_abs_iod(df) )|>
+             map(\(df) get_rel_iod_data_for_chart(df) )|>
+             map(\(df) get_rel_iod_chart(df) )),
+
+#ICB versions
+tar_target(icb_clustered_gp_and_metrics,clustered_gp_and_metrics|>
+             left_join(gp_icb_mapping, join_by(gp_practice_code==practice_code))),
+tar_target(icb_charts,icb_clustered_gp_and_metrics |>
+             split(icb_clustered_gp_and_metrics$icb_code) |>
+             map(\(df) process_metrics(df) )|>
+             map(\(df) calc_iod_rate(df) )|>
+             map(\(df) calc_iod_global_rate(df) )|>
[…]

### 2024-03-08 _targets.R
Tabset the regional plots and swith to use of group_split() from split()

@@ -491,7 +491,7 @@ tar_target(rate_chart_full_out,get_rate_chart(rate_chart_data,"Full outcomes")),
 tar_target(region_clustered_gp_and_metrics,clustered_gp_and_metrics|>
               left_join(gp_icb_mapping, join_by(gp_practice_code==practice_code))),
 tar_target(regional_charts,region_clustered_gp_and_metrics |>
-             split(region_clustered_gp_and_metrics$comm_region_code) |>
+             group_split(comm_region_code)|>
              map(\(df) process_metrics(df) )|>
              map(\(df) calc_iod_rate(df) )|>
              map(\(df) calc_iod_global_rate(df) )|>
@@ -505,7 +505,8 @@ tar_target(regional_charts,region_clustered_gp_and_metrics |>
 tar_target(icb_clustered_gp_and_metrics,clustered_gp_and_metrics|>
              left_join(gp_icb_mapping, join_by(gp_practice_code==practice_code))),
 tar_target(icb_charts,icb_clustered_gp_and_metrics |>
-             split(icb_clustered_gp_and_metrics$icb_code) |>
+             group_split(icb_name)|>
+             #split(icb_clustered_gp_and_metrics$icb_code) |>
              map(\(df) process_metrics(df) )|>
              map(\(df) calc_iod_rate(df) )|>
              map(\(df) calc_iod_global_rate(df) )|>

### 2024-03-11 findings.qmd
Tabset the regional plots and swith to use of group_split() from split()
Report writing

--- 2024-03-06 findings.qmd
@@
 ---
-title: "Inequalities by ethnicity in the CHD pathway"
+title: "Visualising CHD Pathway Inequalities by ethnicity - draft"
 lang: en-GB
@@
   html:
+    self-contained: true
     grid:
-      sidebar-width: 250px
-      body-width: 800px
-      margin-width: 250px
+      sidebar-width: 200px
+      body-width: 950px
+      margin-width: 150px
       gutter-width: 1.5rem
@@
 library(targets)
-
+library(gt)
 library(tidyverse)
+library(grid)
+library(gridExtra)
+library(cowplot)
 ```
 
-## Introduction
+# Introduction
 
-In the earlier report ..........
+In 2022, the Strategy Unit worked with the British Heart Foundation (BHF) to explore ways of visualising socio-economic inequalities as they emerge through the coronary heart disease (CHD) pathway.
+We made these visualisations available so that health and care staff can understand where on the pathway socio-economic inequalities emerge and at which points they are moderated or exacerbated.
+The report [^1]and web-based tool[^2] are available via the BHF and Strategy Unit websites.
 
-## CHD Metrics
+[^1]: <https://www.strategyunitwm.nhs.uk/publications/socio-economic-inequalities-coronary-heart-disease>
 
-jhjkhhjhkj
+[^2]: <https://www.bhf.org.uk/icb-tool>
+
+Whilst this initial analysis was focused on inequalities between socio-economic groups, BHF asked the Strategy Unit if we might explore the feasibility of assessing inequalities across other dimensions of inequality.
+This report explores inequalities by ethnicity along the coronary heart disease (CHD) pathway.
+
+The objectives of the report are to:
+
+1.  Set out the methods by which the Strategy Unit have sought to represent CHD pathway inequalities by ethnicity.
+
+2.  Quantify and illustrate the inequalities by ethnicity over the disease progression and treatment pathway for CHD
+
+The analysis has been conducted by the Strategy Unit on behalf of the British Heart Foundation.
+
+# Inequities in healthcare
+
+The term ‘inequities’ is used to describe unjustifiable differences in rates of access between
+subgroups. An equity analysis must control for levels of need within each population subgroup.
+Having done this, an equitable distribution of services is one where rates of access to a service or
+population follow the distribution of need, such that a patient with a given level of need in one
+subgroup has the same chance of accessing a service as their counterparts with a similar level of
+need in other subgroups. This is the standard that the NHS seeks to achieve. Assessing equity is
+challenging. Further detail about inequalities and inequities in healthcare can be found in a recent
+Strategy Unit report[^3]
+
+[^3]: <https://www.strategyunitwm.nhs.uk/publications/socio-economic-inequalities-access-planned-hospital-care-causes-and-consequences>
+
+In our previous work, visualising socio-economic inequalities, our units of analysis were GP practices.
+The metrics of interest, risk factors, primary prevention interventions, secondary care procedures, outcomes etc, are readily available at GP practice level and there are established methods of assigning GP practices to deciles of deprivation.
+Having calculated a metric value for each decile of GP practices, we used the relative index of inequality (RII) to estimate the scale and direction of inequality.
+The RII can only be used when the dimension of inequality can be expressed as a set of order groups.
+
+Assessing inequalities across ethnic groups is more challenging, since unlike deprivation, ethnicity cannot be expresseed as a set of ordered groups.
+The distribution of patients across ethnic groups is unique to each practice.
+There is no meaningful way to numerically aggregate these distributions into a single variable (e.g.
+% BME) without significant loss of information.
+An alternative method of grouping practices according to the distribution of its patients over ethnic group was required.
+We have used k-medoids clustering to assign practices to one of a small number of groups such that practices within a group have similar distributions of patients over ethnic groups.
+K-medoids is an established, and commonly used unsupervised machine learning technique.
+This clustering approach required data on the ethnicity of a practices registered population.
+We imputed it based on the Lower Super Output Area (LSOA) of residence of GP registrants and the ethnicity distribution of patients living in these LSOAs using the Census 2021 data.
+
+Ethnicity, unlike deprivation, is not an ordered variable.
+The approach in our previous work used the relative index of inequality to measure the degree of inequality.
+This measure relies on the ordered quality of the socio-economic deprivation variable.
+To handle the categorical nature of the ethnicity variable we have used the relative index of disparity to indicate the extent to which the rate of an activity or event varies across groups.
+The index estimates the proportion of events (e.g., admissions) that would need to be redistributed between clusters in order that event rates follow levels of need.
+Further detailed explanations of the methods used in the analysis can be found in the appendix.
+
+# CHD Metrics
+
+This report quantifies and illustrates levels of inequity across 34 metrics at various points along the
+continuum of coronary heart disease progression and over a typical treatment pathway. They are
+shown in the table below grouped by domain (risk factors, risk factor identification, primary
+prevention, disease identification, secondary prevention, tertiary prevention, intermediate and full
+outcomes), which represent the various stages along the pathway. Full definitions and data sources
+for each metric are included in the appendix.
+
+*Table 1 - coronary heart disease pathway metrics*
+
+```{r}
+library(gt)
+metrics_table <- tibble(metric_domain = c("Need","Risk factors","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Primary prevention","Primary prevention","Primary prevention","Primary prevention","Primary prevention","Disease identification","Disease identification","Disease identification","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Intermediate outcome","Intermediate outcome","Intermediate outcome","Full outcomes","Full outcomes","Full outcomes","Full outcomes"),
+  name_metric=c("CHD synthetic prevalence estimates",
+                                      "Smoking synthetic prevalence estimates",
+                                      "Smoking register",
+                                      "Obesity register",
+                                      "Diabetes register",
+                                      "Depression register",
+                                      "CVD risk register",
+                                      "Percentage of patients aged 18 and over with GP recorded CVD (narrow definition), who are currently treated with lipid lowering therapy",
+                                      "Percentage of patients aged 18 and over, with no GP recorded CVD and a GP recorded QRISK score of 10% or more, CKD (G3a to G5), T1 diabetes (aged 40 and over) or T2 diabetes aged 60 and over, who are currently treated with lipid lowering therapy.",
+                                      "Smoking cessation support offered",
+                                      "Exception reporting for Smoking cessation support offered",
+                                      "The percentage of patients aged 45 or over who have a record of blood pressure in the preceding 5 years",
+                                      "CHD register",
+                                      "CT angiography",
+                                      "Electrocardiography",
+                                      "Aspirin, anti-platelet or anti-coagulent",
+                                      "Exception reporting for Aspirin, anti-platelet or anti-coagulent",
+                                      "Flu vaccination",
+                                      "Exception reporting for Flu vaccination",
+                                      "Percentage of patients aged 65 or over
+who received a seasonal influenza vaccination between 1 September 2022 and 31 March 2023",
+                                      " Percentage of patients aged 18 to 64
+years and in a clinical at-risk group who received
+a seasonal influenza vaccination between 1
+September 2022 and 31 March 2023",
+"Referral to cardiology (First outpatient)",
+                                      "Cardiology outpatient DNAs",
+"Elective PCI","Elective CABG","Waiting time for elective PCI / CABG","Elective PCI / CABG patients discharged before trimpoint",
+"Cardiac rehabilitation - Started", "Cardiac rehabilitation - Completed","BP reading < 140/90","Readmission within 30 days of elective PCI / CABG","Emergency admissions for CHD","Deaths in hospital from CHD","Deaths in hospital from CHD <75","Deaths from CHD","Deaths from CHD <75"
+                                      )
+                        )
+
+metrics_table |>
+    gt()|>
+  cols_hide(metric_domain)|>
+ # tab_stubhead(label = md("**Domain**"))|>
+    tab_header(
+    title = md("**Table 1** - *Coronary heart disease pathway metrics*")
+  )|>
+   tab_row_group(
+     id="Need",
+    label = md("**Need**"),
+    rows = 1
+  ) |>
+  tab_row_group(
+    id="Risk factors",
+    label = md("**Risk factors**"),
+    rows = 2
+  ) |>
+  tab_row_group(
+    id="Risk factor identification",
+    label = md("**Risk factor identification**"),
+    rows = 3:7
+  )|>
+  tab_row_group(
+    id="Primary prevention",
+    label = md("**Primary prevention**"),
+    rows = 8:12
+  )|>
+  tab_row_group(
+    id="Disease identification",
+    label = md("**Disease identification**"),
+    rows = 13:15
+  )|>
+  tab_row_group(
+    id="Secondary prevention",
+    label = md("**Secondary prevention**"),
+    rows = 16:23
+  )|>
+    tab_row_group(
+    id="Tertiary prevention",
+    label = md("**Tertiary prevention**"),
+    rows = 24:29
+  )|>
+    tab_row_group(
+    id="Intermediate outcome",
+    label = md("**Intermediate outcome**"),
+    rows = 30:32
+  )|>
+    tab_row_group(
+    id="Full outcomes",
+    label = md("**Full outcomes**"),
+    rows = 33:36
+  )|>
+  row_group_order(groups=c("Need","Risk factors","Risk factor identification","Primary prevention","Disease identification","Secondary prevention","Tertiary prevention","Intermediate outcome","Full outcomes"))
+
+#  cols_label(name_metric=md("**Metric**"))
+
+    #domain_metric=md("**Domain**")
+```
+
+# Clustering
+
+Combining List size by lsoa with 2021 census by ethnicity
 
@@
 
-### Cluster descriptions
+## Cluster descriptions
 
@@
 
-## Relative Index of Disparity
+# Rate of activity
 
-```{r}
-tar_read(rel_iod_chart)
-```
+KEY FINDINGS
 
-```{r}
-tar_read(ci_iod_chart)
-```
-
-## Rate
+The charts below show the rates by cluster for each metric along the CHD pathway.
+The horizontal line is the overall rate.
+Explain how its calculated and how to interpret.
+Mention that confidence ontervals are also shown on the charts.
 
@@
 
-## Appendix
+# Relative Index of Disparity
 
-### Definitions and data sources of pathway metrics
+KEY FINDINGS
 
-### Methods explained
+```{r} tar_read(ci_iod_chart)}
+```
 
-#### CHD synthetic prevalence estimates
+::: {.callout-tip icon="false"}
+## Confidence Intervals
+
+Confidence intervals are shown in grey. There is 95% confidence that the index of disparity is within this range.
+:::
+
+Calc of iod inc bootstrap of confidence intervals
+
+# Regional Analysis
+
+## Clusters
+
+The charts below show the distribution of the patients in each region according to the cluster of their GP practice.
+
+::: panel-tabset
+## London
+
+```{r}
+#1
+region_cluster_charts <- tar_read(region_cluster_charts)
+region_cluster_charts[2]
+```
+
+## South West
+
+```{r}
+#2
+region_cluster_charts[7]
+```
+
+## South East
+
+```{r}
+#3
+region_cluster_charts[6]
+```
+
+## Midlands
+
+```{r}
+#4
+region_cluster_charts[3]
+```
+
+## East of England
+
+```{r}
+#5
+region_cluster_charts[1]
+```
+
+## North West
+
+```{r}
+#6
+region_cluster_charts[5]
+```
+
+## North East and Yorkshire
+
+```{r}
+#7
+region_cluster_charts[4]
+```
+:::
+
+## Relative Index of Disparity
+
+::: panel-tabset
+## London
+
+```{r}
+regional_charts <- tar_read(regional_charts)
+regional_charts[1]
+```
+
+## South West
+
+```{r}
+regional_charts[2]
+```
+
+## South East
+
+```{r}
+regional_charts[3]
+```
+
+## Midlands
+
+```{r}
+regional_charts[4]
+```
+
+## East of England
+
+```{r}
+regional_charts[5]
+```
+
+## North West
+
+```{r}
+regional_charts[6]
+```
+
+## North East and Yorkshire
+
+```{r}
+regional_charts[7]
+```
+:::
+
+Mention ICB charts are in appendix and refer to ICBs of interest under relevant region, hyperlink to appendix?? CIs of ICB????? should the CI be reduced?
+
+# Conclusions
+
+# Appendix
+
+## Definitions and data sources of pathway metrics
+
+The table below sets out the sources of the various metrics as well as the time-period to which they relate, and details of the selection criteria used.
+
+| Domain                     | Metric                                 | Data Source | Year    | Definition and selection criteria/codes                                                 |
+|-------------|---------------------|-----------|--------|-------------------|
+| Need                       | CHD synthetic prevalence estimates     | PHE         | 2015    | Provided directly by OHID (formerly indicator 92847 in FingertipsR)                     |
+| Risk factors               | Smoking synthetic prevalence estimates | GP survey   | 2023    |                                                                                         |
+| Risk factor identification | Smoking register                       | NHSD QOF    | 2022/23 | [Indicator 91280 extracted fromFingertipsR](https://fingertips.phe.org.uk/search/91280) |
+|                            | Obesity register                       | NHSD QOF    |         | [Indicator 93088 extracted fromFingertipsR](https://fingertips.phe.org.uk/search/93088) |
+|                            | Diabetes register                      | NHSD QOF    |         |                                                                                         |
+|                            | Depression register                    | NHSD QOF    |         |                                                                                         |
+| Primary prevention         |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+|                            |                                        |             |         |                                                                                         |
+| Full Outcomes              | Deaths in hospital from CHD            |             |         |                                                                                         |
+|                            | Deaths in hospital from CHD \< 75      |             |         |                                                                                         |
+|                            | Deaths from CHD                        |             |         |                                                                                         |
+|                            | Deaths from CHD \<75                   |             |         |                                                                                         |
+
+## Methods explained
+
+### QOF data
+
+### SUS data
+
+### Cardiac rehabilitation
+
+### Readmission within 30 days
+
+### ONS Death records
+
+### CHD synthetic prevalence estimates
 
@@
 
+<https://the-strategy-unit.github.io/data_science/blogs/posts/2024-01-17_nearest_neighbour.html>
+
 #### Census 2021 Ethnicity by LSOA
@@
 #### Which date should be used for the gp list size dataset?
+
+### K-Medoids Cluster
+
+### Index of Disparity calculations
+
+### Confidence Intervals
+
+Explain how these were done - bootstrap method etc
+
+## Integrated Care Board (ICB) Analysis
+
+#### Relative Index of Disparity
+
+```{r}
+icb_charts <- tar_read(icb_charts)
+icb_charts[1]
+icb_charts[2]
+icb_charts[3]
+icb_charts[4]
+icb_charts[5]
+icb_charts[6]
+icb_charts[7]
+icb_charts[8]
+icb_charts[9]
+icb_charts[10]
+icb_charts[11]
+icb_charts[12]
+icb_charts[13]
+icb_charts[14]
+icb_charts[15]
+icb_charts[16]
+icb_charts[17]
+icb_charts[18]
+icb_charts[19]
+icb_charts[20]
+icb_charts[21]
+icb_charts[22]
+icb_charts[23]
+icb_charts[24]
+icb_charts[25]
+icb_charts[26]
+icb_charts[27]
+icb_charts[28]
+icb_charts[23]
+icb_charts[29]
+icb_charts[30]
+icb_charts[31]
+icb_charts[32]
+icb_charts[33]
+icb_charts[34]
+icb_charts[35]
+icb_charts[36]
+icb_charts[37]
+icb_charts[38]
+icb_charts[39]
+icb_charts[40]
+icb_charts[41]
+icb_charts[42]
+```
+
+### References
+
+Confidence intervals vs statistical significance
+
+<https://academic.oup.com/jrssig/article/16/4/20/7038025>

### 2024-03-11 R/chart_functions.R

@@ -13,66 +13,112 @@ return(chart_iod_data)
 get_rel_iod_chart <- function(chart_iod_data){
 
   rel_iod_chart <- chart_iod_data |>
-  mutate(metric=fct_rev(factor(metric,levels=c("metric2_rel_iod","metric5_rel_iod","metric6_rel_iod","metric7_rel_iod","metric8_rel_iod",
-                                               "metric33_rel_iod","metric9_rel_iod","metric34_rel_iod","metric38_rel_iod","metric11_rel_iod",
-                                               "metric12_rel_iod","metric13_rel_iod","metric14_rel_iod", "metric15_rel_iod","metric16_rel_iod",
-                                               "metric31_rel_iod", "metric17_rel_iod","metric32_rel_iod", "metric39_rel_iod",
-                                               "metric40_rel_iod","metric18_rel_iod", "metric19_rel_iod","metric20_rel_iod","metric21_rel_iod",
-                                               "metric22_rel_iod","metric23_rel_iod", "metric25_rel_iod","metric26_rel_iod","metric27_rel_iod",
-                                               "metric28_rel_iod","metric28b_rel_iod", "metric29_rel_iod","metric29b_rel_iod")))) |>
- 
-    mutate(metric_name=case_when(metric=="metric2_rel_iod"~"Smoking prev est",
-                               metric=="metric5_rel_iod"~"Smoking register",
-                               metric=="metric6_rel_iod"~"Obesity register",
-                               metric=="metric7_rel_iod"~"Diabetes register",
-                               metric=="metric8_rel_iod"~"Depression register",
-                               metric=="metric33_rel_iod"~"45+ BP check < 5 years",
-                               metric=="metric9_rel_iod"~"CVD risk register",
-                               metric=="metric34_rel_iod"~"CVD pat treated with LLT",
-                               metric=="metric38_rel_iod"~"Pat at risk treated with LLT",
-                               metric=="metric11_rel_iod"~"Smoking cessation support offered",
-                               metric=="metric12_rel_iod"~"Except rep. smoke cease sup",
-                               metric=="metric13_rel_iod"~"CHD register",
-                               metric=="metric14_rel_iod"~"CT angiography",
-                               metric=="metric15_rel_iod"~"Electrocardiography",
-                               metric=="metric16_rel_iod"~"Asprin anti-platelet etc",
-                               metric=="metric31_rel_iod"~"Except rep. asprin etc",
[…]

### 2024-03-12 findings.qmd
Further details added to the report

@@
 ---
-title: "Visualising CHD Pathway Inequalities by ethnicity - draft"
+title: "Visualising CHD Pathway Inequalities by ethnicity - DRAFT"
 lang: en-GB
@@
 date-format: "YYYY-MM-DD"
+title-block-banner: "#f9bf07"
+title-block-banner-color: "#333739"
 format:
@@
   chunk_output_type: console
+
+css: styles.css
 ---
@@
 library(gridExtra)
-library(cowplot)
+
 ```
@@
 We made these visualisations available so that health and care staff can understand where on the pathway socio-economic inequalities emerge and at which points they are moderated or exacerbated.
-The report [^1]and web-based tool[^2] are available via the BHF and Strategy Unit websites.
+The report[^1] and web-based tool[^2] are available via the BHF and Strategy Unit websites.
 
@@
 
-2.  Quantify and illustrate the inequalities by ethnicity over the disease progression and treatment pathway for CHD
+2.  Quantify and illustrate the inequalities by ethnicity over the disease progression and treatment pathway for CHD.
 
@@
 
-The term ‘inequities’ is used to describe unjustifiable differences in rates of access between
-subgroups. An equity analysis must control for levels of need within each population subgroup.
-Having done this, an equitable distribution of services is one where rates of access to a service or
-population follow the distribution of need, such that a patient with a given level of need in one
-subgroup has the same chance of accessing a service as their counterparts with a similar level of
-need in other subgroups. This is the standard that the NHS seeks to achieve. Assessing equity is
-challenging. Further detail about inequalities and inequities in healthcare can be found in a recent
-Strategy Unit report[^3]
+The term ‘inequities’ is used to describe unjustifiable differences in rates of access between subgroups.
+An equity analysis must control for levels of need within each population subgroup.
+Having done this, an equitable distribution of services is one where rates of access to a service or population follow the distribution of need, such that a patient with a given level of need in one subgroup has the same chance of accessing a service as their counterparts with a similar level of need in other subgroups.
+This is the standard that the NHS seeks to achieve.
+Assessing equity is challenging.
+Further detail about inequalities and inequities in healthcare can be found in a previous Strategy Unit report[^3]
 
@@
 
-This report quantifies and illustrates levels of inequity across 34 metrics at various points along the
-continuum of coronary heart disease progression and over a typical treatment pathway. They are
-shown in the table below grouped by domain (risk factors, risk factor identification, primary
-prevention, disease identification, secondary prevention, tertiary prevention, intermediate and full
-outcomes), which represent the various stages along the pathway. Full definitions and data sources
-for each metric are included in the appendix.
+This report quantifies and illustrates levels of inequity across 34 metrics at various points along the continuum of coronary heart disease progression and over a typical treatment pathway.
+They are shown in the table below grouped by domain (risk factors, risk factor identification, primary prevention, disease identification, secondary prevention, tertiary prevention, intermediate and full outcomes), which represent the various stages along the pathway.
+Full definitions and data sources for each metric are included in the appendix.
 
@@
   )|>
-  row_group_order(groups=c("Need","Risk factors","Risk factor identification","Primary prevention","Disease identification","Secondary prevention","Tertiary prevention","Intermediate outcome","Full outcomes"))
+  row_group_order(groups=c("Need","Risk factors","Risk factor identification","Primary prevention","Disease identification","Secondary prevention","Tertiary prevention","Intermediate outcome","Full outcomes"))|>
+  cols_label(name_metric="")
 
-#  cols_label(name_metric=md("**Metric**"))
-
-    #domain_metric=md("**Domain**")
 ```
@@
 
-Combining List size by lsoa with 2021 census by ethnicity
+In this analysis we used the k-medoids clustering method to group practices.
+Each practice was assigned to one of a small number of groups such that practices within the group have similar distributions of patients over ethnic groups.
+In order to obtain data on the ethnicity of a practice's registered population we imputed it based on the Lower Super Output Area (LSOA) of residence of GP registrants and the ethnicity distribution of patients living in those LSOAs using the Census 2021 data.
+Further detailed explanations of the method can be found in the appendix.
 
-## Clustered GP Practices
-
-```{r}
-tar_read(cluster2_map)
-```
+The clustering analysis resulted in 5 groups (clusters) of practices which are illustrated in the charts and map below.
 
@@
 
-#### Cluster 1 - Least diverse
+The median percentage of each ethnic group for each cluster was calculated and is presented in these charts to describe their respective diversity.
+
+### Cluster 1 - Least diverse
 
@@
 
-#### Cluster 2
+### Cluster 2
 
@@
 
-#### Cluster 3
+### Cluster 3
 
@@
 
-#### Cluster 4
+### Cluster 4
 
@@
 
-#### Cluster 5 - Most diverse
+### Cluster 5 - Most diverse
 
@@
 
+## Clustered GP Practices
+
+The map below presents each GP practice coloured according to their assigned cluster. Cluster 1 (red) is the least diverse and cluster 5 (orange) is the most diverse.
+
+```{r}
+tar_read(cluster2_map)
+```
+
 # Rate of activity
@@
 Explain how its calculated and how to interpret.
-Mention that confidence ontervals are also shown on the charts.
+Mention that confidence intervals are also shown on the charts.
 
@@
 
-```{r} tar_read(ci_iod_chart)}
+and explain what it means and what it is showing
+
+```{r}
+tar_read(ci_iod_chart)
 ```
@@
 
-Confidence intervals are shown in grey. There is 95% confidence that the index of disparity is within this range.
+Confidence intervals are shown in grey. There is 95% confidence that the index of disparity is within this range. The confidence interval estimation methods are detailed in the appendix.
 :::
-
-Calc of iod inc bootstrap of confidence intervals
 
@@
 
-The charts below show the distribution of the patients in each region according to the cluster of their GP practice.
+The charts below show the distribution of the patients in each NHS region according to the cluster of their GP practice.
 
@@
 
-Mention ICB charts are in appendix and refer to ICBs of interest under relevant region, hyperlink to appendix?? CIs of ICB????? should the CI be reduced?
+refer to ICBs of interest under relevant region
+
+The appendix contains equivalent charts to those above, showing the Index of Disparity for each CHD pathway metric for each of the 42 ICBs.
 
@@
 | Domain                     | Metric                                 | Data Source | Year    | Definition and selection criteria/codes                                                 |
-|-------------|---------------------|-----------|--------|-------------------|
+|---------------|---------------|---------------|---------------|---------------|
 | Need                       | CHD synthetic prevalence estimates     | PHE         | 2015    | Provided directly by OHID (formerly indicator 92847 in FingertipsR)                     |
@@
 
+Metrics taken from the QOF data via Fingertips have been extracted by GP practice and the counts
+(numerator of the measure) from the calculations have been used.
+Exception reporting metrics have used the Personalised Care Adjustment (PCA) data to obtain the
+number of patients by GP practice for whom a PCA has been recorded for the relevant metric.
+Possible reasons for a PCA include; Newly diagnosed/registered, Intervention is clinically
+unsuitable, patient choice, did not respond to offers of care, service not available.
+
 ### SUS data
+
+SUS data was extracted using Transact-SQL and using relevant OPCS, ICD10, HRG, and Treatment
+function codes as detailed in table 2 above.
 
@@
 
+This data was supplied by NACR. The data wasn’t available by GP Practice, so an alternative method
+was followed.
+Practices were allocated to a cluster and data was then supplied aggregated into clusters.
+In some cases, the data had to be redacted due to small numbers in the category.
+For the purposes of this analysis the redacted data item was reassigned an average value.
+The true definition of eligible for rehabilitation is the HES admission and successful discharge of all
+patients with CVD e.g., MI, revascularization, or HF. However, to link with the NACR only those
+records entered into the database can be used so NACR use pseudo eligible as this is as close as
+the audit can get, at a granular level, to eligible. It is though affected by areas without entry to the
+database,
+The data used relates to the 2023 calendar year, which is the most recent complete year, and
+following changes in the way rehabilitation has been delivered due to the Covid pandemic, most
+closely reflects the rehabilitation model now in place.
+
 ### Readmission within 30 days
+
+The methodology used for this metric follows, as far as possible, that detailed by NHS Digital[^4]. The
+readmissions counted relate to those who had an elective PCI / CABG in the period 1/3/2022 to
+28/2/2023.
+
+[^4]: <https://digital.nhs.uk/data-and-information/publications/statistical/ccg-outcomes-indicator-set/specifications/3.2-emergency-readmissions-within-30-days-of-discharge-from-hospital_1_4>
 
@@
 
+ONS death records contain an encrypted HES ID which was used to link to outpatient, inpatient and
+A&E HES records, from which the latest GP practice recorded (on a spine traced record) for the
+patient was then assigned to the death record.
+
 ### CHD synthetic prevalence estimates
 
-CHD synthetic prevalence estimates data has been removed from Fingertips by OHID.
+Public Health England (PHE) CHD synthetic prevalence estimates data was previously available from OHID via Fingertips.
+This data has now been removed from Fingertips and archived, however OHID supplied it via email for this project following a request.
+This data is for 7564 practices from 2015, of which 6297 are current practices.
+In order to assign a prevalence for practices with missing data a nearest neighbours methodology was used to impute the prevalence by taking an average of the prevalence of each practice within a 1.5km radius of the practice without data.
+Further details, including code with sample data, for this methodology are available in a blog post on the Strategy Unit Data Science website[^5].
 
-OHID have sent the archived data. This is for 7564 practices from 2015, of which 6297 are current practices.
+[^5]: <https://the-strategy-unit.github.io/data_science/blogs/posts/2024-01-17_nearest_neighbour.html>
 
-Need to identify practices with missing data and work out how to get a prevalence for them.
+### Ethnicity by GP practice
 
-Also have file from previous project - checked data is the same but need to check practices included - maybe there are more?
+Ethnicity by GP practice was imputed based on the Lower Super Output Area (LSOA) of residence of GP registrants and the ethnicity distribution of patients living in these LSOAs using the Census 2021 data.
 
-nearest neighbours to impute chd prevalence
+#### GP list size by LSOA
 
-<https://the-strategy-unit.github.io/data_science/blogs/posts/2024-01-17_nearest_neighbour.html>
+Since metric data is mainly from the 2022/23 year, the October 2022 list size data was taken from the NHS Digital Patients Registered at a GP practice dataset[^6] as this is the data closest to the mid point in the 22/23 year.
+The GP to ICB mapping file was also downloaded and joined to the list size to easily create analysis by ICB and region.
+
+[^6]: <https://digital.nhs.uk/data-and-information/publications/statistical/patients-registered-at-a-gp-practice/october-2022>
 
@@
 
-#### Which date should be used for the gp list size dataset?
+xxxxxxxxxxxxxx
 
-### K-Medoids Cluster
+explain how 2011 lsoa and 2021 lsoa mis match was handled
+
+### K-Medoids Clustering
+
+K-medoids is an established, and commonly used unsupervised machine learning technique.
+This clustering approach required data on the ethnicity of a practices registered population which was obtained as per the methodology explained above (Ethnicity by GP practice).
+The data used in the clustering was the % of patients in each ethnic group per practice, such that for each practice all the ethnic group %'s added to 100%.
+
+The K-medoids clusterimg was performed using the {cluster} package[^7] in R using partitioning (clustering) of the data into k clusters "around medoids" (PAM), which is a more robust version of K-means.
+
+[^7]: Maechler M, Rousseeuw P, Struyf A, Hubert M, Hornik K (2023). *cluster: Cluster Analysis Basics and Extensions*.
+R package version 2.1.6 — For new features, see the 'NEWS' and the 'Changelog' file in the package source), [https://CRAN.R-project.org/package=cluster](https://cran.r-project.org/package=cluster).
+
+Clustering experiments were conducted using a variety of different combinations of ethnic breakdowns, such as 14 ethnic groups and 5 ethnic groups, as well as variations that excluded the % of white british patients or factored in age group.
+Techniques were also used to attempt to determine the optimal number of clusters such as the elbow method, gap plots, and scree plots.
+Consideration was given to combining some ethnic groups and in doing so a correlation matrix was produced to identify ethnic groups that commonly featured together.
+It was also important to seek a number of clusters that could be defined easily, where characteristics did not overlap each other so that resulting clusters were too similar to each other.
+It was also necessary to ensure that there was a sufficient number of practices in each cluster, too few practices in a cluster could result in issues with small amounts of metric data when calculating activity rates and consequently wide confidence intervals in the index of disparity calculations.
+
+In conclusion, the final version selected and presented here used the percentage of the GP list size in each of the 14 census 2021 ethnic groups and five clusters was felt to give a sufficient number of describable clusters for which the index of disparity calculations could be performed.
+The chart below is an elbow plot generated for the 14 ethnic groups showing an elbow around 4 or 5 clusters.
+
+```{r}
+tar_read(elbow_plot)
+```
+
+Table 3 shows the number of GP practices assigned to each of the 5 clusters and the characteristics of the clusters are descirbed above in the main body of the report.
+
+```{r}
+final_data_full_cats_percent_5_clusters <- tar_read(final_data_full_cats_percent_5_clusters) |> as_tibble()
+
+final_data_full_cats_percent_5_clusters |>
+  group_by(cluster)|>
+  count("practice_code")|>
+  ungroup() |>
+  select(cluster,n)|>
+  gt()|>
+    tab_header(title = md("**Table 3** - *Number of practices per cluster*"))|>
+    cols_label(cluster=md("**Cluster**"),
+             n=md("**Number of practices**"))
+```
 
@@
 
+remember...how , detail - allowing for NEED and how this is done
+
 ### Confidence Intervals
 
-Explain how these were done - bootstrap method etc
+Confidence intervals of 95% are presented in the national charts presented in this report.
+Confidence intervals have been used, rather than calculations of statistical significance as this presents a clearer picture of the likelihood of the disparity and due to concerns with the suitability in this case of statistical significance.
+
+> "Practitioners need to be aware that statistical significance differs from practical importance in that statistical significance is highly dependent upon sample size.
+For a large sample, a statistically significant result is likely to be obtained even when the actual magnitude of an effect is small and of little or no practical importance.
+On the other hand, for a small sample, it is quite likely that insufficient evidence of a statistically significant result will be obtained – even when there is, indeed, an effect of practical importance." (Hahn, G.
+J., Doganaksoy, N., Meeker, W.
+Q., 2019)[^8]
+
+[^8]: <https://academic.oup.com/jrssig/article/16/4/20/7038025>
+
+In order to estimate the confidence intervals for the index of disparity, a bootstrap method was adopted, assuming independence, generating 1000 random samples for each metric for which an index of disparity was calculated.
+The 2.5th and 97.5th percentiles were then taken for each metric to give the upper and lower limits of the 95% confidence interval for each.
+
+An explanation of bootstrap theory and confidence intervals is available in Practical Statistics for Data Scientists (Bruce, P. and Bruce, A., 2017)[^9]
+
+[^9]: <https://www.oreilly.com/library/view/practical-statistics-for/9781491952955/>
 
@@
 
-### References
+## R code on Github
 
-Confidence intervals vs statistical significance
+All the code used in this project is available on the Strategy Unit GitHub[^10].
 
-<https://academic.oup.com/jrssig/article/16/4/20/7038025>
+[^10]: <https://github.com/The-Strategy-Unit/bhf_chd-ethnicity>

### 2024-03-12 findings_draft.qmd
Further details added to the report

--- 2024-03-07 findings.qmd
@@
+self-contained: true
+
 ---
-title: "Inequalities by ethnicity in the CHD pathway"
+title: "Inequalities by ethnicity in the CHD pathway - DRAFT"
 lang: en-GB
@@
   html:
+    self-contained: true
     grid:
@@
 ```
-
-## Introduction
-
-In the earlier report ..........
-
-## CHD Metrics
-
-jhjkhhjhkj
 
@@
 
+::: {.callout-tip icon="false"}
+## Confidence Intervals
+
+Confidence intervals are shown in grey. Please disregard the last 4 metrics (related to patient deaths) at this stage.
+:::
+
 ## Rate
+
+The charts below show the rates by cluster for each metric along the CHD pathway. The horizontal line is the overall rate.
 
@@
 
-## Regions - Relative Index of Disparity
+## Regions
 
-### London
+### Clusters
+
+The charts below show the distribution of the patients in each region according to the cluster of their GP practice.
+
+::: panel-tabset
+## London
+
+```{r}
+#1
+region_cluster_charts <- tar_read(region_cluster_charts)
+region_cluster_charts[2]
+```
+
+## South West
+
+```{r}
+#2
+region_cluster_charts[7]
+```
+
+## South East
+
+```{r}
+#3
+region_cluster_charts[6]
+```
+
+## Midlands
+
+```{r}
+#4
+region_cluster_charts[3]
+```
+
+## East of England
+
+```{r}
+#5
+region_cluster_charts[1]
+```
+
+## North West
+
+```{r}
+#6
+region_cluster_charts[5]
+```
+
+## North East and Yorkshire
+
+```{r}
+#7
+region_cluster_charts[4]
+```
+:::
+
+### Relative Index of Disparity
+
+::: panel-tabset
+## London
 
@@
 
-### South West
+## South West
 
 ```{r}
-
 regional_charts[2]
@@
 
-### South East
+## South East
 
 ```{r}
-
 regional_charts[3]
@@
 
-### Midlands
+## Midlands
 
@@
 
-### East of England
+## East of England
 
@@
 
-### North West
+## North West
 
@@
 
-### North East and Yorkshire
+## North East and Yorkshire
 
@@
 ```
-
-```{r}
-
-```
+:::
 
@@
 
-### Definitions and data sources of pathway metrics
+### Integrated Care Boards (ICBs)
 
-### Methods explained
-
-#### CHD synthetic prevalence estimates
-
-CHD synthetic prevalence estimates data has been removed from Fingertips by OHID.
-
-OHID have sent the archived data. This is for 7564 practices from 2015, of which 6297 are current practices.
-
-Need to identify practices with missing data and work out how to get a prevalence for them.
-
-Also have file from previous project - checked data is the same but need to check practices included - maybe there are more?
-
-nearest neighbours to impute chd prevalence
-
-#### Census 2021 Ethnicity by LSOA
-
-#### Which date should be used for the gp list size dataset?
-
-### ICBs- Relative Index of Disparity
+#### Relative Index of Disparity
 
@@
 icb_charts[12]
+icb_charts[13]
+icb_charts[14]
+icb_charts[15]
+icb_charts[16]
+icb_charts[17]
+icb_charts[18]
+icb_charts[19]
+icb_charts[20]
+icb_charts[21]
+icb_charts[22]
+icb_charts[23]
+icb_charts[24]
+icb_charts[25]
+icb_charts[26]
+icb_charts[27]
+icb_charts[28]
+icb_charts[23]
+icb_charts[29]
+icb_charts[30]
+icb_charts[31]
+icb_charts[32]
+icb_charts[33]
+icb_charts[34]
+icb_charts[35]
+icb_charts[36]
+icb_charts[37]
+icb_charts[38]
+icb_charts[39]
+icb_charts[40]
+icb_charts[41]
+icb_charts[42]
 ```
+
+###

### 2024-03-12 R/clustering_functions.R

@@ -327,6 +327,7 @@ get_cluster2_14_eth_chart <- function(clustered_gp_and_metrics,cluster_num){
     )|>
     left_join(titles)|>
     filter(cluster2==cluster_num)|>
+    mutate(ethnicity_name=fct_reorder(ethnicity_name,desc(med_percent))) |>
     ggplot(aes(x=fct_rev(factor(ethnicity_name)) , y=med_percent, fill=factor(ethnicity_name))) +
     geom_col() +
     coord_flip() +

### 2024-03-12 methods.qmd
Added cluster charts to targets pipeline
Adding more metrics to the data collection
Initialising gh-pages branch

@@
 
-### Attempt 1
+### Notes
 
@@
 <https://www.health.state.mn.us/data/mchs/pubs/raceethn/rankingbyratio20032007.pdf>
+
+## Smoking PCAS
+
+Metric 12 - Exception reporting for Smoking cessation support offered
+
+<https://digital.nhs.uk/data-and-information/publications/statistical/quality-and-outcomes-framework-achievement-prevalence-and-exceptions-data/2022-23>
+
+Donwloaded:
+
+qof-2223-prev-ach-pca-ls-prac.xlsx
+
+## References
+
+<https://journals.sagepub.com/doi/epdf/10.1093/phr/117.3.273>

### 2024-03-12 R/chart_functions.R

@@ -1,0 +1,400 @@
+
+get_rel_iod_data_for_chart <- function(activity_by_type_clusters_stg6){
+  chart_iod_data <- activity_by_type_clusters_stg6 |>
+  select(cluster2,list_size_total,ends_with("rel_iod"))|>
+  pivot_longer(cols=starts_with("metric"),
+               names_to="metric")|>
+  select(-cluster2,-list_size_total)|>
+  dplyr::distinct(metric,value)
+return(chart_iod_data)
+}
+
+
+get_rel_iod_chart <- function(chart_iod_data){
+
+  rel_iod_chart <- chart_iod_data |>
+    mutate(metric_long_name=case_when(metric=="metric2_rel_iod"~"Smoking prev est",
+                                      metric=="metric5_rel_iod"~"Smoking register",
+                                      metric=="metric6_rel_iod"~"Obesity register",
+                                      metric=="metric7_rel_iod"~"Diabetes register",
+                                      metric=="metric8_rel_iod"~"Depression register",
+                                      metric=="metric33_rel_iod"~"45+ BP check < 5 years",
+                                      metric=="metric9_rel_iod"~"CVD risk register",
+                                      metric=="metric34_rel_iod"~"CVD pat treated with LLT",
+                                      metric=="metric38_rel_iod"~"Pat at risk treated with LLT",
+                                      metric=="metric11_rel_iod"~"Smoking cessation support offered",
+                                      metric=="metric12_rel_iod"~"Except rep. smoke cease sup",
+                                      metric=="metric13_rel_iod"~"CHD register",
+                                      metric=="metric14_rel_iod"~"CT angiography",
+                                      metric=="metric15_rel_iod"~"Electrocardiography",
+                                      metric=="metric16_rel_iod"~"Asprin anti-platelet etc",
+                                      metric=="metric31_rel_iod"~"Except rep. asprin etc",
+                                      metric=="metric17_rel_iod"~"Flu vaccination",
+                                      metric=="metric32_rel_iod"~"Except rep. flu vacc",
+                                      metric=="metric39_rel_iod"~"65+ flu vaccination",
+                                      metric=="metric40_rel_iod"~"<65 at risk flu vaccination",
+                                      metric=="metric18_rel_iod"~"Ref to OP cardiology (new)",
+                                      metric=="metric19_rel_iod"~"Cardiology outpatient DNAs",
+                                      metric=="metric20_rel_iod"~"Elective PCI",
+                                      metric=="metric21_rel_iod"~"Elective CABG",
[…]

### 2024-03-12 R/clustering_functions.R

@@ -1,0 +1,573 @@
+################################################################################
+
+# Functions for clustering GP practices by ethnic group
+
+################################################################################
+
+get_full_cats_percents_over45 <- function(gp_lsoa_with_eth_sum_over45perc)
+{
+  full_cats_percents_over45 <- gp_lsoa_with_eth_sum_over45perc |>
+    select(practice_code,gp_sum_est_bangladeshi,gp_sum_est_chinese,gp_sum_est_indian,
+           gp_sum_est_pakistani,gp_sum_est_other_asian,gp_sum_est_blk_african,gp_sum_est_blk_caribbean,
+           gp_sum_est_other_blk,gp_sum_est_all_mixed,gp_sum_est_white_british,gp_sum_est_white_irish,
+           gp_sum_est_white_other,gp_sum_est_other_arab,gp_sum_est_other_other,gp_sum_total,perc_over45) |>
+    mutate(gp_perc_est_bangladeshi=(gp_sum_est_bangladeshi/gp_sum_total)*100,
+           gp_perc_est_chinese=(gp_sum_est_chinese/gp_sum_total)*100,
+           gp_perc_est_indian=(gp_sum_est_indian/gp_sum_total)*100,
+           gp_perc_est_pakistani=(gp_sum_est_pakistani/gp_sum_total)*100,
+           gp_perc_est_other_asian=(gp_sum_est_other_asian/gp_sum_total)*100,
+           gp_perc_est_blk_african=(gp_sum_est_blk_african/gp_sum_total)*100,
+           gp_perc_est_blk_caribbean=(gp_sum_est_blk_caribbean/gp_sum_total)*100,
+           gp_perc_est_other_blk=(gp_sum_est_other_blk/gp_sum_total)*100,
+           gp_perc_est_all_mixed=(gp_sum_est_all_mixed/gp_sum_total)*100,
+           gp_perc_est_white_british=(gp_sum_est_white_british/gp_sum_total)*100,
+           gp_perc_est_white_irish=(gp_sum_est_white_irish/gp_sum_total)*100,
+           gp_perc_est_white_other=(gp_sum_est_white_other/gp_sum_total)*100,
+           gp_perc_est_other_arab=(gp_sum_est_other_arab/gp_sum_total)*100,
+           gp_perc_est_other_other=(gp_sum_est_other_other/gp_sum_total)*100
+    ) |>
+    select(-gp_sum_est_bangladeshi,-gp_sum_est_chinese,-gp_sum_est_indian,
[…]

### 2024-03-12 R/data_collection_functions.R

@@ -1,0 +1,58 @@
+################################################################################
+
+# Functions for various data collection methods
+
+################################################################################
+
+
+#Read an excel data file
+read_excel_file <- function(file){
+  excel_data <- read_excel(file) 
+}
+
+#Read a csv data file
+read_csv_file <- function(file){
+  csv_data <- read_csv(file) |>
+    clean_names()
+}
+
+
+
+#Get Fingertips data Value
+get_my_fingertips_gp_data <- function(ind,year){
+  Metric <- fingertips_data(IndicatorID = ind,AreaTypeID = "All") |>
+    filter(AreaType=="GPs", Timeperiod==year) |>
+    select(AreaCode,Value)
+}
+
+#Get Fingertips data count
+get_my_fingertips_gp_data_count <- function(ind,year){
+  Metric <- fingertips_data(IndicatorID = ind,AreaTypeID = "All") |>
+    filter(AreaType=="GPs", Timeperiod==year) |>
+    select(AreaCode,Count,Denominator)
+}
+
+#Get QOF via Excel file
+read_qof_excel_file <- function(file){
+  excel_qof_data <- read_xlsx(file,skip=11,.name_repair = "unique_quiet") 
+}
+
+
+
+#Connect to Server and get data
+get_data_via_server <- function(){
+  StrategicWorking_DB <- dbConnect(odbc(), 
+                                   driver = "SQL Server",
+                                   server="MLCSU-BI-SQL",
+                                   database="StrategicWorking")
+  server_metric<- dbFetch(dbSendQuery(StrategicWorking_DB,"SELECT [PRACTICE_CODE]
+,[VALUE]
+FROM [StrategicWorking].[adhoc].[QOF_ACHIEVEMENT_2021_v2]
+where
+indicator_code = 'CHD007'
+and
+measure = 'PCAS'
+order by PRACTICE_CODE")) |>
+    clean_names()
+  #colnames(server_metric)[colnames(server_metric) == 'VALUE'] <- 'Metric32'
+}

### 2024-03-12 R/data_processing_functions.R

@@ -1,0 +1,216 @@
+
+
+# Process Census 2021 ethnicity data to remove wales and organise ethnicities
+process_census21data <- function(data){
+  data |>
+    filter(str_detect(lsoa_code, '^E'))|> 
+    group_by(lsoa_code) |> 
+    summarise(sum_all = sum(total_all_usual_residents),
+              all_asian = sum(asian_asian_british_or_asian_welsh),
+              bangladeshi = sum(asian_asian_british_or_asian_welsh_bangladeshi),
+              chinese = sum(asian_asian_british_or_asian_welsh_chinese),
+              indian =  sum(asian_asian_british_or_asian_welsh_indian),
+              pakistani =  sum(asian_asian_british_or_asian_welsh_pakistani),
+              other_asian =  sum(asian_asian_british_or_asian_welsh_other_asian),
+              all_black = sum(black_black_british_black_welsh_caribbean_or_african),
+              blk_african = sum(black_black_british_black_welsh_caribbean_or_african_african),
+              blk_caribbean = sum(black_black_british_black_welsh_caribbean_or_african_caribbean),
+              other_blk = sum(black_black_british_black_welsh_caribbean_or_african_other_black),
+              all_mixed = sum(mixed_or_multiple_ethnic_groups),
+              mixed_other = sum(mixed_or_multiple_ethnic_groups_other_mixed_or_multiple_ethnic_groups),
+              mixed_white_asian = sum(mixed_or_multiple_ethnic_groups_white_and_asian),
+              mixed_white_blk_african = sum(mixed_or_multiple_ethnic_groups_white_and_black_african),
+              mixed_white_blk_caribbean = sum(mixed_or_multiple_ethnic_groups_white_and_black_caribbean),
+              all_white = sum(white),
+              white_british = sum(white_english_welsh_scottish_northern_irish_or_british),
+              white_irish = sum(white_irish),
+              white_gypsy_or_irish_traveller = sum(white_gypsy_or_irish_traveller),
+              white_roma = sum(white_roma),
+              white_other_white = sum(white_other_white),
+              all_other = sum(other_ethnic_group),
+              other_arab = sum(other_ethnic_group_arab),
+              other_other= sum(other_ethnic_group_any_other_ethnic_group)
+    ) |>
[…]

### 2024-03-12 R/iod_calc_functions.R

@@ -1,0 +1,112 @@
+get_one_metric <- function(activity_by_type_clusters_stg6,metric_id,denominator) {
+  one_metric_data <- activity_by_type_clusters_stg6 |>
+    select(cluster2,all_of(denominator),starts_with(metric_id))
+  return(one_metric_data)
+}
+
+compute_iod <- function(one_metric_data) {
+  # Compute index of disparity
+  data <- one_metric_data|>
+    mutate(upper_ci=(numerator+(sqrt(numerator)*1.96))*(1/denominator),
+           lower_ci=(numerator-(sqrt(numerator)*1.96))*(1/denominator),
+           rate=numerator/denominator,
+           se=(rate-upper_ci)/1.96,
+           sd=(rate-mean(numerator)),
+           global_rate=sum(numerator)/sum(denominator),
+           diff = rate-global_rate,
+           abs_diff=abs(diff*denominator))
+  iod_abs <- sum(data$abs_diff)/2
+  iod_rel <- sum(data$abs_diff)/(2*sum(data$numerator))
+  return(iod_rel)
+}
+
+get_rates_per_cluster <- function(one_metric_data){
+  data <- one_metric_data
+  i<-1
+  n<-as.numeric(nrow(data))
+  
+  for (i in 1:n){
+    num=data$numerator[i]
+    denom=data$denominator[i]
+    rate=num/denom
+    upper_ci=(num+(sqrt(num)*1.96))*(1/denom)
+    se=(upper_ci-rate)/1.96
+    rates <- rnorm(n=1000,rate,se)
+    if(i==1)
+    { rate_clusters <- as.data.frame(rates) }
+    else 
+    { rate_clusters <- rate_clusters|> bind_cols(as.data.frame(rates))}
+    i=i+1
+  }
+  
+  rate_clusters <- rate_clusters |>
+    rename(rates1=`rates...1`,
+           rates2=`rates...2`,
+           rates3=`rates...3`,
+           rates4=`rates...4`,
+           rates5=`rates...5`
+    )|>
+    mutate(num1=rates1*data$denominator[1],
+           num2=rates2*data$denominator[2],
+           num3=rates3*data$denominator[3],
+           num4=rates4*data$denominator[4],
+           num5=rates5*data$denominator[5],
+           denom1=data$denominator[1],
+           denom2=data$denominator[2],
+           denom3=data$denominator[3],
[…]

### 2024-03-12 R/metric_processing_functions.R

@@ -1,0 +1,792 @@
+################################################################################
+
+# Functions to pre-process metrics
+
+################################################################################
+
+# splits practices into those with and those without prev, finds neighbours
+# of those missing prev and populates with mean of neighbours
+get_missing_chd_prevalence <- function(metric1,gp_history_short,joined_gp_history_and_chd_prev,gp_list_summary,gp_geocoded){
+  
+  joined <- gp_list_summary |>
+    left_join(joined_gp_history_and_chd_prev) |>
+    #where there is no matching value from orig in the gp list summary then flag these 
+    mutate(has_orig_prev=case_when(Value>=0 ~ 1, .default =  0))
+  
+  orig_with_geocode_and_id <- gp_geocoded |>
+    select(org_code,practice_name,postcode,has_orig_prev,geometry) |>
+    mutate(id=row_number())
+  
+  no_prev <- orig_with_geocode_and_id |>
+    filter(has_orig_prev == 0)
+  
+  with_prev <- orig_with_geocode_and_id |>
+    filter(has_orig_prev == 1)
+  
+  no_prev |>
+    select(no_prev_postcode = postcode) |>
+    st_join(with_prev, st_nearest_feature)
+  
+  no_prev_with_neighbours <-  st_sf(no_prev) |>
+    select(no_prev_postcode = postcode,no_prev_org_code=org_code) |>
+    # finds all the gp's in 1.5 km
+    st_join(st_sf(with_prev), st_is_within_distance, 1500)
+  
+  no_prev_with_neighbours_and_neighbours_prev <- no_prev_with_neighbours |>
+    left_join(joined_gp_history_and_chd_prev, join_by(org_code)) |>
+    group_by(no_prev_org_code) |>
+    summarise(prev_est=mean(Value)) |>
+    st_drop_geometry(select(no_prev_org_code,prev_est))
+  
+  metric1_updated <- joined |>
+    left_join(no_prev_with_neighbours_and_neighbours_prev,join_by(org_code==no_prev_org_code)) |>
+    mutate(chd_prev_to_use = case_when(Value>=0 ~ Value,
+                                       .default = prev_est)
+    ) |>
+    select(org_code,chd_prev_to_use)
+  return(metric1_updated)
+}
+
+
+
+#process metric 23 to work out whether los of cabg and pci is below trimpoint
+metric23_below_trimpoint <- function(metric23,metric23trimpoints){
[…]

### 2024-03-12 TestScript.R

@@ -1,0 +1,266 @@
+library(dplyr)
+library(stringr)
+library(MLID)
+library(tidyverse)
+
+lsoa_eth_sum <- ethnicities |>  
+  filter(str_detect(LSOA, '^E'))|> 
+  group_by(LSOA) |> 
+  summarise(sum_persons = sum(Persons),
+                                      sum_whitebritish = sum(WhiteBrit),
+                                      sum_irish = sum(Irish),
+                                      sum_otherwhite = sum(OtherWhite),
+                                      sum_othermixed = sum(Mixed),
+                                      sum_otherindian = sum(Indian),
+                                      sum_otherpakistani = sum(Pakistani),
+                                      sum_otherbangladeshi = sum(Bangladeshi),
+                                      sum_chinese = sum(Chinese),
+                                      sum_otherasian = sum(OtherAsian),
+                                      sum_blkafrican = sum(BlkAfrican),
+                                      sum_blkcaribbean = sum(BlkCaribbean),
+                                      sum_otherblk = sum(OtherBlk),
+                                      sum_arab = sum(Arab),
+                                      sum_other = sum(Other)
+                                      ) |> 
+  arrange()
+
+
+Densham <- gp_reg_pat_prac_lsoa_all |> 
+  filter(PRACTICE_CODE == "A81002") |> 
+  rename(LSOA = LSOA_CODE)
+
+
+Test <- gp_reg_pat_prac_lsoa_all |> 
+  rename(LSOA = LSOA_CODE)
+
+Densham_join <- Densham |> 
+  left_join(lsoa_eth_sum)
+
+Test_join <- Test |> 
+  left_join(lsoa_eth_sum)
+
+Test_Eth <- Test_join |> 
+  mutate(WhiteBritishEst = (sum_whitebritish/sum_persons)*NUMBER_OF_PATIENTS) |>
+  mutate(IrishEst = (sum_irish/sum_persons)*NUMBER_OF_PATIENTS) |>
+  mutate(OtherWhiteEst = (sum_otherwhite/sum_persons)*NUMBER_OF_PATIENTS) |>
[…]

### 2024-03-12 _targets.R

@@ -1,0 +1,540 @@
+# Created by use_targets().
+
+# Load packages required to define the pipeline:
+library(targets)
+
+# library(tarchetypes) # Load other packages as needed.
+
+# Set target options:
+tar_option_set(
+  packages = c("tibble","fingertipsR","readxl","tidyverse","utils","janitor","forcats",
+               "readr","visNetwork","odbc","stringr","MLID","sf","tidygeocoder","patchwork",
+               "cluster","factoextra","purrr","broom","glue","RColorBrewer","leaflet","treemapify"), # packages that your targets need to run
+  format = "rds"
+
+ 
+  # Set other options as needed.
+)
+
+# tar_make_clustermq() is an older (pre-{crew}) way to do distributed computing
+# in {targets}, and its configuration for your machine is below.
+options(clustermq.scheduler = "multiprocess")
+options(scipen=999)
+
+
+# Run the R scripts in the R/ folder with your custom functions:
+tar_source()
+# source("other_functions.R") # Source other scripts as needed.
+
+# Target list:
+list(
+  #Get data
+  tar_target(data_path1, "data/metric14.csv", format = "file"), # SUS data metric 14
+  tar_target(metric14,read_csv_file(data_path1)|> 
+               rename(metric14=number_ct_angiography)),
+  tar_target(data_path2, "data/metric15.csv", format = "file"), # SUS data metric 15
+  tar_target(metric15,read_csv_file(data_path2)|> 
+               rename(metric15=number_electrocardiography)),
+  tar_target(data_path3, "data/metric18.csv", format = "file"), # SUS data metric 18
+  tar_target(metric18,read_csv_file(data_path3)|> 
+               rename(metric18=cardi_op_attend)),
+  tar_target(data_path4, "data/metric19.csv", format = "file"), # SUS data metric 19
+  tar_target(metric19,read_csv_file(data_path4)|> 
+               rename(metric19=cardi_opdna)),
+  tar_target(data_path5, "data/metric20.csv", format = "file"), # SUS data metric 20
+  tar_target(metric20,read_csv_file(data_path5)|> 
+               rename(metric20=elective_pci)),
+  tar_target(data_path6, "data/metric21.csv", format = "file"), # SUS data metric 21
+  tar_target(metric21,read_csv_file(data_path6)|> 
+               rename(metric21=elective_cabg)),
+  
[…]

### 2024-03-12 clustering_generate_data.R

@@ -1,0 +1,997 @@
+
+##### Clustering tests#####
+library(factoextra)
+library(cluster)
+library(dplyr)
+library(targets)
+library(tidyverse)
+library(sf)
+library(tidygeocoder)
+library(leaflet)
+library(viridisLite)
+library(RColorBrewer)
+
+
+
+gp_reg_pat_prac <- tar_read(gp_reg_pat_prac_lsoa) |> as_tibble() |>
+  group_by(practice_code) |>
+  summarise(list_tot= sum(number_of_patients))
+
+gp_lsoa_with_eth_sum <- tar_read(gp_lsoa_with_eth_sum) |> as_tibble()
+
+full_cats <- gp_lsoa_with_eth_sum |>
+  select(practice_code,gp_sum_est_bangladeshi,gp_sum_est_chinese,gp_sum_est_indian,
+         gp_sum_est_pakistani,gp_sum_est_other_asian,gp_sum_est_blk_african,gp_sum_est_blk_caribbean,
+         gp_sum_est_other_blk,gp_sum_est_all_mixed,gp_sum_est_white_british,gp_sum_est_white_irish,
+         gp_sum_est_white_other,gp_sum_est_other_arab,gp_sum_est_other_other)
+
+five_cats <- gp_lsoa_with_eth_sum |>
+  select(practice_code,gp_sum_est_all_asian,gp_sum_est_all_black, gp_sum_est_all_mixed, gp_sum_est_all_white,gp_sum_est_all_other)
+
+#Move the practice code into row names
+full_cats <-
+  full_cats |>
+  remove_rownames() |>
+  column_to_rownames(var = 'practice_code')
+
+#Move the practice code into row names
+five_cats <-
+  five_cats |>
+  remove_rownames() |>
+  column_to_rownames(var = 'practice_code')
+
+
+#remove rows with missing values
+full_cats <- na.omit(full_cats)
+five_cats <- na.omit(five_cats)
+
+#scale each variable to have a mean of 0 and sd of 1
+# full_cats <- scale(full_cats)
+# five_cats <- scale(five_cats)
+
+scale_full_cats <- scale(full_cats)
+scale_five_cats <- scale(five_cats)
+
+# v <- prcomp(scale_full_cats)$x[,1]
+# scale_full_cats <- scale_full_cats[order(v),]
+# 
+# v <- prcomp(scale_five_cats)$x[,1]
[…]

### 2024-03-12 postcode.R

@@ -1,0 +1,110 @@
+library(targets)
+library(tidyverse)
+library(sf)
+library(tidygeocoder)
+library(leaflet)
+
+
+metric1 <- tar_read(metric1) |> as_tibble()
+gp_history <- tar_read(gp_history) |> as_tibble()
+gp_reg_pat_prac_lsoa <- tar_read(gp_reg_pat_prac_lsoa) |> as_tibble()
+gp_history_short <- tar_read(gp_history_short) |> as_tibble()
+joined_gp_history_and_chd_prev <- tar_read(joined_gp_history_and_chd_prev) |> as_tibble()
+gp_list_summary <- tar_read(gp_list_summary) |> as_tibble()
+gp_geocoded <- tar_read(gp_geocoded) |> as_tibble()
+
+gp_list_summary <- gp_reg_pat_prac_lsoa |> 
+  filter(sex=="ALL") |>
+  group_by(practice_code,practice_name) |>
+  summarise(number_of_patients=sum(number_of_patients)) |>
+  rename(org_code=practice_code)
+
+gp_history_short <- gp_history |> select(org_code, name, postcode,open_date,close_date,join_date,left_date)
+
+gp_list_summary <- gp_list_summary |> left_join(gp_history_short)
+
+orig <- metric1 |> select(AreaCode,AreaName, Value, IsCurrent) |>
+   rename(org_code=AreaCode) |>
+   left_join(gp_history_short)
+
+#prevalence is in orig 
+
+#join the orig and the gp list summary
+
+joined <- gp_list_summary |>
+  left_join(orig) |>
+#where there is no matching value from orig in the gp list summary then flag these 
+mutate(has_orig_prev=case_when(Value>=0 ~ 1, .default =  0))
+
+# pull those without a prev into their own table and geocode 
+
+# missing_prev <- joined |> filter(has_orig_prev==0) |>
+#   select(org_code,practice_name,postcode) |>
+#   mutate(id = row_number()) |>
+#   geocode(postalcode = postcode) |>
+#   st_as_sf(coords = c("long", "lat"), crs = 4326,na.fail=FALSE)
+
+# nothing more than a check that the points look valid
+# leaflet(missing_prev) |>
+#   addTiles() |>
+#   addMarkers()
+
+#find all the practices with a prevalence and geocode
+# orig_with_prev <- joined |>
+#   filter(has_orig_prev==1) |>
[…]

### 2024-03-12 processing_dev.R

@@ -1,0 +1,155 @@
+#process metrics ready to calc
+library(tidyverse)
+library(fingertipsR)
+library(readxl)
+
+clustered_gp_and_metrics <- targets::tar_read(clustered_gp_and_metrics) |> dplyr::as_tibble()
+activity_by_type_clusters_stg1<- targets::tar_read(activity_by_type_clusters_stg1) |> dplyr::as_tibble()
+activity_by_type_clusters_stg2<- targets::tar_read(activity_by_type_clusters_stg2) |> dplyr::as_tibble()
+activity_by_type_clusters_stg3<- targets::tar_read(activity_by_type_clusters_stg3) |> dplyr::as_tibble()
+activity_by_type_clusters_stg4<- targets::tar_read(activity_by_type_clusters_stg4) |> dplyr::as_tibble()
+activity_by_type_clusters_stg5<- targets::tar_read(activity_by_type_clusters_stg5) |> dplyr::as_tibble()
+activity_by_type_clusters_stg6<- targets::tar_read(activity_by_type_clusters_stg6) |> dplyr::as_tibble()
+ncdes_data <- tar_read(ncdes_data) |> as_tibble()
+gp_over45_perc <- tar_read(gp_over45_perc) |> as_tibble()
+gp_reg_pat_prac_sing_age_male <- tar_read(gp_reg_pat_prac_sing_age_male) |> as_tibble()
+gp_reg_pat_prac_sing_age_female <- tar_read(gp_reg_pat_prac_sing_age_female) |> as_tibble()
+
+
+
+activity_by_type_clusters_stg7 <- activity_by_type_clusters_stg6 |>
+  select(cluster2,list_size_total,ends_with("rel_iod"))|>
+         #metric28_rel_iod,metric29_rel_iod,metric31_rel_iod,metric32_rel_iod,metric33_rel_iod)|>
+  pivot_longer(cols=starts_with("metric"),
+               names_to="metric")|>
+  filter(cluster2==1)|>
+  select(-cluster2,-list_size_total)
+
+
+activity_by_type_clusters_stg7 |>
+  mutate(metric=fct_rev(factor(metric,levels=c("metric2_rel_iod","metric5_rel_iod","metric6_rel_iod","metric7_rel_iod","metric8_rel_iod",
+                                               "metric33_rel_iod","metric9_rel_iod","metric34_rel_iod","metric38_rel_iod","metric11_rel_iod",
[…]

### 2024-03-12 tests.R

@@ -1,0 +1,49 @@
+# load packages
+library(targets)
+library(dplyr)
+
+source("R/functions.R")
+
+
+use_targets()
+
+new_metric_16 <- tar_read(qof_chd_2223) |> as_tibble()
+
+metric_16 <- tar_read(metric16) |> as_tibble()
+
+metric_13 <- tar_read(metric13) |> as_tibble()
+
+new_metric_16_short <-new_metric_16 |> select(1,2,3,6,7,32,33,35,38)
+
+joined <- tar_read(joined) |> as_tibble()
+
+
+new_metric_16$patients_receiving_intervention_percent_38
+
+gp_lsoa <- tar_read(gp_lsoa) |> as_tibble()
+
+lsoa_eth_sum <- tar_read(lsoa_eth_sum) |> as_tibble()
+
+lsoa_lookup <- tar_read(lsoa_lookup) |> as_tibble()
+
+gp_lsoa_with_eth_sum <- tar_read(gp_lsoa_with_eth_sum) |> as_tibble()
+
+metric1 <- tar_read(metric1) |> as_tibble()
+gp_history <- tar_read(gp_history) |> as_tibble()
+
+metric40 <- tar_read(metric40) |> as_tibble()
+
+metric25b <- tar_read(metric25b) |> as_tibble()
+
+metric32 <- tar_read(metric32) |> as_tibble()
+
+indid <- 92304
+smok <- fingertips_data(IndicatorID = indid,AreaTypeID ="All")
+Metric05 <- Metric05 |> filter(AreaType=="GPs")
+Metric05 <- Metric05 %>% select(AreaCode,Value, Count)
+colnames(Metric05)[colnames(Metric05) == 'Value'] <- 'Metric05Prev'
+colnames(Metric05)[colnames(Metric05) == 'Count'] <- 'Metric05Count'
+#View(Metric05)
+
+
+

### 2024-03-12 worked_example.R

@@ -1,0 +1,100 @@
+one_metric_example_data <- tar_read(activity_by_type_clusters_stg6) |>
+  as_tibble()|>
+  select(cluster2,metric1_total,starts_with("metric29"))|>
+  rename(denominator=metric1_total,numerator=metric29_total)
+
+
+# Function to compute index of disparity
+compute_id <- function(one_metric_example_data) {
+  # Compute index of disparity
+  data <- one_metric_example_data|>
+    mutate(upper_ci=(numerator+(sqrt(numerator)*1.96))*(1/denominator),
+           lower_ci=(numerator-(sqrt(numerator)*1.96))*(1/denominator),
+           rate=numerator/denominator,
+           se=(rate-upper_ci)/1.96,
+           sd=(rate-mean(numerator)),
+           global_rate=sum(numerator)/sum(denominator),
+           diff = rate-global_rate,
+           abs_diff=abs(diff*denominator))
+  iod_abs <- sum(data$abs_diff)/2
+  iod_rel <- sum(data$abs_diff)/(2*sum(data$numerator))
+  return(iod_rel)
+}
+
+
+data <- one_metric_example_data
+i<-1
+n<-as.numeric(nrow(data))
+for (i in 1:n){
+  num=data$numerator[i]
+  denom=data$denominator[i]
+  rate=num/denom
+  upper_ci=(num+(sqrt(num)*1.96))*(1/denom)
+  se=(upper_ci-rate)/1.96
+  rates <- rnorm(n=1000,rate,se)
+  if(i==1)
+   { rate_clusters <- as.data.frame(rates) }
+  else 
+  { rate_clusters <- rate_clusters|> bind_cols(as.data.frame(rates))}
+  i=i+1
+}
+
+rate_clusters <- rate_clusters |>
+  rename(rates1=`rates...1`,
+         rates2=`rates...2`,
+         rates3=`rates...3`,
+         rates4=`rates...4`,
+         rates5=`rates...5`
+        )|>
+  mutate(num1=rates1*data$denominator[1],
+         num2=rates2*data$denominator[2],
+         num3=rates3*data$denominator[3],
+         num4=rates4*data$denominator[4],
+         num5=rates5*data$denominator[5],
+         denom1=data$denominator[1],
[…]

### 2024-03-12 findings.qmd
Updated findings_draft with re-arranged order of presentation and adding title banner and logo

@@
 
-This report quantifies and illustrates levels of inequity across 34 metrics at various points along the continuum of coronary heart disease progression and over a typical treatment pathway.
+This report quantifies and illustrates levels of inequity across 35 metrics at various points along the continuum of coronary heart disease progression and over a typical treatment pathway.
 They are shown in the table below grouped by domain (risk factors, risk factor identification, primary prevention, disease identification, secondary prevention, tertiary prevention, intermediate and full outcomes), which represent the various stages along the pathway.
@@
 ::: panel-tabset
-## Five categories
+## Five ethnic groups
 
@@
 
-## Fourteen categories
+## Fourteen ethnic groups
 
@@
 
-The map below presents each GP practice coloured according to their assigned cluster. Cluster 1 (red) is the least diverse and cluster 5 (orange) is the most diverse.
+The interactive map below presents each GP practice coloured according to their assigned cluster. Cluster 1 (red) is the least diverse and cluster 5 (orange) is the most diverse.
 
@@
 
-Metrics taken from the QOF data via Fingertips have been extracted by GP practice and the counts
-(numerator of the measure) from the calculations have been used.
-Exception reporting metrics have used the Personalised Care Adjustment (PCA) data to obtain the
-number of patients by GP practice for whom a PCA has been recorded for the relevant metric.
-Possible reasons for a PCA include; Newly diagnosed/registered, Intervention is clinically
-unsuitable, patient choice, did not respond to offers of care, service not available.
+Metrics taken from the QOF data via Fingertips have been extracted by GP practice and the counts (numerator of the measure) from the calculations have been used.
+Exception reporting metrics have used the Personalised Care Adjustment (PCA) data to obtain the number of patients by GP practice for whom a PCA has been recorded for the relevant metric.
+Possible reasons for a PCA include; Newly diagnosed/registered, Intervention is clinically unsuitable, patient choice, did not respond to offers of care, service not available.
 
@@
 
-SUS data was extracted using Transact-SQL and using relevant OPCS, ICD10, HRG, and Treatment
-function codes as detailed in table 2 above.
+SUS data was extracted using Transact-SQL and using relevant OPCS, ICD10, HRG, and Treatment function codes as detailed in table 2 above.
 
@@
 
-This data was supplied by NACR. The data wasn’t available by GP Practice, so an alternative method
-was followed.
+This data was supplied by NACR.
+The data wasn’t available by GP Practice, so an alternative method was followed.
 Practices were allocated to a cluster and data was then supplied aggregated into clusters.
@@
 For the purposes of this analysis the redacted data item was reassigned an average value.
-The true definition of eligible for rehabilitation is the HES admission and successful discharge of all
-patients with CVD e.g., MI, revascularization, or HF. However, to link with the NACR only those
-records entered into the database can be used so NACR use pseudo eligible as this is as close as
-the audit can get, at a granular level, to eligible. It is though affected by areas without entry to the
-database,
-The data used relates to the 2023 calendar year, which is the most recent complete year, and
-following changes in the way rehabilitation has been delivered due to the Covid pandemic, most
-closely reflects the rehabilitation model now in place.
+The true definition of eligible for rehabilitation is the HES admission and successful discharge of all patients with CVD e.g., MI, revascularization, or HF.
+However, to link with the NACR only those records entered into the database can be used so NACR use pseudo eligible as this is as close as the audit can get, at a granular level, to eligible.
+It is though affected by areas without entry to the database, The data used relates to the 2023 calendar year, which is the most recent complete year, and following changes in the way rehabilitation has been delivered due to the Covid pandemic, most closely reflects the rehabilitation model now in place.
 
@@
 
-The methodology used for this metric follows, as far as possible, that detailed by NHS Digital[^4]. The
-readmissions counted relate to those who had an elective PCI / CABG in the period 1/3/2022 to
-28/2/2023.
+The methodology used for this metric follows, as far as possible, that detailed by NHS Digital[^4].
+The readmissions counted relate to those who had an elective PCI / CABG in the period 1/3/2022 to 28/2/2023.
 
@@
 
-ONS death records contain an encrypted HES ID which was used to link to outpatient, inpatient and
-A&E HES records, from which the latest GP practice recorded (on a spine traced record) for the
-patient was then assigned to the death record.
+ONS death records contain an encrypted HES ID which was used to link to outpatient, inpatient and A&E HES records, from which the latest GP practice recorded (on a spine traced record) for the patient was then assigned to the death record.
 
@@
 
-xxxxxxxxxxxxxx
+The 2021 Census Ethnicity data by LSOA uses the 2021 LSOAs.
+The GP list size data uses the 2011 LSOAs.
+Some LSOAs that have gp list population assigned to them do not feature in the census data of ethnicity by lsoa.
+Need to check if these are the changes from 2011 to 2021 census.
+
+Possible solution would be to work out the average proportions for all the other LSOAs for the practice and distribute the gp list population for the lsoas that have no match proportionally across the ethnic groups.
+
+Another option would be where lsoas in 2021 have a parent lsoa in 2011 then apply the average ethnicity % for the child lsoas to the parent (this is when one 2011 lsoa has been split into multiple 2021 lsoas.) Where the lsoas in 2021 have a child lsoa in 2011 then apply the ethnicity % for the 2021 parent to the 2011 children (this is when several lsoas have been combined into one new lsoa)
+
+join lsoa_lookup to census 21eth using the 2021 lsoa code in the lookup dataset.
+
+Have opted for the latter solution.
+
+Imputing data in Census 2021:
+
+<https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/methodologies/itemeditingandimputationprocessforcensus2021englandandwales>
+
+Variation in age groups for different ethnicities:
+
+<https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/>
+
+<https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/#download-the-data>
 
@@
 
-The K-medoids clusterimg was performed using the {cluster} package[^7] in R using partitioning (clustering) of the data into k clusters "around medoids" (PAM), which is a more robust version of K-means.
+The K-medoids clustering was performed using the {cluster} package[^7] in R using partitioning (clustering) of the data into k clusters "around medoids" (PAM), which is a more robust version of K-means.
+The algorithm is based on the search for k representative objects or medoids among the observations of the dataset.
+After finding a set of k medoids, k clusters are constructed by assigning each observation to the nearest medoid.
+The goal is to find k representative objects which minimise the sum of the dissimilarities of the observations to their closest representative object.[^8]
 
@@
 
-Clustering experiments were conducted using a variety of different combinations of ethnic breakdowns, such as 14 ethnic groups and 5 ethnic groups, as well as variations that excluded the % of white british patients or factored in age group.
+[^8]: <https://cran.r-project.org/web/packages/cluster/cluster.pdf>
+
+Clustering experiments were conducted using a variety of different combinations of ethnic breakdowns, such as 14 ethnic groups and 5 ethnic groups, as well as variations that excluded the % of White British patients or factored in age group.
 Techniques were also used to attempt to determine the optimal number of clusters such as the elbow method, gap plots, and scree plots.
@@
 It was also important to seek a number of clusters that could be defined easily, where characteristics did not overlap each other so that resulting clusters were too similar to each other.
-It was also necessary to ensure that there was a sufficient number of practices in each cluster, too few practices in a cluster could result in issues with small amounts of metric data when calculating activity rates and consequently wide confidence intervals in the index of disparity calculations.
+It was also necessary to ensure that there was a sufficient number of practices in each cluster, too few practices in a cluster could result in issues with small amounts of metric data when calculating activity rates and consequently lead to wide confidence intervals in the index of disparity calculations.
 
-In conclusion, the final version selected and presented here used the percentage of the GP list size in each of the 14 census 2021 ethnic groups and five clusters was felt to give a sufficient number of describable clusters for which the index of disparity calculations could be performed.
-The chart below is an elbow plot generated for the 14 ethnic groups showing an elbow around 4 or 5 clusters.
+In conclusion, the final version selected, and presented here, used the percentage of the GP list size in each of the 14 census 2021 ethnic groups and five clusters was felt to give a sufficient number of describable clusters for which the index of disparity calculations could be performed.
+The chart below is an elbow plot generated for the 14 ethnic groups showing an elbow at around 4 or 5 clusters.
 
@@
 
-Table 3 shows the number of GP practices assigned to each of the 5 clusters and the characteristics of the clusters are descirbed above in the main body of the report.
+Table 3 shows the number of GP practices assigned to each of the 5 clusters and the characteristics of the clusters are described in the main body of the report above.
 
@@
 J., Doganaksoy, N., Meeker, W.
-Q., 2019)[^8]
+Q., 2019)[^9]
 
-[^8]: <https://academic.oup.com/jrssig/article/16/4/20/7038025>
+[^9]: <https://academic.oup.com/jrssig/article/16/4/20/7038025>
 
@@
 
-An explanation of bootstrap theory and confidence intervals is available in Practical Statistics for Data Scientists (Bruce, P. and Bruce, A., 2017)[^9]
+An explanation of bootstrap theory and confidence intervals is available in Practical Statistics for Data Scientists (Bruce, P. and Bruce, A., 2017)[^10]
 
-[^9]: <https://www.oreilly.com/library/view/practical-statistics-for/9781491952955/>
+[^10]: <https://www.oreilly.com/library/view/practical-statistics-for/9781491952955/>
 
@@
 
-All the code used in this project is available on the Strategy Unit GitHub[^10].
+All the code used in this project is available on the Strategy Unit GitHub[^11].
 
-[^10]: <https://github.com/The-Strategy-Unit/bhf_chd-ethnicity>
+[^11]: <https://github.com/The-Strategy-Unit/bhf_chd-ethnicity>

### 2024-03-12 findings.qmd
Additional report content

@@
 library(gt)
-metrics_table <- tibble(metric_domain = c("Need","Risk factors","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Primary prevention","Primary prevention","Primary prevention","Primary prevention","Primary prevention","Disease identification","Disease identification","Disease identification","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Intermediate outcome","Intermediate outcome","Intermediate outcome","Full outcomes","Full outcomes","Full outcomes","Full outcomes"),
+metrics_table <- tibble(metric_domain = c("Need","Risk factors","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Primary prevention","Primary prevention","Primary prevention","Primary prevention","Disease identification","Disease identification","Disease identification","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Intermediate outcome","Intermediate outcome","Intermediate outcome","Full outcomes","Full outcomes","Full outcomes","Full outcomes"),
   name_metric=c("CHD synthetic prevalence estimates",
@@
                                       "Depression register",
+                                      "The percentage of patients aged 45 or over who have a record of blood pressure in the preceding 5 years",
                                       "CVD risk register",
@@
                                       "Exception reporting for Smoking cessation support offered",
-                                      "The percentage of patients aged 45 or over who have a record of blood pressure in the preceding 5 years",
+
                                       "CHD register",
@@
 ```
+
+::: {.callout-note icon="false"}
+## Cardiac rehabilitation metrics
+
+Cardiac rehabilitation metrics will be added once they become available from NACR (expected before 31/3/24)
+:::
 
@@
 
-KEY FINDINGS
+The charts below show the activity rates by cluster for each metric along the CHD pathway.
+The horizontal line is the overall rate.
+The 95% confidence intervals are shown at the top of each bar.
+These activity rates take account of need by using CHD prevalence (or list size for risk factor) as the denominator of the calculation.
+Further detail on the methodology can be found in the appendix.
 
-The charts below show the rates by cluster for each metric along the CHD pathway.
-The horizontal line is the overall rate.
-Explain how its calculated and how to interpret.
-Mention that confidence intervals are also shown on the charts.
+In some cases these upper and lower limits are so small as to not be visible (e.g.
+Obesity register), whereas for other metrics (e.g.
+Readmission within 30 days of a PCI or CABG) they are much wider.
+Metrics with larger volumes of activity data give rise to a greater confidence in the calculated rate (e.g.
+limits are closer together) when compared to metrics with smaller volumes of activity data where there is less confidence in the calculated rate (e.g.
+limits are further apart).
+
+In calculating these rates for each cluster within each metric and then comparing the rates by cluster to the global rate for the metric it is possible to identify how activity rates vary across the 5 clusters.
+For there to be equity between the clusters the rates would need to be the same.
+The charts below present the metrics grouped according to pathway domain.
 
 ### Risk factors
+
+::: callout-tip
+## Key Findings
+
+-   Activity rates vary very little for the risk factor metric % patients 45+ with a record of BP \< 5 years
+
+-   Rates of patients on the diabetes register are highest in the most diverse cluster
+
+-   Rates of patients on the depression register are highest in the least diverse cluster and lowest in the most diverse cluster
+:::
+
+#### Risk factors
 
@@
 
-### Risk factor identification
+#### Risk factor identification
 
@@
 ### Primary prevention
+
+::: callout-tip
+## Key Findings
+
+-   The activity rate for CVD patients treated with LLT is highest for those patients whose practice is in the least diverse cluster and lowest for patients whose practice is in cluster 4.
+
+-   A similar pattern applies for at risk patients treated with LLT, activity rates being highest for those in the least diverse cluster.
+:::
 
@@
 
+::: callout-tip
+## Key Findings
+
+-   Activity rates vary considerably across the clusters in the disease identification metrics, with differing patterns for each metric
+
+-   Patients from clusters 4 and 5 (most diverse) are less likely to be recorded on the CHD register
+
+-   Clusters 5 and 4 have lower rates of CT angiography and electrocardiography
+:::
+
 ```{r}
@@
 ### Secondary prevention
+
+::: callout-tip
+## Key Findings
+
+-   There is little variation between clusters 1 to 4 in referral rates to outpatient cardiology, although the referrals are lower for the most diverse cluster (cluster 5)
+
+-   Flu vaccination rates for 65+ patients are highest among patients whose practice is in the least diverse cluster. the two most diverse clusters having the lowest rates
+
+-   Amongst patients under 65 who are at risk, flu vaccination rates are less varied between the clusters than for 65+ patients. Cluster 4 has the lowest rate of vaccination
+
+-   Cardiology outpatient DNA rates are higher the more diverse the cluster
+:::
 
@@
 
+::: callout-tip
+## Key Findings
+
+-   All 4 tertiary prevention metrics follow the same pattern of activity rates with the highest rates being in the least diverse cluster (cluster 1), followed by clusters 2 and then cluster 5 (the most diverse) and cluster 3.
+Cluster 4 has the lowest activity rates.
+:::
+
 ```{r}
@@
 ### Intermediate outcome
+
+::: callout-tip
+## Key Findings
+
+-   The least diverse cluster has the highest rates of BP readings \< 140/90.
+
+-   Emergency admissions for CHD and readmissions within 30 days of a PCI or CABG are highest for the least diverse cluster and lowest for cluster 4.
+:::
 
@@
 
+::: callout-tip
+## Key Findings
+
+...to be determined - awaiting death data check
+:::
+
 ```{r}
@@
 
-KEY FINDINGS
+The relative index of deprivation for the metric is expressed here as a %, such that the % indicates the amount of activity that would need to be redistributed between clusters in order that event rates follow levels of need.
+The detail of the methodology used for its calculation can be found in the appendix.
 
-and explain what it means and what it is showing
+The chart below presents the metrics along the CHD pathway from risk factor and identification through to outcome measures of death from CHD.
+For each metric the index of disparity is shown as a point estimate (yellow dot) with the upper and lower confidence intervals in grey.
+
+::: callout-tip
+## Key Findings
+
+-   The greatest disparity is flu vaccinations for patients aged 65+ where the relative index of disparity is 16.70% (95% ci - 16.64%, 16.70%), indicating that 16.7% of activity needs to be redistributed so that rates reflect need.
+:::
 
@@
 :::
+
+## Routes to equity
+
+A previous Strategy Unit report "Strategies to reduce inequalities in access to planned hospital procedures"[^4] highlights in chapter 1, three of the many routes from inequity to equity: levelling-up, levelling-down and zero-sum redistribution.
+Using the zero-sum redistribution route as an example, along with the metric flu vaccinations for patients aged 65+, this illustration shows how equity could be achieved by increasing and decreasing the number of vaccinations delivered in each cluster to give the same activity rate per cluster.
+
+[^4]: <https://www.midlandsdecisionsupport.nhs.uk/knowledge-library/strategies-to-reduce-inequalities-in-access-to-planned-hospital-procedures/>
+
+|  Cluster  | Number of vaccinations | CHD synthetic prevalence | Rate (No. of vaccinations / CHD synthetic prevalence) | Global rate |  Change in number of vaccinations to match global rate  |
+|:----------:|-----------:|-----------:|:----------:|:----------:|:----------:|
+|           |                    *n* |                      *d* |                        *r=n/d*                        | *gr=𝚺n/𝚺d*  |                      *(r-gr) x d*                       |
+|     1     |              3,356,002 |                1,026,826 |                         3.27                          |    2.33     |                        -961,239                         |
+|     2     |              2,638,698 |                  924,088 |                         2.86                          |    2.33     |                        -483,540                         |
+|     3     |              1,862,430 |                  958,825 |                         1.94                          |    2.33     |                        +373,741                         |
+|     4     |                476,616 |                  478,728 |                         1.00                          |    2.33     |                        +639,873                         |
+|     5     |                332,728 |                  327,542 |                         1.02                          |    2.33     |                        +431,166                         |
+| **Total** |          **8,666,474** |            **3,716,009** |                       **2.33**                        |             |                          **0**                          |
+
+: In reality, in attempting to achieve equity for this metric, it is unlikely that the offer of vaccination would be removed from patients whose practice is in clusters 1 and 2.
+More likely would be an approach to focus public health and vaccination roll-out campaigns on practices in clusters 3, 4 and 5.
+
+Each of the metrics along the CHD pathway with the higher relative indicies of disparity would almost certainly require a variety of different routes from inequity to equity given their differing nature.
+The more effective interventions on the pathway might be levelled-up, whilst those of more limited value might be levelled-down.
 
@@
 
+::: callout-tip
+## Key Findings
+
+-   The greatest disparity
+:::
+
 ::: panel-tabset
@@
 
-refer to ICBs of interest under relevant region
-
 The appendix contains equivalent charts to those above, showing the Index of Disparity for each CHD pathway metric for each of the 42 ICBs.
@@
 # Conclusions
+
+.......
 
@@
 
-The methodology used for this metric follows, as far as possible, that detailed by NHS Digital[^4].
+The methodology used for this metric follows, as far as possible, that detailed by NHS Digital[^5].
 The readmissions counted relate to those who had an elective PCI / CABG in the period 1/3/2022 to 28/2/2023.
 
-[^4]: <https://digital.nhs.uk/data-and-information/publications/statistical/ccg-outcomes-indicator-set/specifications/3.2-emergency-readmissions-within-30-days-of-discharge-from-hospital_1_4>
+[^5]: <https://digital.nhs.uk/data-and-information/publications/statistical/ccg-outcomes-indicator-set/specifications/3.2-emergency-readmissions-within-30-days-of-discharge-from-hospital_1_4>
 
@@
 In order to assign a prevalence for practices with missing data a nearest neighbours methodology was used to impute the prevalence by taking an average of the prevalence of each practice within a 1.5km radius of the practice without data.
-Further details, including code with sample data, for this methodology are available in a blog post on the Strategy Unit Data Science website[^5].
+Further details, including code with sample data, for this methodology are available in a blog post on the Strategy Unit Data Science website[^6].
 
-[^5]: <https://the-strategy-unit.github.io/data_science/blogs/posts/2024-01-17_nearest_neighbour.html>
+[^6]: <https://the-strategy-unit.github.io/data_science/blogs/posts/2024-01-17_nearest_neighbour.html>
 
@@
 
-Since metric data is mainly from the 2022/23 year, the October 2022 list size data was taken from the NHS Digital Patients Registered at a GP practice dataset[^6] as this is the data closest to the mid point in the 22/23 year.
+Since metric data is mainly from the 2022/23 year, the October 2022 list size data was taken from the NHS Digital Patients Registered at a GP practice dataset[^7] as this is the data closest to the mid point in the 22/23 year.
 The GP to ICB mapping file was also downloaded and joined to the list size to easily create analysis by ICB and region.
 
-[^6]: <https://digital.nhs.uk/data-and-information/publications/statistical/patients-registered-at-a-gp-practice/october-2022>
+[^7]: <https://digital.nhs.uk/data-and-information/publications/statistical/patients-registered-at-a-gp-practice/october-2022>
 
@@
 
+The 2021 Census Ethnicity data was downloaded from the .gov.uk website.[^8] Summary details of the census ethnicity data is also available and this includes descriptions of the standardised list of ethnic groups and the differences from the 2011 ethnic groups.[^9]
+
+[^8]: <https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/#download-the-data>
+
+[^9]: <https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/>
+
 The 2021 Census Ethnicity data by LSOA uses the 2021 LSOAs.
 The GP list size data uses the 2011 LSOAs.
-Some LSOAs that have gp list population assigned to them do not feature in the census data of ethnicity by lsoa.
-Need to check if these are the changes from 2011 to 2021 census.
+This results in gaps when the datasets are combined.
+To overcome this the following method was used:
 
-Possible solution would be to work out the average proportions for all the other LSOAs for the practice and distribute the gp list population for the lsoas that have no match proportionally across the ethnic groups.
+-   An LSOA lookup showing 2011 and 2021 LSOA parent and child relationships was joined to the 2021 Census Ethnicity data
 
-Another option would be where lsoas in 2021 have a parent lsoa in 2011 then apply the average ethnicity % for the child lsoas to the parent (this is when one 2011 lsoa has been split into multiple 2021 lsoas.) Where the lsoas in 2021 have a child lsoa in 2011 then apply the ethnicity % for the 2021 parent to the 2011 children (this is when several lsoas have been combined into one new lsoa)
+-   Where a 2021 LSOAs has a parent LSOA in 2011 then we applied the average ethnicity % for the child LSOAs to the parent (this is when one 2011 LSOA has been split into multiple 2021 LSOAs).
 
-join lsoa_lookup to census 21eth using the 2021 lsoa code in the lookup dataset.
+-   Where the LSOAs in 2021 have a child LSOA in 2011 then we applied the ethnicity % for the 2021 parent to the 2011 children (this is when several LSOAs have been combined into one new LSOA).
 
-Have opted for the latter solution.
+Further details about item editing and imputation processes for Census 2021, England and Wales are available on the ONS website.
+This details techniques used by ONS with the aim of arriving at a fully populated clean database.
+These techniques include manual imputation and nearest neighbour donor imputation.
+For the England and Wales Census 2021 the imputation rate for ethnic group was 1.3%.[^10]
 
-Imputing data in Census 2021:
-
-<https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/methodologies/itemeditingandimputationprocessforcensus2021englandandwales>
-
-Variation in age groups for different ethnicities:
-
-<https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/>
-
-<https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/#download-the-data>
-
-explain how 2011 lsoa and 2021 lsoa mis match was handled
+[^10]: <https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/methodologies/itemeditingandimputationprocessforcensus2021englandandwales>
 
@@
 
-The K-medoids clustering was performed using the {cluster} package[^7] in R using partitioning (clustering) of the data into k clusters "around medoids" (PAM), which is a more robust version of K-means.
+The K-medoids clustering was performed using the {cluster} package[^11] in R using partitioning (clustering) of the data into k clusters "around medoids" (PAM), which is a more robust version of K-means.
 The algorithm is based on the search for k representative objects or medoids among the observations of the dataset.
 After finding a set of k medoids, k clusters are constructed by assigning each observation to the nearest medoid.
-The goal is to find k representative objects which minimise the sum of the dissimilarities of the observations to their closest representative object.[^8]
+The goal is to find k representative objects which minimise the sum of the dissimilarities of the observations to their closest representative object.[^12]
 
-[^7]: Maechler M, Rousseeuw P, Struyf A, Hubert M, Hornik K (2023). *cluster: Cluster Analysis Basics and Extensions*.
+[^11]: Maechler M, Rousseeuw P, Struyf A, Hubert M, Hornik K (2023). *cluster: Cluster Analysis Basics and Extensions*.
 R package version 2.1.6 — For new features, see the 'NEWS' and the 'Changelog' file in the package source), [https://CRAN.R-project.org/package=cluster](https://cran.r-project.org/package=cluster).
 
-[^8]: <https://cran.r-project.org/web/packages/cluster/cluster.pdf>
+[^12]: <https://cran.r-project.org/web/packages/cluster/cluster.pdf>
 
@@
 
-remember...how , detail - allowing for NEED and how this is done
+#### Rate of activity
+
+Calculations of the index of disparity begin by calculating an activity rate using a numerator of the activity ( eg number of elective procedures) divided by a denominator, which for metrics other than risk factors has been taken to be CHD synthetic prevalence (as a determinant of need).
+
+Where the metric is in the risk factor domain the denominator used is GP practice list size (with the exception of the percentage of patients aged 45 or over who have a record of blood pressure in the preceding 5 years, where the 45+ list size has been used).
+
+In calculating these rates for each cluster within each metric and then comparing these rates by cluster to the global rate for the metric it is possible to identify how activity rates vary across the 5 clusters
+
+#### Index of disparity
+
+The relative index of disparity indicates the extent to which the rate of an activity or event varies across groups.
+Having calculated the rates for each cluster within each metric the disparity is calculated by taking the differences for each cluster from the global rate and then determining the amount of activity by which this cluster varies from the global.
+This absolute difference is then summed for the 5 clusters and divided by twice the sum of the numerators.
+This produces a relative index of deprivation for the metric, expressed as a %, such that the % indicates the amount of activity that would need to be redistributed between clusters in order that event rates follow levels of need.
+
+Index of disparity = (𝚺 \| r~(1–n)~ - R\| / n) / R\*100
+
+where r = group rate and R = total population rate
+
+A report by Pearcy and Keppel (2002) presents the Index of Disparity as a summary measure of disparity across population groups and advocates its use for groups defined in terms of categories such as ethnicity.[^13]
+
+[^13]: <https://journals.sagepub.com/doi/epdf/10.1093/phr/117.3.273>
 
@@
 J., Doganaksoy, N., Meeker, W.
-Q., 2019)[^9]
+Q., 2019)[^14]
 
-[^9]: <https://academic.oup.com/jrssig/article/16/4/20/7038025>
+[^14]: <https://academic.oup.com/jrssig/article/16/4/20/7038025>
 
@@
 
-An explanation of bootstrap theory and confidence intervals is available in Practical Statistics for Data Scientists (Bruce, P. and Bruce, A., 2017)[^10]
+An explanation of bootstrap theory and confidence intervals is available in Practical Statistics for Data Scientists (Bruce, P. and Bruce, A., 2017)[^15]
 
-[^10]: <https://www.oreilly.com/library/view/practical-statistics-for/9781491952955/>
+[^15]: <https://www.oreilly.com/library/view/practical-statistics-for/9781491952955/>
 
@@
 
-All the code used in this project is available on the Strategy Unit GitHub[^11].
+All the code used in this project is available on the Strategy Unit GitHub[^16].
 
-[^11]: <https://github.com/The-Strategy-Unit/bhf_chd-ethnicity>
+[^16]: <https://github.com/The-Strategy-Unit/bhf_chd-ethnicity>

### 2024-03-12 R/chart_functions.R

@@ -388,13 +388,14 @@ get_region_cluster_chart <- function(region_cluster_chart_data){
   palette <- brewer.pal(5, "Set1")
   options(scipen=999)
   region_cluster_chart<- region_cluster_chart_data |>
-    ggplot(aes(x=as.factor(cluster2),y=format(cluster_list_size, scientific = FALSE), fill=as.factor(cluster2) ))+
+    ggplot(aes(x=as.factor(cluster2),y=cluster_list_size, fill=as.factor(cluster2) ))+
     geom_bar(stat = "identity") +
     scale_fill_manual(values = palette)+
     theme_light() +
     theme(legend.position="none")+
     xlab("Cluster (1=Least diverse, 5=Most diverse)")+
     ylab("List Size")+
+    scale_y_continuous(limits = c(0,NA))+
     labs(title = paste("List Size by Cluster"))
   return(region_cluster_chart)
 }

### 2024-03-12 R/functions.R

@@ -129,4 +129,11 @@ join_over45_to_gp_lsoa_with_eth_sum <- function(gp_lsoa_with_eth_sum,gp_over45_p
   return(gp_lsoa_with_eth_sum_over45perc)
 }
 
-
+show_in_excel <- function(.data){
+  
+  tmp <- paste0(tempfile(), ".csv")
+  
+  write.csv(.data, tmp)
+  
+  fs::file_show(path = tmp)
+}

### 2024-03-13 draft_report.qmd
draft report

--- 2024-03-12 findings.qmd
@@
 
-This report quantifies and illustrates levels of inequity across 35 metrics at various points along the continuum of coronary heart disease progression and over a typical treatment pathway.
+This report quantifies and illustrates levels of inequity across 33 metrics at various points along the continuum of coronary heart disease progression and over a typical treatment pathway.
 They are shown in the table below grouped by domain (risk factors, risk factor identification, primary prevention, disease identification, secondary prevention, tertiary prevention, intermediate and full outcomes), which represent the various stages along the pathway.
@@
 
-...to be determined - awaiting death data check
+...to be determined - awaiting death data checks
 :::
@@
 
--   The greatest disparity is flu vaccinations for patients aged 65+ where the relative index of disparity is 16.70% (95% ci - 16.64%, 16.70%), indicating that 16.7% of activity needs to be redistributed so that rates reflect need.
+-   The greatest disparity along the CHD pathway is flu vaccinations for patients aged 65+ where the relative index of disparity is 16.70% (95% ci - 16.64%, 16.70%), indicating that 16.7% of activity needs to be redistributed so that rates reflect need.
+
+-   The most equitably distributed metric is blood pressure checks within the last 5 years for patients aged 45+ with a relative index of disparity of 0.8% (95% ci - 0.82%, 0.78%).
+
+-   There are 7 metrics where the index of disparity is less than 5%, as well as blood pressure checks, these include metrics such as referrals to outpatient cardiology, outpatient DNAs and waiting times for elective procedures.
+
+-   Out of the 33 metrics along the CHD pathway, 10 have an index of disparity of greater than 10%.
+
+-   In general (but not exclusively) these greater levels of disparity occur in the disease identification and secondary prevention pathway domains and consequently the outcome metrics related to deaths from CHD.
 :::
@@
 A previous Strategy Unit report "Strategies to reduce inequalities in access to planned hospital procedures"[^4] highlights in chapter 1, three of the many routes from inequity to equity: levelling-up, levelling-down and zero-sum redistribution.
-Using the zero-sum redistribution route as an example, along with the metric flu vaccinations for patients aged 65+, this illustration shows how equity could be achieved by increasing and decreasing the number of vaccinations delivered in each cluster to give the same activity rate per cluster.
+Using the zero-sum redistribution route as an example, along with the metric flu vaccinations for patients aged 65+, the illustration below shows how equity could be achieved by increasing and decreasing the number of vaccinations delivered in each cluster to give the same activity rate per cluster.
 
@@
 
-: In reality, in attempting to achieve equity for this metric, it is unlikely that the offer of vaccination would be removed from patients whose practice is in clusters 1 and 2.
+In reality, in attempting to achieve equity for this metric, it is unlikely that the offer of vaccination would be removed from patients whose practice is in clusters 1 and 2.
 More likely would be an approach to focus public health and vaccination roll-out campaigns on practices in clusters 3, 4 and 5.
@@
 
--   The greatest disparity
+-   Looking at a regional level, the greatest disparity is in CT angiography in the South West, Midland, and the North West.
 :::
@@
 
-.......
+To be added as part of finalisation of the report.
 
@@
 
-| Domain                     | Metric                                 | Data Source | Year    | Definition and selection criteria/codes                                                 |
-|---------------|---------------|---------------|---------------|---------------|
-| Need                       | CHD synthetic prevalence estimates     | PHE         | 2015    | Provided directly by OHID (formerly indicator 92847 in FingertipsR)                     |
-| Risk factors               | Smoking synthetic prevalence estimates | GP survey   | 2023    |                                                                                         |
-| Risk factor identification | Smoking register                       | NHSD QOF    | 2022/23 | [Indicator 91280 extracted fromFingertipsR](https://fingertips.phe.org.uk/search/91280) |
-|                            | Obesity register                       | NHSD QOF    |         | [Indicator 93088 extracted fromFingertipsR](https://fingertips.phe.org.uk/search/93088) |
-|                            | Diabetes register                      | NHSD QOF    |         |                                                                                         |
-|                            | Depression register                    | NHSD QOF    |         |                                                                                         |
-| Primary prevention         |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-| Full Outcomes              | Deaths in hospital from CHD            |             |         |                                                                                         |
-|                            | Deaths in hospital from CHD \< 75      |             |         |                                                                                         |
-|                            | Deaths from CHD                        |             |         |                                                                                         |
-|                            | Deaths from CHD \<75                   |             |         |                                                                                         |
+...
 
@@
 
+The charts below are equivalent charts to those in the regional analysis section of the report, showing the Index of Disparity for each CHD pathway metric for each of the 42 ICBs.
+There is considerable variation in the levels of disparity across the ICBs as well as the metrics along the pathway.
+
 ```{r}

### 2024-03-13 R/chart_functions.R

@@ -29,8 +29,8 @@ get_rel_iod_chart <- function(chart_iod_data){
                                       metric=="metric15_rel_iod"~"Electrocardiography",
                                       metric=="metric16_rel_iod"~"Asprin anti-platelet etc",
                                       metric=="metric31_rel_iod"~"Except rep. asprin etc",
-                                      metric=="metric17_rel_iod"~"Flu vaccination",
-                                      metric=="metric32_rel_iod"~"Except rep. flu vacc",
+                                      # metric=="metric17_rel_iod"~"Flu vaccination",
+                                      # metric=="metric32_rel_iod"~"Except rep. flu vacc",
                                       metric=="metric39_rel_iod"~"65+ flu vaccination",
                                       metric=="metric40_rel_iod"~"<65 at risk flu vaccination",
                                       metric=="metric18_rel_iod"~"Ref to OP cardiology (new)",
@@ -51,8 +51,9 @@ get_rel_iod_chart <- function(chart_iod_data){
                                                                      "Depression register", "45+ BP check < 5 years","CVD risk register","CVD pat treated with LLT",
                                                                      "Pat at risk treated with LLT","Smoking cessation support offered",
                                                                      "Except rep. smoke cease sup","CHD register","CT angiography","Electrocardiography",
-                                                                     "Asprin anti-platelet etc","Except rep. asprin etc", "Flu vaccination",
-                                                                     "Except rep. flu vacc", "65+ flu vaccination","<65 at risk flu vaccination",
+                                                                     "Asprin anti-platelet etc","Except rep. asprin etc", 
+                                                                     #"Flu vaccination","Except rep. flu vacc", 
+                                                                     "65+ flu vaccination","<65 at risk flu vaccination",
                                                                      "Ref to OP cardiology (new)", "Cardiology outpatient DNAs","Elective PCI",
                                                                      "Elective CABG","Waiting time for elective PCI / CABG","PCI/CABG disch < trimpoint",
                                                                      "BP reading < 140/90 (<80 yrs CHD)","Readmit<30 days of PCI / CABG",
@@ -74,8 +75,8 @@ get_rel_iod_chart <- function(chart_iod_data){
[…]

### 2024-03-13 R/metric_processing_functions.R

@@ -69,21 +69,25 @@ metric23_below_trimpoint <- function(metric23,metric23trimpoints){
   return(metric23_updated)
 }
 
-combine_metrics17_and_32 <- function(metric32,metric17){
-  metric32_updated <- metric17 |>
-    left_join(metric32)|>
-    select(-metric17_num)|>
-    rename(metric32_denom=metric17_denom)
-  
-  return(metric32_updated)
-}
+# combine_metrics17_and_32 <- function(metric32,metric17){
+#   metric32_updated <- metric17 |>
+#     left_join(metric32)|>
+#     select(-metric17_num)|>
+#     rename(metric32_denom=metric17_denom)
+#   
+#   return(metric32_updated)
+# }
 
 add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final_data_full_cats_percent_5_clusters,
                             gp_16andover_pop,gp_list_45over,
                             metric1_updated,metric2,metric5,metric6,metric7,metric8,metric9,metric11,metric12,metric13,metric13b,
-                            metric14,metric15,metric16,metric16b,metric17,metric18,metric19,metric20,
+                            metric14,metric15,metric16,metric16b,
+                            #metric17,
+                            metric18,metric19,metric20,
                             metric21,metric22,metric23_updated,metric25b,metric26,metric27,metric28,metric28b,
-                            metric29,metric29b,metric31,metric32_updated,metric33,metric34,metric38,
+                            metric29,metric29b,metric31,
+                            #metric32_updated,
+                            metric33,metric34,metric38,
                             metric39,metric40){
   clustered_gp_and_metrics <-
     final_data_full_cats_percent_over45_5_clusters |> rename(cluster1=cluster)|>
@@ -117,7 +121,7 @@ add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final
     left_join(metric15)|>
     left_join(metric16)|>
     left_join(metric16b)|>
-    left_join(metric17)|>
+   # left_join(metric17)|>
     left_join(metric18)|>
     left_join(metric19)|>
     left_join(metric20)|>
@@ -135,7 +139,7 @@ add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final
     left_join(metric29b)|>
[…]

### 2024-03-13 _targets.R

@@ -157,10 +157,10 @@ list(
                rename(gp_practice_code=AreaCode,
                       metric16_num=Count,
                       metric16_denom=Denominator)), # metric 16
-  tar_target(metric17,get_my_fingertips_gp_data_count(91000,"2020/21")|>
-               rename(gp_practice_code=AreaCode,
-                      metric17_num=Count,
-                      metric17_denom=Denominator)), # metric 17
+  # tar_target(metric17,get_my_fingertips_gp_data_count(91000,"2020/21")|>
+  #              rename(gp_practice_code=AreaCode,
+  #                     metric17_num=Count,
+  #                     metric17_denom=Denominator)), # metric 17
 
 # metric 9 has been removed from Fingertips :-(  this was the original code
 # tar_target(metric9,get_my_fingertips_gp_data(92589,"2019/20")), # metric 9
@@ -200,10 +200,10 @@ list(
                           rename(gp_practice_code=practice_code,
                                  metric25_num=numerator_40,
                                  metric25_denom=denominator_plus_pc_as_45)),   # metric 25b 22/23
-  tar_target(metric32,get_data_via_server() |>
-               rename(gp_practice_code=practice_code,
-                      metric32_num=value)), # metric 32 20/21
-  tar_target(metric32_updated,combine_metrics17_and_32(metric32,metric17)),
+  # tar_target(metric32,get_data_via_server() |>
+  #              rename(gp_practice_code=practice_code,
+  #                     metric32_num=value)), # metric 32 20/21
+  # tar_target(metric32_updated,combine_metrics17_and_32(metric32,metric17)),
   tar_target(metric13b,qof_chd_2223 |> 
                           select(4,8,6,7) |> #15
                           rename(gp_practice_code=practice_code,
@@ -291,10 +291,13 @@ tar_target(clustered_gp_and_metrics,
                            gp_16andover_pop,gp_list_45over,
                            metric1_updated,metric2,metric5,metric6,metric7,metric8,metric9,metric11,
                            metric12,metric13,metric13b,metric14,metric15,metric16,
-                           metric16b,metric17,metric18,metric19,metric20,
+                           metric16b,
+                           #metric17,
[…]

### 2024-03-13 deaths_qa.R

@@ -1,0 +1,90 @@
+clustered_gp_and_metrics <- tar_read(clustered_gp_and_metrics)|>as_tibble()
+activity_by_type_clusters_stg6 <- tar_read(activity_by_type_clusters_stg6)|>as_tibble()
+gp_reg_pat_prac_sing_age_female<- tar_read(gp_reg_pat_prac_sing_age_female)|>as_tibble()
+gp_reg_pat_prac_sing_age_male<- tar_read(gp_reg_pat_prac_sing_age_male)|>as_tibble()
+gp_reg_pat_prac_lsoa <- tar_read(gp_reg_pat_prac_lsoa)|>as_tibble()
+
+need <- activity_by_type_clusters_stg6|>select(cluster2,metric1_total)|>rename(chdprev_need=metric1_total)
+
+gp_75over_perc <- gp_reg_pat_prac_sing_age_male |>
+  rbind(gp_reg_pat_prac_sing_age_female)|>
+  filter(age != "ALL") |>
+  group_by(org_code,age) |>
+  summarise(number_of_patients=sum(number_of_patients)) |>
+  dplyr::mutate(age=as.numeric(case_when(age=="95+" ~ 95,
+                                         .default = as.numeric(age))
+  )) |>
+  mutate(over_75=case_when(age>= 75 ~1,
+                           .default = 0)) |>
+  group_by(org_code,over_75)|>
+  summarise(number_of_patients=sum(number_of_patients))|>
+  group_by(org_code) |>
+  mutate(total_patients=sum(number_of_patients))|>
+  ungroup() |>
+  mutate(perc=(number_of_patients/total_patients)*100)|>
+  filter(over_75==1) |>
+  select(org_code,perc)|>
+  rename(gp_practice_code=org_code,perc_over75=perc)
+
+gp_16over_female_num <- gp_reg_pat_prac_sing_age_female |>
+  filter(age != "ALL") |>
+  group_by(org_code,age) |>
+  summarise(number_of_patients=sum(number_of_patients)) |>
+  dplyr::mutate(age=as.numeric(case_when(age=="95+" ~ 95,
+                                         .default = as.numeric(age))
+  )) |>
+  mutate(over_16=case_when(age>= 16 ~1,
+                           .default = 0)) |>
+  group_by(org_code,over_16)|>
+  summarise(number_of_patients=sum(number_of_patients))|>
+  filter(over_16==1)|>
+  select(-over_16)|>
+  rename(gp_practice_code=org_code,num_females=number_of_patients)
+
+
[…]

### 2024-03-14 draft_report.qmd
Corrected error in text

@@
 
-The relative index of deprivation for the metric is expressed here as a %, such that the % indicates the amount of activity that would need to be redistributed between clusters in order that event rates follow levels of need.
+The relative index of disparity for the metric is expressed here as a %, such that the % indicates the amount of activity that would need to be redistributed between clusters in order that event rates follow levels of need.
 The detail of the methodology used for its calculation can be found in the appendix.

### 2024-03-14 findings.qmd
Corrected error in text

@@
 | Domain                     | Metric                                                                                                                                                                                                                                                 | Data Source | Year    | Definition and selection criteria/codes                                                  |
-|---------------|---------------|---------------|---------------|---------------|
+|-------------|-----------------|-------------|-------------|----------------|
 | Need                       | CHD synthetic prevalence estimates                                                                                                                                                                                                                     | PHE         | 2015    | Provided directly by OHID (formerly indicator 92847 in FingertipsR)                      |
-| Risk factors               | Smoking synthetic prevalence estimates | GP survey   | 2023    |                                                                                         |
-| Risk factor identification | Smoking register                       | NHSD QOF    | 2022/23 | [Indicator 91280 extracted fromFingertipsR](https://fingertips.phe.org.uk/search/91280) |
-|                            | Obesity register                       | NHSD QOF    |         | [Indicator 93088 extracted fromFingertipsR](https://fingertips.phe.org.uk/search/93088) |
-|                            | Diabetes register                      | NHSD QOF    |         |                                                                                         |
-|                            | Depression register                    | NHSD QOF    |         |                                                                                         |
-| Primary prevention         |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-| Full Outcomes              | Deaths in hospital from CHD            |             |         |                                                                                         |
-|                            | Deaths in hospital from CHD \< 75      |             |         |                                                                                         |
-|                            | Deaths from CHD                        |             |         |                                                                                         |
-|                            | Deaths from CHD \<75                   |             |         |                                                                                         |
+| Risk factors               | Smoking synthetic prevalence estimates                                                                                                                                                                                                                 | GP Survey   | 2023    |                                                                                          |
+| Risk factor identification | Smoking register                                                                                                                                                                                                                                       | NHSD QOF    | 2022/23 | [Indicator 91280 extracted from FingertipsR](https://fingertips.phe.org.uk/search/91280) |
+|                            | Obesity register                                                                                                                                                                                                                                       | NHSD QOF    | 2022/23 | [Indicator 93088 extracted from FingertipsR](https://fingertips.phe.org.uk/search/93088) |
+|                            | Diabetes register                                                                                                                                                                                                                                      | NHSD QOF    | 2022/23 | [Indicator 241 extracted from FingertipsR](https://fingertips.phe.org.uk/search/241)     |
+|                            | Depression register                                                                                                                                                                                                                                    | NHSD QOF    | 2022/23 | [Indicator 848 extracted from FingertipsR](https://fingertips.phe.org.uk/search/848)     |
+|                            | The percentage of patients aged 45 or over who have a record of blood pressure in the preceding 5 years                                                                                                                                                | NHSD QOF    | 2022/23 | [Indicator 91262 extracted from FingertipsR](https://fingertips.phe.org.uk/search/91262) |
+|                            | CVD risk register                                                                                                                                                                                                                                      | NHSD QOF    | 2019/20 | Formerly indicator 92589 extracted from FingertipsR                                      |
+| Primary prevention         | Percentage of patients aged 18 and over with GP recorded CVD (narrow definition), who are currently treated with lipid lowering therapy                                                                                                                | CVD Prevent | 2022/23 | <https://www.cvdprevent.nhs.uk/home>                                                     |
+|                            | Percentage of patients aged 18 and over, with no GP recorded CVD and a GP recorded QRISK score of 10% or more, CKD (G3a to G5), T1 diabetes (aged 40 and over) or T2 diabetes aged 60 and over, who are currently treated with lipid lowering therapy.
+| CVD Prevent | 2022/23 | <https://www.cvdprevent.nhs.uk/home>                                                     |
+|                            | Smoking cessation support offered                                                                                                                                                                                                                      | NHSD QOF    | 2022/23 | [Indicator 90619 extracted from FingertipsR](https://fingertips.phe.org.uk/search/90619) |
+|                            | Exception reporting for Smoking cessation support offered                                                                                                                                                                                              | NHSD QOF    | 2022/23 | [Indicator 90619 extracted from FingertipsR](https://fingertips.phe.org.uk/search/90619) |
+| Disease Identification     | CHD register                                                                                                                                                                                                                                           | NHSD QOF    | 2022/23 | [Indicator 273 extracted from FingertipsR](https://fingertips.phe.org.uk/search/273)     |
+|                            | CT angiography                                                                                                                                                                                                                                         | SUS         | 2022/23 | OPCS Code: U102 - Cardiac                                                                |
+
+```         
+                                                                                                                                                                                                                                                                                                                    computed tomography angiography.
+                                                                                                                                                                                                                                                                                                                    Elective only                                                                                           |
+```
+
+|                            \| Electrocardiography \| SUS \| 2022/23 \| OPCS: U19 & U34. Elective only \|
+| Secondary Prevention \| Aspirin, anti-platelet or anti-coagulent \| NHSD QOF \| 2022/23 \| [Indicator 90999 extracted from FingertipsR](https://fingertips.phe.org.uk/search/90999) \|
+|                            \| Exception reporting for Aspirin, anti-platelet or anti-coagulent \| NHSD QOF \| 2022/23 \| [Indicator 90999 extracted from FingertipsR](https://fingertips.phe.org.uk/search/90999) \|
+|                            \| Percentage of patients aged 65 or over who received a seasonal influenza vaccination between 1 September 2022 and 31 March 2023 \| IIF \| March 2023 \| <https://www.england.nhs.uk/primary-care/primary-care-networks/network-contract-des/iif/> \|
+|                            \| Percentage of patients aged 18 to 64 years and in a clinical at-risk group who received a seasonal influenza vaccination between 1 September 2022 and 31 March 2023 \| IIF \| March 2023 \| <https://www.england.nhs.uk/primary-care/primary-care-networks/network-contract-des/iif/> \|
+|                            \| Referral to cardiology (First outpatient) \| SUS \| 2022/23 \| TFC = 320 \|
+|                            \| Cardiology outpatient DNAs \| SUS \| 2022/23 \| TFC = 320 \|
+| Tertiary Prevention \| Elective PCI \| SUS \| 2022/23 \| OPCS Code: K49, K50, K75. FCE = 1 \|
+|                            \| Elective CABG \| SUS \| 2022/23 \| OPCS Code: K40-K46, FCE = 1 \|
+|                            \| Waiting time for elective PCI / CABG \| SUS \| 2022/23 \| Elective admissions. OPCS Codes as per metrics 20 and 21 \|
+|                            \| Elective PCI / CABG patients discharged before trimpoint \| SUS \| 2022/23 \| Main PCI and CABG Trim points for 2022/23 linked for HRG \|
+|                            \| Cardiac rehabilitation - Started \| NACR \| 2023 \| Data supplied by ICB. Aggregated ACS and HF patients. \|
+|                            \| Cardiac rehabilitation - Completed \| NACR \| 2023 \| Data supplied by ICB. Aggregated ACS and HF patients. \|
+| Intermediate Outcomes \| Last BP reading of patients (\<80 yrs, with CHD) in last 12mths is \<=140/90 \| NHSD QOF \| 2022/23 \| CHD008 \|
+|                            \| Readmission within 30 days of elective PCI / CABG \| SUS \| 2022/23 (MarFeb) \| No.
+of emerg.
+spells up to 31/03/2023 within 0-29 days (inclusive) of the last, previous discharge from hospital / No.
+of finished spells with discharge date between 01/03/2022 and 28/02/2023 Exclude: TFC = 501, 560, 610, OPCS starting with O, Classpat = 1 (Ord.), Any diagnosis = C00\*-C97\*, D37\*-D48\* \|
+|                            \| Emergency admissions for CHD \| SUS \| 2022/23 \| PRIMARY diagnosis = I20 or I21 or I22 or I23 or I24 or I25 \|
+| Full Outcomes \| Deaths in hospital from CHD \| SUS \| 2022/23 \| Following any admission Elective or Emergency. PRIMARY diagnosis = I20 or I21 or I22 or I23 or I24 or I25 \|
+|                            \| Deaths in hospital from CHD \< 75 \| SUS \| 2022/23 \| Following any admission Elective or Emergency.
+PRIMARY diagnosis = I20 or I21 or I22 or I23 or I24 or I25.
+Age \<=75 \|
+|                            \| Deaths from CHD \| ONS Death records \| 2022/23 \| Underlying Cause of Death = I20 or I21 or I22 or I23 or I24 or I25 \|
+|                            \| Deaths from CHD \<75 \| ONS Death records \| 2022/23 \| Underlying Cause of Death = I20 or I21 or I22 or I23 or I24 or I25. Age \<=75 \|
 

### 2024-03-14 draft_report.qmd
Repaired broken table in appendix

@@
 
+*Table 2 - achieving equity via zero-sum redistribution*
+
 |  Cluster  | Number of vaccinations | CHD synthetic prevalence | Rate (No. of vaccinations / CHD synthetic prevalence) | Global rate | Change in number of vaccinations to match global rate |
@@
 
--   Looking at a regional level, the greatest disparity is in CT angiography in the South West, Midland, and the North West.
+-   Looking at a regional level, the greatest disparity is in CT angiography in the South West, Midlands, and the North West.
 :::
@@
 
-...
+```{r}
+library(gt)
+metrics_table <- tibble(metric_domain = c("Need","Risk factors","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Primary prevention","Primary prevention","Primary prevention","Primary prevention","Disease identification","Disease identification","Disease identification","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Intermediate outcome","Intermediate outcome","Intermediate outcome","Full outcomes","Full outcomes","Full outcomes","Full outcomes"),
+  name_metric=c("CHD synthetic prevalence estimates",
+                                      "Smoking synthetic prevalence estimates",
+                                      "Smoking register",
+                                      "Obesity register",
+                                      "Diabetes register",
+                                      "Depression register",
+                                      "The percentage of patients aged 45 or over who have a record of blood pressure in the preceding 5 years",
+                                      "CVD risk register",
+                                      "Percentage of patients aged 18 and over with GP recorded CVD (narrow definition), who are currently treated with lipid lowering therapy",
+                                      "Percentage of patients aged 18 and over, with no GP recorded CVD and a GP recorded QRISK score of 10% or more, CKD (G3a to G5), T1 diabetes (aged 40 and over) or T2 diabetes aged 60 and over, who are currently treated with lipid lowering therapy",
+                                      "Smoking cessation support offered",
+                                      "Exception reporting for Smoking cessation support offered",
+                                      "CHD register",
+                                      "CT angiography",
+                                      "Electrocardiography",
+                                      "Aspirin, anti-platelet or anti-coagulent",
+                                      "Exception reporting for Aspirin, anti-platelet or anti-coagulent",
+                                      "Percentage of patients aged 65 or over
+who received a seasonal influenza vaccination
+between 1 September 2022 and 31 March 2023",
+                                      " Percentage of patients aged 18 to 64
+years and in a clinical at-risk group who received
+a seasonal influenza vaccination between 1
+September 2022 and 31 March 2023",
+"Referral to cardiology (First outpatient)",
+                                      "Cardiology outpatient DNAs",
+"Elective PCI","Elective CABG","Waiting time for elective PCI / CABG","Elective PCI / CABG patients discharged before trimpoint",
+"Cardiac rehabilitation - Started", "Cardiac rehabilitation - Completed","BP reading < 140/90","Readmission within 30 days of elective PCI / CABG","Emergency admissions for CHD","Deaths in hospital from CHD","Deaths in hospital from CHD <75","Deaths from CHD","Deaths from CHD <75"
+                                      ),
+data_source=c("PHE","GP Survey","NHSD QOF","NHSD QOF","NHSD QOF","NHSD QOF","NHSD QOF","NHSD QOF","CVD Prevent","CVD Prevent",
+              "NHSD QOF","NHSD QOF","NHSD QOF","SUS","SUS","NHSD QOF","NHSD QOF","IIF","IIF","SUS",
+              "SUS","SUS","SUS","SUS","SUS","NACR","NACR","NHSD QOF","SUS","SUS",
+              "SUS","SUS","ONS Death records","ONS Death records"),
+year=c("2015","2023","2022/23","2022/23","2022/23","2022/23","2022/23","2019/20","2022/23","2022/23",
+              "2022/23","2022/23","2022/23","2022/23","2022/23","2022/23","2022/23","March 2023","March 2023","2022/23",
+              "2022/23","2022/23","2022/23","2022/23","2022/23","2023","2023","2022/23","2022/23 (Mar-Feb)","2022/23",
+              "2022/23","2022/23","2022/23","2022/23"),
+definition=c("Provided directly by OHID (formerly indicator 92847 in FingertipsR)","","Indicator 91280 extracted from FingertipsR","Indicator 93088 extracted from FingertipsR","Indicator 241 extracted from FingertipsR","Indicator 848 extracted from FingertipsR","Indicator 91262 extracted from FingertipsR","Formerly indicator 92589 extracted from FingertipsR","https://www.cvdprevent.nhs.uk/home","https://www.cvdprevent.nhs.uk/home",
+              "Indicator 90619 extracted from FingertipsR","Indicator 90619 extracted from FingertipsR","Indicator 273 extracted from FingertipsR","OPCS Code: U102 - Cardiac computed tomography angiography.
+Elective only","OPCS: U19 & U34.
+Elective only","Indicator 90999 extracted from FingertipsR","Indicator 90999 extracted from FingertipsR","https://www.england.nhs.uk/primary-care/primary-care-networks/network-contract-des/iif/","https://www.england.nhs.uk/primary-care/primary-care-networks/network-contract-des/iif/","TFC = 320",
+              "TFC = 320","OPCS Code: K49, K50, K75.
+FCE = 1","OPCS Code: K40-K46, FCE = 1","Elective admissions.
+OPCS Codes as per metrics 20 and 21","Main PCI and CABG Trim points for 2022/23 linked for HRG","Data supplied by ICB.
+Aggregated ACS and HF patients","Data supplied by ICB.
+Aggregated ACS and HF patients","CHD008","No.
+of emerg.
+spells up to 31/03/2023 within 0-29 days (inclusive) of the last, previous discharge from hospital / No.
+of finished spells with discharge date between 01/03/2022 and 28/02/2023 Exclude: TFC = 501, 560, 610, OPCS starting with O, Classpat = 1 (Ord.), Any diagnosis = C00*-C97*, D37*-D48*","PRIMARY diagnosis = I20 or I21 or I22 or I23 or I24 or I25",
+              "Following any admission Elective or Emergency.
+PRIMARY diagnosis = I20 or I21 or I22 or I23 or I24 or I25","Following any admission Elective or Emergency.
+PRIMARY diagnosis = I20 or I21 or I22 or I23 or I24 or I25.
+Age <=75","Underlying Cause of Death = I20 or I21 or I22 or I23 or I24 or I25","Underlying Cause of Death = I20 or I21 or I22 or I23 or I24 or I25.
+Age <=75")
+                        )
+
+metrics_table |>
+    gt()|>
+    tab_header(
+    title = md("**Table 3** - *Metric definitions and data sources*"))|>
+    cols_label(name_metric=md("**Metric**"),
+             metric_domain=md("**Domain**"),
+             data_source=md("**Data Source**"),
+             year=md("**Year**"),
+             definition=md("**Definition and selection criteria/codes**"))
+```
 
@@
   gt()|>
-    tab_header(title = md("**Table 3** - *Number of practices per cluster*"))|>
+    tab_header(title = md("**Table 4** - *Number of practices per cluster*"))|>
     cols_label(cluster=md("**Cluster**"),

### 2024-03-14 findings.qmd
Repaired broken table in appendix

--- 2024-03-12 findings.qmd
@@
 A previous Strategy Unit report "Strategies to reduce inequalities in access to planned hospital procedures"[^4] highlights in chapter 1, three of the many routes from inequity to equity: levelling-up, levelling-down and zero-sum redistribution.
-Using the zero-sum redistribution route as an example, along with the metric flu vaccinations for patients aged 65+, this illustration shows how equity could be achieved by increasing and decreasing the number of vaccinations delivered in each cluster to give the same activity rate per cluster.
+Using the zero-sum redistribution route as an example, along with the metric flu vaccinations for patients aged 65+, the illustration below shows how equity could be achieved by increasing and decreasing the number of vaccinations delivered in each cluster to give the same activity rate per cluster.
 
 [^4]: <https://www.midlandsdecisionsupport.nhs.uk/knowledge-library/strategies-to-reduce-inequalities-in-access-to-planned-hospital-procedures/>
+
+*Table 2 - achieving equity via zero-sum redistribution*
 
@@
 
--   The greatest disparity
+-   Looking at a regional level, the greatest disparity is in CT angiography in the South West, Midlands, and the North West.
 :::
@@
 
-| Domain                     | Metric                                 | Data Source | Year    | Definition and selection criteria/codes                                                 |
-|---------------|---------------|---------------|---------------|---------------|
-| Need                       | CHD synthetic prevalence estimates     | PHE         | 2015    | Provided directly by OHID (formerly indicator 92847 in FingertipsR)                     |
-| Risk factors               | Smoking synthetic prevalence estimates | GP survey   | 2023    |                                                                                         |
-| Risk factor identification | Smoking register                       | NHSD QOF    | 2022/23 | [Indicator 91280 extracted fromFingertipsR](https://fingertips.phe.org.uk/search/91280) |
-|                            | Obesity register                       | NHSD QOF    |         | [Indicator 93088 extracted fromFingertipsR](https://fingertips.phe.org.uk/search/93088) |
-|                            | Diabetes register                      | NHSD QOF    |         |                                                                                         |
-|                            | Depression register                    | NHSD QOF    |         |                                                                                         |
-| Primary prevention         |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-| Full Outcomes              | Deaths in hospital from CHD            |             |         |                                                                                         |
-|                            | Deaths in hospital from CHD \< 75      |             |         |                                                                                         |
-|                            | Deaths from CHD                        |             |         |                                                                                         |
-|                            | Deaths from CHD \<75                   |             |         |                                                                                         |
+```{r}
+library(gt)
+metrics_table <- tibble(metric_domain = c("Need","Risk factors","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Primary prevention","Primary prevention","Primary prevention","Primary prevention","Disease identification","Disease identification","Disease identification","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Intermediate outcome","Intermediate outcome","Intermediate outcome","Full outcomes","Full outcomes","Full outcomes","Full outcomes"),
+  name_metric=c("CHD synthetic prevalence estimates",
+                                      "Smoking synthetic prevalence estimates",
+                                      "Smoking register",
+                                      "Obesity register",
+                                      "Diabetes register",
+                                      "Depression register",
+                                      "The percentage of patients aged 45 or over who have a record of blood pressure in the preceding 5 years",
+                                      "CVD risk register",
+                                      "Percentage of patients aged 18 and over with GP recorded CVD (narrow definition), who are currently treated with lipid lowering therapy",
+                                      "Percentage of patients aged 18 and over, with no GP recorded CVD and a GP recorded QRISK score of 10% or more, CKD (G3a to G5), T1 diabetes (aged 40 and over) or T2 diabetes aged 60 and over, who are currently treated with lipid lowering therapy",
+                                      "Smoking cessation support offered",
+                                      "Exception reporting for Smoking cessation support offered",
+                                      "CHD register",
+                                      "CT angiography",
+                                      "Electrocardiography",
+                                      "Aspirin, anti-platelet or anti-coagulent",
+                                      "Exception reporting for Aspirin, anti-platelet or anti-coagulent",
+                                      "Percentage of patients aged 65 or over
+who received a seasonal influenza vaccination
+between 1 September 2022 and 31 March 2023",
+                                      " Percentage of patients aged 18 to 64
+years and in a clinical at-risk group who received
+a seasonal influenza vaccination between 1
+September 2022 and 31 March 2023",
+"Referral to cardiology (First outpatient)",
+                                      "Cardiology outpatient DNAs",
+"Elective PCI","Elective CABG","Waiting time for elective PCI / CABG","Elective PCI / CABG patients discharged before trimpoint",
+"Cardiac rehabilitation - Started", "Cardiac rehabilitation - Completed","BP reading < 140/90","Readmission within 30 days of elective PCI / CABG","Emergency admissions for CHD","Deaths in hospital from CHD","Deaths in hospital from CHD <75","Deaths from CHD","Deaths from CHD <75"
+                                      ),
+data_source=c("PHE","GP Survey","NHSD QOF","NHSD QOF","NHSD QOF","NHSD QOF","NHSD QOF","NHSD QOF","CVD Prevent","CVD Prevent",
+              "NHSD QOF","NHSD QOF","NHSD QOF","SUS","SUS","NHSD QOF","NHSD QOF","IIF","IIF","SUS",
+              "SUS","SUS","SUS","SUS","SUS","NACR","NACR","NHSD QOF","SUS","SUS",
+              "SUS","SUS","ONS Death records","ONS Death records"),
+year=c("2015","2023","2022/23","2022/23","2022/23","2022/23","2022/23","2019/20","2022/23","2022/23",
+              "2022/23","2022/23","2022/23","2022/23","2022/23","2022/23","2022/23","March 2023","March 2023","2022/23",
+              "2022/23","2022/23","2022/23","2022/23","2022/23","2023","2023","2022/23","2022/23 (Mar-Feb)","2022/23",
+              "2022/23","2022/23","2022/23","2022/23"),
+definition=c("Provided directly by OHID (formerly indicator 92847 in FingertipsR)","","Indicator 91280 extracted from FingertipsR","Indicator 93088 extracted from FingertipsR","Indicator 241 extracted from FingertipsR","Indicator 848 extracted from FingertipsR","Indicator 91262 extracted from FingertipsR","Formerly indicator 92589 extracted from FingertipsR","https://www.cvdprevent.nhs.uk/home","https://www.cvdprevent.nhs.uk/home",
+              "Indicator 90619 extracted from FingertipsR","Indicator 90619 extracted from FingertipsR","Indicator 273 extracted from FingertipsR","OPCS Code: U102 - Cardiac computed tomography angiography.
+Elective only","OPCS: U19 & U34.
+Elective only","Indicator 90999 extracted from FingertipsR","Indicator 90999 extracted from FingertipsR","https://www.england.nhs.uk/primary-care/primary-care-networks/network-contract-des/iif/","https://www.england.nhs.uk/primary-care/primary-care-networks/network-contract-des/iif/","TFC = 320",
+              "TFC = 320","OPCS Code: K49, K50, K75.
+FCE = 1","OPCS Code: K40-K46, FCE = 1","Elective admissions.
+OPCS Codes as per metrics 20 and 21","Main PCI and CABG Trim points for 2022/23 linked for HRG","Data supplied by ICB.
+Aggregated ACS and HF patients","Data supplied by ICB.
+Aggregated ACS and HF patients","CHD008","No.
+of emerg.
+spells up to 31/03/2023 within 0-29 days (inclusive) of the last, previous discharge from hospital / No.
+of finished spells with discharge date between 01/03/2022 and 28/02/2023 Exclude: TFC = 501, 560, 610, OPCS starting with O, Classpat = 1 (Ord.), Any diagnosis = C00*-C97*, D37*-D48*","PRIMARY diagnosis = I20 or I21 or I22 or I23 or I24 or I25",
+              "Following any admission Elective or Emergency.
+PRIMARY diagnosis = I20 or I21 or I22 or I23 or I24 or I25","Following any admission Elective or Emergency.
+PRIMARY diagnosis = I20 or I21 or I22 or I23 or I24 or I25.
+Age <=75","Underlying Cause of Death = I20 or I21 or I22 or I23 or I24 or I25","Underlying Cause of Death = I20 or I21 or I22 or I23 or I24 or I25.
+Age <=75")
+                        )
+
+metrics_table |>
+    gt()|>
+    tab_header(
+    title = md("**Table 3** - *Metric definitions and data sources*"))|>
+    cols_label(name_metric=md("**Metric**"),
+             metric_domain=md("**Domain**"),
+             data_source=md("**Data Source**"),
+             year=md("**Year**"),
+             definition=md("**Definition and selection criteria/codes**"))
+
+```
 

### 2024-03-14 draft_report.qmd
Updated details re rehabilitation metrics

@@
               "2022/23","2022/23","2022/23","2022/23","2022/23","2022/23","2022/23","March 2023","March 2023","2022/23",
-              "2022/23","2022/23","2022/23","2022/23","2022/23","2023","2023","2022/23","2022/23 (Mar-Feb)","2022/23",
+              "2022/23","2022/23","2022/23","2022/23","2022/23","2021 & 2022","2021 & 2022","2022/23","2022/23 (Mar-Feb)","2022/23",
               "2022/23","2022/23","2022/23","2022/23"),
@@
 For the purposes of this analysis the redacted data item was reassigned an average value.
-The true definition of eligible for rehabilitation is the HES admission and successful discharge of all patients with CVD e.g., MI, revascularization, or HF.
-However, to link with the NACR only those records entered into the database can be used so NACR use pseudo eligible as this is as close as the audit can get, at a granular level, to eligible.
-It is though affected by areas without entry to the database, The data used relates to the 2023 calendar year, which is the most recent complete year, and following changes in the way rehabilitation has been delivered due to the Covid pandemic, most closely reflects the rehabilitation model now in place.
+The data used relates to the 2021 & 2022 calendar years, which are the most recent complete years, and following changes in the way rehabilitation has been delivered due to the Covid pandemic, most closely reflects the rehabilitation model now in place.
 

### 2024-04-12 draft_report.qmd
fixing

@@
 |  Cluster  | Number of vaccinations | CHD synthetic prevalence | Rate (No. of vaccinations / CHD synthetic prevalence) | Global rate | Change in number of vaccinations to match global rate |
-|:----------:|-----------:|-----------:|:----------:|:----------:|:----------:|
+|:---------:|-----------------------:|-------------------------:|:-----------------------------------------------------:|:-----------:|:-----------------------------------------------------:|
 |           |                    *n* |                      *d* |                        *r=n/d*                        | *gr=𝚺n/𝚺d*  |                     *(r-gr) x d*                      |

### 2024-04-12 R/chart_functions.R

@@ -238,7 +238,6 @@ get_ci_iod_chart <- function(iod_with_ci){
 
 
 get_rate_chart_data <- function(activity_by_type_clusters_stg6){
-  options(scipen=999)
   rate_data <- activity_by_type_clusters_stg6 |>
   select(cluster2,ends_with("_rate"),ends_with("ci"))|>
   select(-ends_with("global_rate"),-ends_with("diff_rate"))|>
@@ -280,9 +279,9 @@ get_rate_chart_data <- function(activity_by_type_clusters_stg6){
                              metric=="metric23_lower_ci"~"Tertiary prevention",
                            metric=="metric25_upper_ci"|metric=="metric26_upper_ci"|metric=="metric27_upper_ci"|
                              metric=="metric25_lower_ci"|metric=="metric26_lower_ci"|metric=="metric27_lower_ci"~"Intermediate outcome",
-                           metric=="metric28_upper_ci"|metric=="metric28b_upper_ci"|metric=="metric29_upper_ci"|
-                             metric=="metric29b_upper_ci"|metric=="metric28_lower_ci"|metric=="metric28b_lower_ci"|metric=="metric29_lower_ci"|
-                             metric=="metric29b_er_ci"~"Full outcomes",
+                           metric=="metric28_upper_ci"|metric=="metric28b_upper_ci"|metric=="metric29_upper_ci"| metric=="metric29b_upper_ci"|
+                             metric=="metric28_lower_ci"|metric=="metric28b_lower_ci"|metric=="metric29_lower_ci"|
+                             metric=="metric29b_lower_ci"~"Full outcomes",
                            .default="other"))|>
   mutate(metric_name=case_when(metric=="metric2_rate"|metric=="metric2_lower_ci"|metric=="metric2_upper_ci"~"Smoking prev est",
                                metric=="metric5_rate"|metric=="metric5_lower_ci"|metric=="metric5_upper_ci"~"Smoking register",
@@ -368,6 +367,7 @@ rate_chart <- rate_chart_data|>
   geom_bar(stat = "identity") +
   geom_errorbar( aes(x=as.factor(cluster2), ymin=lower_ci, ymax=upper_ci), width=0.4, colour="grey", alpha=0.9, size=0.5) +
   scale_fill_manual(values = palette)+
+ # scale_y_continuous(comma(rate,digits=4,format="g"))+#added this line
   theme_light() +
   theme(legend.position="none")+
   xlab("Cluster (1=Least diverse, 5=Most diverse)") +

### 2024-04-12 R/clustering_functions.R

@@ -276,7 +276,7 @@ get_cluster2_chart <- function(clustered_gp_and_metrics){
     ylim(0,25) +
     labs(title="Median Percent by Ethnicity") +
     theme(axis.text.y=element_text(size=6))+
-    guides(fill = FALSE) 
+    guides(fill = "none") 
   
   return(cluster2_chart)
   
@@ -337,7 +337,7 @@ get_cluster2_14_eth_chart <- function(clustered_gp_and_metrics,cluster_num){
     ylim(0,25) +
     labs(title="Median Percent by Ethnicity") +
     theme(axis.text.y=element_text(size=6))+
-    guides(fill = FALSE) 
+    guides(fill = "none") 
   
   return(cluster2_14_eth_chart)
   

### 2024-04-12 R/metric_processing_functions.R

@@ -69,25 +69,21 @@ metric23_below_trimpoint <- function(metric23,metric23trimpoints){
   return(metric23_updated)
 }
 
-# combine_metrics17_and_32 <- function(metric32,metric17){
-#   metric32_updated <- metric17 |>
-#     left_join(metric32)|>
-#     select(-metric17_num)|>
-#     rename(metric32_denom=metric17_denom)
-#   
-#   return(metric32_updated)
-# }
+combine_metrics17_and_32 <- function(metric32,metric17){
+  metric32_updated <- metric17 |>
+    left_join(metric32)|>
+    select(-metric17_num)|>
+    rename(metric32_denom=metric17_denom)
+  
+  return(metric32_updated)
+}
 
 add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final_data_full_cats_percent_5_clusters,
                             gp_16andover_pop,gp_list_45over,
                             metric1_updated,metric2,metric5,metric6,metric7,metric8,metric9,metric11,metric12,metric13,metric13b,
-                            metric14,metric15,metric16,metric16b,
-                            #metric17,
-                            metric18,metric19,metric20,
+                            metric14,metric15,metric16,metric16b,metric17,metric18,metric19,metric20,
                             metric21,metric22,metric23_updated,metric25b,metric26,metric27,metric28,metric28b,
-                            metric29,metric29b,metric31,
-                            #metric32_updated,
-                            metric33,metric34,metric38,
+                            metric29,metric29b,metric31,metric32_updated,metric33,metric34,metric38,
                             metric39,metric40){
   clustered_gp_and_metrics <-
     final_data_full_cats_percent_over45_5_clusters |> rename(cluster1=cluster)|>
@@ -121,7 +117,7 @@ add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final
     left_join(metric15)|>
     left_join(metric16)|>
     left_join(metric16b)|>
-   # left_join(metric17)|>
+    left_join(metric17)|>
     left_join(metric18)|>
     left_join(metric19)|>
     left_join(metric20)|>
@@ -139,7 +135,7 @@ add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final
     left_join(metric29b)|>
[…]

### 2024-04-12 _targets_old.R

@@ -1,0 +1,363 @@
+# Created by use_targets().
+
+# Load packages required to define the pipeline:
+library(targets)
+
+# library(tarchetypes) # Load other packages as needed.
+
+# Set target options:
+tar_option_set(
+  packages = c("tibble","fingertipsR","readxl","tidyverse","utils","janitor","forcats",
+               "readr","visNetwork","odbc","stringr","MLID","sf","tidygeocoder",
+               "cluster","factoextra","purrr","broom","glue","RColorBrewer","leaflet","treemapify"), # packages that your targets need to run
+  format = "rds"
+
+ 
+  # Set other options as needed.
+)
+
+# tar_make_clustermq() is an older (pre-{crew}) way to do distributed computing
+# in {targets}, and its configuration for your machine is below.
+options(clustermq.scheduler = "multiprocess")
+
+
+# Run the R scripts in the R/ folder with your custom functions:
+tar_source()
+# source("other_functions.R") # Source other scripts as needed.
+
+# Target list:
+list(
+  #Get data
+  tar_target(data_path1, "data/metric14.csv", format = "file"), # SUS data metric 14
+  tar_target(metric14,read_csv_file(data_path1)|> 
+               rename(metric14=number_ct_angiography)),
+  tar_target(data_path2, "data/metric15.csv", format = "file"), # SUS data metric 15
+  tar_target(metric15,read_csv_file(data_path2)|> 
+               rename(metric15=number_electrocardiography)),
+  tar_target(data_path3, "data/metric18.csv", format = "file"), # SUS data metric 18
+  tar_target(metric18,read_csv_file(data_path3)|> 
+               rename(metric18=cardi_op_attend)),
+  tar_target(data_path4, "data/metric19.csv", format = "file"), # SUS data metric 19
+  tar_target(metric19,read_csv_file(data_path4)|> 
+               rename(metric19=cardi_opdna)),
+  tar_target(data_path5, "data/metric20.csv", format = "file"), # SUS data metric 20
+  tar_target(metric20,read_csv_file(data_path5)|> 
+               rename(metric20=elective_pci)),
+  tar_target(data_path6, "data/metric21.csv", format = "file"), # SUS data metric 21
+  tar_target(metric21,read_csv_file(data_path6)|> 
+               rename(metric21=elective_cabg)),
+  
+  tar_target(data_path7, "data/metric22.csv", format = "file"), # SUS data metric 22
[…]

### 2024-04-12 deaths_qa.R

@@ -1,8 +1,12 @@
+library(targets)
+library(janitor)
+
 clustered_gp_and_metrics <- tar_read(clustered_gp_and_metrics)|>as_tibble()
 activity_by_type_clusters_stg6 <- tar_read(activity_by_type_clusters_stg6)|>as_tibble()
 gp_reg_pat_prac_sing_age_female<- tar_read(gp_reg_pat_prac_sing_age_female)|>as_tibble()
 gp_reg_pat_prac_sing_age_male<- tar_read(gp_reg_pat_prac_sing_age_male)|>as_tibble()
 gp_reg_pat_prac_lsoa <- tar_read(gp_reg_pat_prac_lsoa)|>as_tibble()
+clusters_for_nacr <-  tar_read(clusters_for_nacr)|>as_tibble()
 
 need <- activity_by_type_clusters_stg6|>select(cluster2,metric1_total)|>rename(chdprev_need=metric1_total)
 
@@ -88,3 +92,203 @@ death_data <- clustered_gp_and_metrics |>
   select(-white,-asian,-black,-mixed,-other)|>
   left_join(need)|>
   mutate(chd_prev_perc=chdprev_need/list_size)
+
+
+age_sex_gp_reg_pat <-gp_reg_pat_prac_sing_age_female|>
+  rbind(gp_reg_pat_prac_sing_age_male)
+
+
+all_cause_deaths_2223 <- read_csv_file("data/all_cause_deaths_2223.csv")
+
+all_cause_deaths_2223 |>
+  group_by(gpnull)|>
+  summarise(deaths=sum(num_deaths))
+
+all_cause_deaths_2223 |>
+  filter(age_at_death!="NA")|>
+  group_by(gpnull)|>
+  summarise(deaths=sum(num_deaths))
+
+all_cause_deaths_2223_data <- all_cause_deaths_2223|>
+  filter(age_at_death!="NA")|>
+  #filter(chd==1)|>
+  mutate(age_group=case_when(age_at_death<1~'<1',
+                   age_at_death>=1 & age_at_death <5 ~'1-4',
+                   age_at_death>=5 & age_at_death <10 ~'5-9',
+                   age_at_death>=10 & age_at_death <15 ~'10-14',
+                   age_at_death>=15 & age_at_death <20 ~'15-19',
+                   age_at_death>=20 & age_at_death <25 ~'20-24',
+                   age_at_death>=25 & age_at_death <30 ~'25-29',
[…]

### 2024-04-12 worked_example.R

@@ -1,7 +1,9 @@
 one_metric_example_data <- tar_read(activity_by_type_clusters_stg6) |>
   as_tibble()|>
-  select(cluster2,metric1_total,starts_with("metric29"))|>
-  rename(denominator=metric1_total,numerator=metric29_total)
+  select(cluster2,metric1_total,starts_with("metric29b"))|>
+  rename(denominator=metric1_total,numerator=metric29b_total)
+
+
 
 
 # Function to compute index of disparity
@@ -10,9 +12,9 @@ compute_id <- function(one_metric_example_data) {
   data <- one_metric_example_data|>
     mutate(upper_ci=(numerator+(sqrt(numerator)*1.96))*(1/denominator),
            lower_ci=(numerator-(sqrt(numerator)*1.96))*(1/denominator),
+           se=(numerator-upper_ci)/1.96,
+           sd=(numerator-mean(numerator))
            rate=numerator/denominator,
-           se=(rate-upper_ci)/1.96,
-           sd=(rate-mean(numerator)),
            global_rate=sum(numerator)/sum(denominator),
            diff = rate-global_rate,
            abs_diff=abs(diff*denominator))
@@ -22,79 +24,5 @@ compute_id <- function(one_metric_example_data) {
 }
 
 
-data <- one_metric_example_data
-i<-1
-n<-as.numeric(nrow(data))
-for (i in 1:n){
-  num=data$numerator[i]
-  denom=data$denominator[i]
-  rate=num/denom
-  upper_ci=(num+(sqrt(num)*1.96))*(1/denom)
-  se=(upper_ci-rate)/1.96
-  rates <- rnorm(n=1000,rate,se)
-  if(i==1)
-   { rate_clusters <- as.data.frame(rates) }
-  else 
-  { rate_clusters <- rate_clusters|> bind_cols(as.data.frame(rates))}
-  i=i+1
-}
 
-rate_clusters <- rate_clusters |>
-  rename(rates1=`rates...1`,
-         rates2=`rates...2`,
-         rates3=`rates...3`,
-         rates4=`rates...4`,
-         rates5=`rates...5`
-        )|>
-  mutate(num1=rates1*data$denominator[1],
-         num2=rates2*data$denominator[2],
[…]

### 2024-04-14 draft_report.qmd
Updated text for alteration to calculations

--- 2024-03-14 draft_report.qmd
@@
 
--   The greatest disparity along the CHD pathway is flu vaccinations for patients aged 65+ where the relative index of disparity is 16.70% (95% ci - 16.64%, 16.70%), indicating that 16.7% of activity needs to be redistributed so that rates reflect need.
+-   The greatest disparity along the CHD pathway is flu vaccinations for patients aged 65+ where the relative index of disparity is 17.14% (95% ci - 17.11%, 17.17%), indicating that 17.1% of activity needs to be redistributed so that rates reflect need.
 
@@
 
--   Out of the 33 metrics along the CHD pathway, 10 have an index of disparity of greater than 10%.
+-   Out of the 33 metrics along the CHD pathway, 11 have an index of disparity of greater than 10%.
 
@@
 |:----------:|-----------:|-----------:|:----------:|:----------:|:----------:|
-|           |                    *n* |                      *d* |                        *r=n/d*                        | *gr=𝚺n/𝚺d*  |                     *(r-gr) x d*                      |
-|     1     |              3,356,002 |                1,026,826 |                         3.27                          |    2.33     |                       -961,239                        |
-|     2     |              2,638,698 |                  924,088 |                         2.86                          |    2.33     |                       -483,540                        |
-|     3     |              1,862,430 |                  958,825 |                         1.94                          |    2.33     |                       +373,741                        |
-|     4     |                476,616 |                  478,728 |                         1.00                          |    2.33     |                       +639,873                        |
-|     5     |                332,728 |                  327,542 |                         1.02                          |    2.33     |                       +431,166                        |
-| **Total** |          **8,666,474** |            **3,716,009** |                       **2.33**                        |             |                         **0**                         |
+|           |                    *n* |                      *d* |                        *r=n/d*                        | *gr=𝚺n/𝚺d*  |                     *(gr-r) x d*                      |
+|     1     |              3,356,002 |                1,228,453 |                         2.73                          |    1.92     |                      -1,001,284                       |
+|     2     |              2,638,698 |                1,124,040 |                         2.35                          |    1.92     |                       -484,121                        |
+|     3     |              1,862,430 |                1,181,796 |                         1.58                          |    1.92     |                        402,855                        |
+|     4     |                476,616 |                  571,049 |                         0.83                          |    1.92     |                        617,979                        |
+|     5     |                332,728 |                  415,950 |                         0.80                          |    1.92     |                        464,571                        |
+| **Total** |          **8,666,474** |            **4,521,288** |                         1.92                          |             |                         **0**                         |
 

### 2024-04-14 R/chart_functions.R

@@ -367,7 +367,7 @@ rate_chart <- rate_chart_data|>
   geom_bar(stat = "identity") +
   geom_errorbar( aes(x=as.factor(cluster2), ymin=lower_ci, ymax=upper_ci), width=0.4, colour="grey", alpha=0.9, size=0.5) +
   scale_fill_manual(values = palette)+
-  scale_y_continuous(comma(rate,digits=4,format="g"))+#added this line
+  #scale_y_continuous(comma(rate,digits=4,format="g"))+#added this line
   theme_light() +
   theme(legend.position="none")+
   xlab("Cluster (1=Least diverse, 5=Most diverse)") +

### 2024-04-14 R/functions.R

@@ -123,7 +123,7 @@ return(gp_16andover_pop)
 }
 
 get_allage_list <- function(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)
-{gp_list_allage <- gp_reg_pat_prac_sing_age_male |>
+{gp_list_all_age <- gp_reg_pat_prac_sing_age_male |>
   rbind(gp_reg_pat_prac_sing_age_female)|>
   filter(age == "ALL") |>
   group_by(org_code) |>
@@ -132,7 +132,7 @@ get_allage_list <- function(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_a
   select(org_code,number_of_patients)|>
   rename(practice_code=org_code)
 
-return(gp_list_allage)
+return(gp_list_all_age)
 }
 
 

### 2024-04-14 _targets.R

@@ -91,6 +91,7 @@ list(
   tar_target(gp_over45_perc,get_over45_perc(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)),
   tar_target(gp_list_45over,get_45over_list(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)),
   tar_target(gp_16andover_pop,get_gp_16andover_pop(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)),
+  tar_target(gp_list_all_age,get_allage_list(gp_reg_pat_prac_sing_age_male,gp_reg_pat_prac_sing_age_female)),
  
   tar_target(data_path15, "data/LSOA_(2011)_to_LSOA_(2021)_to_Local_Authority_District_(2022)_Lookup_for_England_and_Wales.csv", format = "file"), #lookup
   tar_target(lsoa_lookup,read_csv_file(data_path15)),
@@ -288,7 +289,7 @@ tar_target(metric40, ncdes_data|>
 tar_target(clustered_gp_and_metrics,
            add_all_metrics(final_data_full_cats_percent_over45_5_clusters,
                            final_data_full_cats_percent_5_clusters,
-                           gp_16andover_pop,gp_list_45over,
+                           gp_16andover_pop,gp_list_45over,gp_list_all_age,
                            metric1_updated,metric2,metric5,metric6,metric7,metric8,metric9,metric11,
                            metric12,metric13,metric13b,metric14,metric15,metric16,
                            metric16b,

### 2024-04-15 draft_report.qmd
import metric 24 from csv and process

@@
 |  Cluster  | Number of vaccinations | CHD synthetic prevalence | Rate (No. of vaccinations / CHD synthetic prevalence) | Global rate | Change in number of vaccinations to match global rate |
-|:----------:|-----------:|-----------:|:----------:|:----------:|:----------:|
+|:---------:|-----------------------:|-------------------------:|:-----------------------------------------------------:|:-----------:|:-----------------------------------------------------:|
 |           |                    *n* |                      *d* |                        *r=n/d*                        | *gr=𝚺n/𝚺d*  |                     *(gr-r) x d*                      |

### 2024-04-15 R/metric_processing_functions.R

@@ -78,13 +78,25 @@ metric23_below_trimpoint <- function(metric23,metric23trimpoints){
 #   return(metric32_updated)
 # }
 
+metric24_process <- function(metric24){
+  
+  metric24_processed <- metric24 |> group_by(cluster) |>
+    summarise(started=sum(started),
+              completed=sum(completed))
+    
+    
+    return(metric24_processed)
+  
+}
+
 add_all_metrics <- function(final_data_full_cats_percent_over45_5_clusters,final_data_full_cats_percent_5_clusters,
                             gp_16andover_pop,gp_list_45over,gp_list_all_age,
                             metric1_updated,metric2,metric5,metric6,metric7,metric8,metric9,metric11,metric12,metric13,metric13b,
                             metric14,metric15,metric16,metric16b,
                             #metric17,
                             metric18,metric19,metric20,
-                            metric21,metric22,metric23_updated,metric25b,metric26,metric27,metric28,metric28b,
+                            metric21,metric22,metric23_updated,
+                            metric25b,metric26,metric27,metric28,metric28b,
                             metric29,metric29b,metric31,
                             #metric32_updated,
                             metric33,metric34,metric38,

### 2024-04-15 _targets.R

@@ -56,6 +56,10 @@ list(
   tar_target(data_path8, "data/metric23.csv", format = "file"), # SUS data metric 23
   tar_target(metric23,read_csv_file(data_path8)),
   
+  tar_target(data_path35, "data/nacr_data.csv", format = "file"), #nacr data metric 24
+  tar_target(metric24,read_csv_file(data_path35)),
+  tar_target(metric24_processed,metric24_process(metric24)),
+  
   tar_target(data_path34, "data/metric26.csv", format = "file"), # SUS data metric 26
   tar_target(metric26,read_csv_file(data_path34)|> 
                rename(metric26=num)),
@@ -416,7 +420,7 @@ tar_target(metric20_rates_per_cluster,get_rates_per_cluster(metric20_data)),
 tar_target(metric21_rates_per_cluster,get_rates_per_cluster(metric21_data)),
 tar_target(metric22_rates_per_cluster,get_rates_per_cluster(metric22_data)),
 tar_target(metric23_rates_per_cluster,get_rates_per_cluster(metric23_data)),
-#24
+#24 tar_target(metric24_rates_per_cluster,xxxxxx(metric24_data)),
 tar_target(metric25_rates_per_cluster,get_rates_per_cluster(metric25_data)),
 tar_target(metric26_rates_per_cluster,get_rates_per_cluster(metric26_data)),
 tar_target(metric27_rates_per_cluster,get_rates_per_cluster(metric27_data)),

### 2024-04-15 draft_report.qmd

--- 2024-04-14 draft_report.qmd
@@
 |  Cluster  | Number of vaccinations | CHD synthetic prevalence | Rate (No. of vaccinations / CHD synthetic prevalence) | Global rate | Change in number of vaccinations to match global rate |
-|:----------:|-----------:|-----------:|:----------:|:----------:|:----------:|
+|:---------:|-----------------------:|-------------------------:|:-----------------------------------------------------:|:-----------:|:-----------------------------------------------------:|
+
 |           |                    *n* |                      *d* |                        *r=n/d*                        | *gr=𝚺n/𝚺d*  |                     *(gr-r) x d*                      |

### 2024-04-16 draft_report.qmd
Incorporated cardiac rehab data to project

--- 2024-04-14 draft_report.qmd
@@
 ```
-
-::: {.callout-note icon="false"}
-## Cardiac rehabilitation metrics
-
-Cardiac rehabilitation metrics will be added once they become available from NACR (expected before 31/3/24)
-:::
 
@@
 
--   All 4 tertiary prevention metrics follow the same pattern of activity rates with the highest rates being in the least diverse cluster (cluster 1), followed by clusters 2 and then cluster 5 (the most diverse) and cluster 3.
+-   All the tertiary prevention metrics, except waiting time for elective PCI/CABG, follow broadly the same pattern of activity rates with the highest rates being in the least diverse cluster (cluster 1), followed by cluster 2 and then clusters 5 (the most diverse) and 3.
 Cluster 4 has the lowest activity rates.
+-   The rate for waiting time for elective PCI/CABG is greatest in the most diverse cluster (cluster 5) and lowest in cluster 4, although overall there is little variation between the clusters.
 :::
@@
 |  Cluster  | Number of vaccinations | CHD synthetic prevalence | Rate (No. of vaccinations / CHD synthetic prevalence) | Global rate | Change in number of vaccinations to match global rate |
-|:----------:|-----------:|-----------:|:----------:|:----------:|:----------:|
+|:---------:|----------:|----------:|:------------:|:---------:|:------------:|
 |           |                    *n* |                      *d* |                        *r=n/d*                        | *gr=𝚺n/𝚺d*  |                     *(gr-r) x d*                      |

### 2024-04-16 R/chart_functions.R

@@ -39,6 +39,8 @@ get_rel_iod_chart <- function(chart_iod_data){
                                       metric=="metric21_rel_iod"~"Elective CABG",
                                       metric=="metric22_rel_iod"~"Waiting time for elective PCI / CABG",
                                       metric=="metric23_rel_iod"~"PCI/CABG disch < trimpoint",
+                                      metric=="metric24a_rel_iod"~"Cardiac rehabilitation started",
+                                      metric=="metric24b_rel_iod"~"Cardiac rehabilitation completed",
                                       metric=="metric25_rel_iod"~"BP reading < 140/90 (<80 yrs CHD)",
                                       metric=="metric26_rel_iod"~"Readmit<30 days of PCI / CABG",
                                       metric=="metric27_rel_iod"~"Emergency admissions for CHD",
@@ -56,6 +58,7 @@ get_rel_iod_chart <- function(chart_iod_data){
                                                                      "65+ flu vaccination","<65 at risk flu vaccination",
                                                                      "Ref to OP cardiology (new)", "Cardiology outpatient DNAs","Elective PCI",
                                                                      "Elective CABG","Waiting time for elective PCI / CABG","PCI/CABG disch < trimpoint",
+                                                                     "Cardiac rehabilitation started","Cardiac rehabilitation completed",
                                                                      "BP reading < 140/90 (<80 yrs CHD)","Readmit<30 days of PCI / CABG",
                                                                      "Emergency admissions for CHD","CHD Hospital Deaths","CHD Hospital Deaths <75", 
                                                                      "CHD Deaths","CHD Deaths <75")))) |>
@@ -85,7 +88,8 @@ get_rel_iod_chart <- function(chart_iod_data){
                                      metric == 'metric21_rel_iod' ~ '3°c prevent',
                                      metric == 'metric22_rel_iod' ~ '3°c prevent',                              
                                      metric == 'metric23_rel_iod' ~ '3°c prevent',   
-                                     # 24   
+                                     metric == 'metric24a_rel_iod' ~ '3°c prevent',  
+                                     metric == 'metric24b_rel_iod' ~ '3°c prevent', 
                                      metric == 'metric25_rel_iod' ~ 'Inter. outc.',                                 
[…]

### 2024-04-16 R/metric_processing_functions.R

@@ -288,9 +288,9 @@ process_metrics <-function(clustered_gp_and_metrics){
 
 calc_iod_rate <-function(activity_by_type_clusters_stg1,metric24_processed){
   activity_by_type_clusters_stg2 <- activity_by_type_clusters_stg1 |>
-  #  left_join(metric24_processed)|>
-   # rename(metric24a_num_total=started,
-    #       metric24b_num_total=completed)|>
+    left_join(metric24_processed)|>
+    rename(metric24a_num_total=started,
+           metric24b_num_total=completed)|>
     mutate(metric2_rate=metric2_num_total/list_size_total,
            metric5_rate=metric5_num_total/list_size_total,
            metric6_rate=metric6_total/list_size_total,
@@ -311,8 +311,8 @@ calc_iod_rate <-function(activity_by_type_clusters_stg1,metric24_processed){
            metric21_rate=metric21_total/metric1_total,
            metric22_rate=metric22_total/metric1_total,
            metric23_rate=metric23_total/metric1_total,
-    #       metric24a_rate=metric24a_num_total/metric1_total,
-     #      metric24b_rate=metric24b_num_total/metric1_total,
+           metric24a_rate=metric24a_num_total/metric1_total,
+           metric24b_rate=metric24b_num_total/metric1_total,
            metric25_rate=metric25_num_total/metric1_total, 
            metric26_rate=metric26_total/metric1_total,
            metric27_rate=metric27_total/metric1_total,
@@ -365,10 +365,10 @@ calc_iod_rate <-function(activity_by_type_clusters_stg1,metric24_processed){
            metric22_lower_ci=(metric22_total-(sqrt(metric22_total*1.96)))*(1/metric1_total),
            metric23_upper_ci=(metric23_total+(sqrt(metric23_total*1.96)))*(1/metric1_total),
            metric23_lower_ci=(metric23_total-(sqrt(metric23_total*1.96)))*(1/metric1_total),
- #          metric24a_upper_ci=(metric24a_num_total+(sqrt(metric24a_num_total*1.96)))*(1/metric1_total),
-  #         metric24a_lower_ci=(metric24a_num_total-(sqrt(metric24a_num_total*1.96)))*(1/metric1_total),
-   #        metric24b_upper_ci=(metric24b_num_total+(sqrt(metric24b_num_total*1.96)))*(1/metric1_total),
[…]

### 2024-04-16 draft_report.qmd
Edited text related to and removed regional and ICB versions of metrics 24a and 24b

--- 2024-04-15 draft_report.qmd
@@
 
--   Out of the 33 metrics along the CHD pathway, 11 have an index of disparity of greater than 10%.
+-   Out of the 33 metrics along the CHD pathway, 13 have an index of disparity of greater than 10%.
 
--   In general (but not exclusively) these greater levels of disparity occur in the disease identification and secondary prevention pathway domains and consequently the outcome metrics related to deaths from CHD.
+-   In general (but not exclusively) these greater levels of disparity occur in the disease identification and secondary and tertiary prevention pathway domains and consequently the outcome metrics related to deaths from CHD.
 :::

### 2024-04-16 R/chart_functions.R

@@ -9,6 +9,12 @@ get_rel_iod_data_for_chart <- function(activity_by_type_clusters_stg6){
 return(chart_iod_data)
 }
 
+remove_metric24_for_regional_charts <- function(data){
+  data <- data|>
+    filter(metric!=("metric24a_rel_iod"))|>
+    filter(metric!=("metric24b_rel_iod"))
+  return(data)
+}
 
 get_rel_iod_chart <- function(chart_iod_data){
 

### 2024-04-16 _targets.R

@@ -518,6 +518,7 @@ tar_target(regional_charts,region_clustered_gp_and_metrics |>
              map(\(df) calc_iod_diff(df) )|>
              map(\(df) calc_abs_iod(df) )|>
              map(\(df) get_rel_iod_data_for_chart(df) )|>
+             map(\(df) remove_metric24_for_regional_charts(df))|>
              map(\(df) get_rel_iod_chart(df) )),
 tar_target(region_cluster_chart_data,get_region_cluster_chart_data(region_clustered_gp_and_metrics)),
 tar_target(region_cluster_charts,region_cluster_chart_data|>
@@ -538,6 +539,7 @@ tar_target(icb_charts,icb_clustered_gp_and_metrics |>
              map(\(df) calc_iod_diff(df) )|>
              map(\(df) calc_abs_iod(df) )|>
              map(\(df) get_rel_iod_data_for_chart(df) )|>
+             map(\(df) remove_metric24_for_regional_charts(df))|>
              map(\(df) get_rel_iod_chart(df) ))
 
 )

### 2024-04-17 draft_report.qmd
Updated methodology relating to NACR data

--- 2024-04-16 draft_report.qmd
@@
 
-This data was supplied by NACR.
+This data was supplied by NACR from an extract taken in March 2024.
 The data wasn’t available by GP Practice, so an alternative method was followed.
-Practices were allocated to a cluster and data was then supplied aggregated into clusters.
-In some cases, the data had to be redacted due to small numbers in the category.
-For the purposes of this analysis the redacted data item was reassigned an average value.
-The data used relates to the 2021 & 2022 calendar years, which are the most recent complete years, and following changes in the way rehabilitation has been delivered due to the Covid pandemic, most closely reflects the rehabilitation model now in place.
+NACR was supplied with a list of practices, ICBs and the cluster to which they had been allocated by the Strategy Unit (SU).
+The data was then supplied to the SU aggregated into clusters and ICBs.
+The data relates to patients who started rehabilitation in 2021 or 2022 calendar years and were ACS or HF according to the NHS England reporting.
+These two years are the most recent complete years, and following changes in the way rehabilitation has been delivered due to the Covid pandemic, most closely reflect the rehabilitation model now in place.
+Patients needed to have GP Practice code recorded to be matched to a cluster.
+Those that didn't equated to 42637 patients that had either NO GP CODE completed, no matching GP Code or GMC Number added, representing 48% of patient records.
+Within the data included in this analysis, there are 11 clusters across 9 ICBs that have no rehabilitation data despite there being practices in the cluster.
+Due to these high levels of missing data the two cardiac rehabilitation metrics have been included in the presentations of the index of disparity and activity rates at a national level, but excluded from the regional and ICB level presentations.
 

### 2024-04-17 draft_report.qmd
Add age/sex profile of clusters and renumber tables

@@
 
+## Age, Sex and Deprivation Profile of Clusters
+
+describe
+
+```{r}
+gp_reg_pat_prac_sing_age_female<- tar_read(gp_reg_pat_prac_sing_age_female)|>as_tibble()
+gp_reg_pat_prac_sing_age_male<- tar_read(gp_reg_pat_prac_sing_age_male)|>as_tibble()
+clusters_for_nacr <- tar_read(clusters_for_nacr)|>as_tibble()
+
+females_gp_reg_pat <- gp_reg_pat_prac_sing_age_female |>
+  filter(age != "ALL") |>
+  mutate(age=as.numeric(case_when(age=="95+" ~ 95,
+                                  .default = as.numeric(age))))|>
+  mutate(age_group=case_when(age<18~'<18',
+                             age>=18 & age <45 ~'18-44',
+                             age>=45 & age <65 ~'45-64',
+                             age>=65 & age <75 ~'65-74',
+                             age>=75 ~'75+'))|>
+  select(-age)|>
+  mutate(sex=2)
+
+males_gp_reg_pat <- gp_reg_pat_prac_sing_age_male |>
+  filter(age != "ALL") |>
+  mutate(age=as.numeric(case_when(age=="95+" ~ 95,
+                                  .default = as.numeric(age))))|>
+  mutate(age_group=case_when(age<18~'<18',
+                             age>=18 & age <45 ~'18-44',
+                             age>=45 & age <65 ~'45-64',
+                             age>=65 & age <75 ~'65-74',
+                             age>=75 ~'75+'))|>
+  select(-age)|>
+  mutate(sex=1)
+
+all_gp_reg_pat <- males_gp_reg_pat|>
+  rbind(females_gp_reg_pat)|>
+  rename(gp_practice_code=org_code)|>
+  group_by(sex,age_group,gp_practice_code)|>
+  summarise(number_of_patients=sum(number_of_patients))|>
+  ungroup()|>
+  left_join(clusters_for_nacr|>select(gp_practice_code,cluster))|>
+  group_by(sex,age_group,cluster)|>
+  summarise(number_of_patients=sum(number_of_patients))|>
+  ungroup()|>
+  mutate(sex_name=case_when(sex==1~'Male',
+                            sex==2~'Female'))|>
+  select(-sex)|>
+  group_by(cluster)|>
+  mutate(patients_total_cluster=sum(number_of_patients))|>
+  mutate(perc=number_of_patients/patients_total_cluster)|>
+  select(-number_of_patients,-patients_total_cluster)
+
+all_gp_reg_pat |>
+    pivot_wider(names_from=c(cluster,sex_name),
+                names_sep = "_",
+                values_from = perc
+                )|>
+    select(age_group, starts_with("1"), starts_with("2"), starts_with("3"),starts_with("4"),starts_with("5")) |>
+  arrange(factor(age_group, 
+                 levels = c("<18","18-44","45-64","65-74","75+")))|>
+    gt(rowname_col = "age_group")|>
+    tab_spanner_delim(delim = "_") |>
+    tab_stubhead(label=md("**Age Group**"))|>
+    tab_options(heading.title.font.size = 18,
+                heading.title.font.weight = "bolder",
+                column_labels.font.weight = "bold")|>
+    fmt_percent(decimals = 2)|>
+    tab_header(title = md("**Table 2** - *Age / Sex Breakdown by Cluster*"))
+
+```
+
+Table 3 = imd
+
 ## Clustered GP Practices
@@
 
-*Table 2 - achieving equity via zero-sum redistribution*
+*Table 4 - achieving equity via zero-sum redistribution*
 
 |  Cluster  | Number of vaccinations | CHD synthetic prevalence | Rate (No. of vaccinations / CHD synthetic prevalence) | Global rate | Change in number of vaccinations to match global rate |
-|:---------:|----------:|----------:|:------------:|:---------:|:------------:|
+|:----------:|-----------:|-----------:|:----------:|:----------:|:----------:|
 |           |                    *n* |                      *d* |                        *r=n/d*                        | *gr=𝚺n/𝚺d*  |                     *(gr-r) x d*                      |
@@
     tab_header(
-    title = md("**Table 3** - *Metric definitions and data sources*"))|>
+    title = md("**Table 5** - *Metric definitions and data sources*"))|>
     cols_label(name_metric=md("**Metric**"),
@@
 
-SUS data was extracted using Transact-SQL and using relevant OPCS, ICD10, HRG, and Treatment function codes as detailed in table 2 above.
+SUS data was extracted using Transact-SQL and using relevant OPCS, ICD10, HRG, and Treatment function codes as detailed in table 5 above.
 
@@
 Patients needed to have GP Practice code recorded to be matched to a cluster.
-Those that didn't equated to 42637 patients that had either NO GP CODE completed, no matching GP Code or GMC Number added, representing 48% of patient records.
+Those that didn't equated to 42,637 patients that had either NO GP CODE completed, no matching GP Code or GMC Number added, representing 48% of patient records.
 Within the data included in this analysis, there are 11 clusters across 9 ICBs that have no rehabilitation data despite there being practices in the cluster.
@@
 
-Table 3 shows the number of GP practices assigned to each of the 5 clusters and the characteristics of the clusters are described in the main body of the report above.
+Table 6 shows the number of GP practices assigned to each of the 5 clusters and the characteristics of the clusters are described in the main body of the report above.
 
@@
   gt()|>
-    tab_header(title = md("**Table 4** - *Number of practices per cluster*"))|>
+    tab_header(title = md("**Table 6** - *Number of practices per cluster*"))|>
     cols_label(cluster=md("**Cluster**"),

### 2024-04-17 deaths_qa.R

@@ -290,5 +290,32 @@ all_gp_reg_pat <- males_gp_reg_pat|>
   select(-sex)|>
   group_by(cluster)|>
   mutate(patients_total_cluster=sum(number_of_patients))|>
-  mutate(perc=number_of_patients/patients_total_cluster)
+  mutate(perc=number_of_patients/patients_total_cluster)|>
+  select(-number_of_patients,-patients_total_cluster)
+  
+  all_gp_reg_pat|>
+  gt(rowname_col = "age_group")|>
+  tab_stubhead(label=md("**Age Group**")) |>
+  tab_spanner(label = "Cluster", columns = matches("cluster")) |>
+  cols_label(
+    perc = md("**Percentage**"),
+    sex_name = md("**Sex**"))|>
+  fmt_number(columns = "perc", decimals = 2)
 
+library(gt)
+  
+  all_gp_reg_pat |>
+    pivot_wider(names_from=c(cluster,sex_name),
+                names_sep = "_",
+                values_from = perc
+                )|>
+    select(age_group, starts_with("1"), starts_with("2"), starts_with("3"),starts_with("4"),starts_with("5")) |>
+    gt(rowname_col = "age_group")|>
+    tab_spanner_delim(delim = "_") |>
+    tab_stubhead(label=md("**Age Group**"))|>
+    tab_options(heading.title.font.size = 18,
+                heading.title.font.weight = "bolder",
+                column_labels.font.weight = "bold")|>
+    fmt_percent(decimals = 2)|>
+    tab_header(title = "Age / Sex Breakdown by Cluster")
+  

### 2024-04-18 draft_report.qmd
Adding IMD analysis of clustered practices

@@
 
-describe
+Table 2 shows that the most diverse clusters are younger and the least diverse clusters are older.
+In particular almost one quarter of cluster 5 are under 18, compared to 19% of cluster 1 and 12% of cluster 1 are aged 75+ compared to only 4% of clusters 4 and 5.
 
@@
 Table 3 = imd
+
+```{r}
+source("R/get_imd_data_by_gpprac.R")
+
+patientweighted_practice_imd |>
+  mutate(gp_imd_quantile = case_when(gp_imd_decile<=2 ~ 1,
+                                     gp_imd_decile==3 ~2,
+                                     gp_imd_decile==4 ~2,
+                                     gp_imd_decile==5 ~3,
+                                     gp_imd_decile==6 ~3,
+                                     gp_imd_decile==7 ~4,
+                                     gp_imd_decile==8 ~4,
+                                     gp_imd_decile>=9 ~5,
+                                     .default = 0
+  ))|>
+  group_by(gp_imd_quantile,cluster)|>
+  summarise(patients=sum(patients))|>
+  gt()
+```
 

### 2024-04-18 R/get_imd_data_by_gpprac.R

@@ -1,0 +1,43 @@
+library(readxl)
+library(tidyr)
+library(tidyverse)
+
+#Get the IMD Deciles for each GP Practice
+
+GPpatients_by_LSOA <- read_excel(here::here("data", "patients by LSOA to GP.xlsx"))
+gp_reg_pat_prac_lsoa <-tar_read(gp_reg_pat_prac_lsoa)|>as_tibble()
+
+imd2019 <- read_excel(here::here("data", "IMD2019.xlsx"))
+
+
+#practice_to_CCGSTP_lookup <- read_excel(here::here("data", "gp-reg-pat-prac-map-jul22.xlsx"))
+
+
+#appends LSOA IMD score to GPpatients_by_LSOA
+gp_reg_pat_prac_lsoa<- gp_reg_pat_prac_lsoa%>% 
+  left_join(select(imd2019,`LSOA code (2011)`,`Index of Multiple Deprivation (IMD) Score`), 
+            by = c("lsoa_code" = "LSOA code (2011)"))
+
+#creates new field of LSOA IMDscore multiplied by the number of patients from that LSOA
+gp_reg_pat_prac_lsoa<-gp_reg_pat_prac_lsoa|>
+  mutate(scoretimespoplsoa = `Index of Multiple Deprivation (IMD) Score`*number_of_patients)
+
+#groups data by Practice and sums number of patients and IMD score
+patientweighted_practice_imd<-gp_reg_pat_prac_lsoa|>
+  filter(!is.na(scoretimespoplsoa))|>
+  group_by (practice_code)|>
+  summarise(total_imd = sum(scoretimespoplsoa),patients = sum(number_of_patients)) |>
+  arrange()|>
+  
+  #then calculates the average IMD (total IMD score/total Patients) 
+  
+  mutate(average_imd = total_imd/patients)|>
+  #then calculates IMDDecile 
+  mutate(gp_imd_decile = ntile(-average_imd, 10))
+  #adds in organisational data
+  # left_join(select(practice_to_CCGSTP_lookup,`CCG_CODE`,`CCG_NAME`,`STP_CODE`,`STP_NAME`,`COMM_REGION_NAME`,`ICB22ons`,`ICB22`,`ICB22NM`,`New Sub-ICB location name`,`Reg22NM`,`PRACTICE_CODE`), 
+  #           by = c("PRACTICE_CODE" = "PRACTICE_CODE"))
+
[…]

### 2024-04-18 draft_report.qmd
Adding IMD analysis of clustered practices

--- 2024-04-17 draft_report.qmd
@@
 
-Table 3 = imd
+Table 3 shows that ...
+
+```{r}
+source("R/get_imd_data_by_gpprac.R")
+
+patientweighted_practice_imd |>
+  mutate(gp_imd_quantile = case_when(gp_imd_decile<=2 ~ 1,
+                                     gp_imd_decile==3 ~2,
+                                     gp_imd_decile==4 ~2,
+                                     gp_imd_decile==5 ~3,
+                                     gp_imd_decile==6 ~3,
+                                     gp_imd_decile==7 ~4,
+                                     gp_imd_decile==8 ~4,
+                                     gp_imd_decile>=9 ~5,
+                                     .default = 0
+  ))|>
+  group_by(gp_imd_quantile,cluster)|>
+  summarise(patients=sum(patients))|>
+  ungroup()|>
+  group_by(cluster)|>
+  mutate(cluster_total=sum(patients),
+         perc=patients/cluster_total)|>
+  gt()
+```
 

### 2024-04-18 R/get_imd_data_by_gpprac.R

@@ -1,0 +1,44 @@
+library(readxl)
+library(tidyr)
+library(tidyverse)
+
+#Get the IMD Deciles for each GP Practice
+
+GPpatients_by_LSOA <- read_excel(here::here("data", "patients by LSOA to GP.xlsx"))
+gp_reg_pat_prac_lsoa <-tar_read(gp_reg_pat_prac_lsoa)|>as_tibble()
+clusters_for_nacr <- tar_read(clusters_for_nacr)|>as_tibble()
+
+imd2019 <- read_excel(here::here("data", "IMD2019.xlsx"))
+
+
+#practice_to_CCGSTP_lookup <- read_excel(here::here("data", "gp-reg-pat-prac-map-jul22.xlsx"))
+
+
+#appends LSOA IMD score to GPpatients_by_LSOA
+gp_reg_pat_prac_lsoa<- gp_reg_pat_prac_lsoa%>% 
+  left_join(select(imd2019,`LSOA code (2011)`,`Index of Multiple Deprivation (IMD) Score`), 
+            by = c("lsoa_code" = "LSOA code (2011)"))
+
+#creates new field of LSOA IMDscore multiplied by the number of patients from that LSOA
+gp_reg_pat_prac_lsoa<-gp_reg_pat_prac_lsoa|>
+  mutate(scoretimespoplsoa = `Index of Multiple Deprivation (IMD) Score`*number_of_patients)
+
+#groups data by Practice and sums number of patients and IMD score
+patientweighted_practice_imd<-gp_reg_pat_prac_lsoa|>
+  filter(!is.na(scoretimespoplsoa))|>
+  group_by (practice_code)|>
+  summarise(total_imd = sum(scoretimespoplsoa),patients = sum(number_of_patients)) |>
+  arrange()|>
+  
+  #then calculates the average IMD (total IMD score/total Patients) 
+  
+  mutate(average_imd = total_imd/patients)|>
+  #then calculates IMDDecile 
+  mutate(gp_imd_decile = ntile(-average_imd, 10))
+  #adds in organisational data
+  # left_join(select(practice_to_CCGSTP_lookup,`CCG_CODE`,`CCG_NAME`,`STP_CODE`,`STP_NAME`,`COMM_REGION_NAME`,`ICB22ons`,`ICB22`,`ICB22NM`,`New Sub-ICB location name`,`Reg22NM`,`PRACTICE_CODE`), 
[…]

### 2024-04-23 draft_report.qmd
Add imd cluster details

@@
 
+```{r}
+source("R/get_imd_data_by_gpprac.R")
+
+patientweighted_practice_imd |>
+  mutate(gp_imd_quantile = case_when(gp_imd_decile<=2 ~ 1,
+                                     gp_imd_decile==3 ~2,
+                                     gp_imd_decile==4 ~2,
+                                     gp_imd_decile==5 ~3,
+                                     gp_imd_decile==6 ~3,
+                                     gp_imd_decile==7 ~4,
+                                     gp_imd_decile==8 ~4,
+                                     gp_imd_decile>=9 ~5,
+                                     .default = 0
+  ))|>
+  group_by(gp_imd_quantile,cluster)|>
+  summarise(patients=sum(patients))|>
+  gt()
+```
+
 ## Clustered GP Practices

### 2024-04-23 R/get_imd_data_by_gpprac.R

@@ -6,7 +6,10 @@ library(tidyverse)
 
 GPpatients_by_LSOA <- read_excel(here::here("data", "patients by LSOA to GP.xlsx"))
 gp_reg_pat_prac_lsoa <-tar_read(gp_reg_pat_prac_lsoa)|>as_tibble()
+<<<<<<< HEAD
 clusters_for_nacr <- tar_read(clusters_for_nacr)|>as_tibble()
+=======
+>>>>>>> 49354d3e4cd309302906afba47f11c705a0d55ee
 
 imd2019 <- read_excel(here::here("data", "IMD2019.xlsx"))
 

### 2024-04-23 draft_report.qmd
Style table of imd by cluster

--- 2024-04-18 draft_report.qmd
@@
 
-Table 3 = imd
+Table 3 shows that
+
+```{r}
+source("R/get_imd_data_by_gpprac.R")
+
+patientweighted_practice_imd |>
+  mutate(gp_imd_quantile = case_when(gp_imd_decile<=2 ~ 1,
+                                     gp_imd_decile==3 ~2,
+                                     gp_imd_decile==4 ~2,
+                                     gp_imd_decile==5 ~3,
+                                     gp_imd_decile==6 ~3,
+                                     gp_imd_decile==7 ~4,
+                                     gp_imd_decile==8 ~4,
+                                     gp_imd_decile>=9 ~5,
+                                     .default = 0
+  ))|>
+  group_by(gp_imd_quantile,cluster)|>
+  summarise(patients=sum(patients))|>
+  ungroup()|>
+  group_by(cluster)|>
+  mutate(cluster_total=sum(patients),
+         perc=patients/cluster_total)|>
+  select(-patients,-cluster_total)|>
+  pivot_wider(names_from=c(cluster),
+              names_sep = "_",
+              values_from = perc
+  )|>
+  select(gp_imd_quantile, starts_with("1"), starts_with("2"), starts_with("3"),starts_with("4"),starts_with("5")) |>
+  arrange(factor(gp_imd_quantile, 
+                 levels = c("1","2","3","4","5")))|>
+
+  gt(rowname_col = "gp_imd_quantile")|>
+  tab_stubhead(label=md("**IMD Quantile**"))|>
+  tab_options(heading.title.font.size = 18,
+              heading.title.font.weight = "bolder",
+              column_labels.font.weight = "bold")|>
+    tab_stubhead(label=md("**IMD Quantile**"))|>
+  fmt_percent(columns=c(starts_with("1"), starts_with("2"), starts_with("3"),starts_with("4"),starts_with("5")),decimals = 2)|>
+  tab_header(title = md("**Table 3** - *Index of Multiple Deprivation Breakdown by Cluster*"))
+```
 

### 2024-04-23 draft_report.qmd
Add column spanner headings to gt tables

@@
     tab_spanner_delim(delim = "_") |>
+    tab_spanner(label = "Cluster", columns = c(starts_with("1"), starts_with("2"), starts_with("3"),starts_with("4"),starts_with("5")))|>
     tab_stubhead(label=md("**Age Group**"))|>
@@
 
-Table 3 shows that
+Table 3 shows that the most diverse cluster has the greatest proportion of patients in practices in the most deprived quantile.
+Cluster 2 has the greatest proportion of patients in practices in the least deprived quantile.
 
@@
     tab_stubhead(label=md("**IMD Quantile**"))|>
+  tab_spanner(label = "Cluster", columns = c(starts_with("1"), starts_with("2"), starts_with("3"),starts_with("4"),starts_with("5")))|>
   fmt_percent(columns=c(starts_with("1"), starts_with("2"), starts_with("3"),starts_with("4"),starts_with("5")),decimals = 2)|>
   tab_header(title = md("**Table 3** - *Index of Multiple Deprivation Breakdown by Cluster*"))
-```
-
-```{r}
-source("R/get_imd_data_by_gpprac.R")
-
-patientweighted_practice_imd |>
-  mutate(gp_imd_quantile = case_when(gp_imd_decile<=2 ~ 1,
-                                     gp_imd_decile==3 ~2,
-                                     gp_imd_decile==4 ~2,
-                                     gp_imd_decile==5 ~3,
-                                     gp_imd_decile==6 ~3,
-                                     gp_imd_decile==7 ~4,
-                                     gp_imd_decile==8 ~4,
-                                     gp_imd_decile>=9 ~5,
-                                     .default = 0
-  ))|>
-  group_by(gp_imd_quantile,cluster)|>
-  summarise(patients=sum(patients))|>
-  gt()
 ```

### 2024-04-23 findings.qmd
update

--- 2024-03-12 findings.qmd
@@
 ---
-title: "Visualising CHD Pathway Inequalities by ethnicity - DRAFT"
+title: "Inequalities by ethnicity in the CHD pathway"
 lang: en-GB
@@
 date-format: "YYYY-MM-DD"
-title-block-banner: "#f9bf07"
-title-block-banner-color: "#333739"
 format:
   html:
-    self-contained: true
     grid:
-      sidebar-width: 200px
-      body-width: 950px
-      margin-width: 150px
+      sidebar-width: 250px
+      body-width: 800px
+      margin-width: 250px
       gutter-width: 1.5rem
@@
   chunk_output_type: console
-
-css: styles.css
 ---
@@
 library(targets)
-library(gt)
+
 library(tidyverse)
-library(grid)
-library(gridExtra)
-
 ```
 
-# Introduction
-
-In 2022, the Strategy Unit worked with the British Heart Foundation (BHF) to explore ways of visualising socio-economic inequalities as they emerge through the coronary heart disease (CHD) pathway.
-We made these visualisations available so that health and care staff can understand where on the pathway socio-economic inequalities emerge and at which points they are moderated or exacerbated.
-The report[^1] and web-based tool[^2] are available via the BHF and Strategy Unit websites.
-
-[^1]: <https://www.strategyunitwm.nhs.uk/publications/socio-economic-inequalities-coronary-heart-disease>
-
-[^2]: <https://www.bhf.org.uk/icb-tool>
-
-Whilst this initial analysis was focused on inequalities between socio-economic groups, BHF asked the Strategy Unit if we might explore the feasibility of assessing inequalities across other dimensions of inequality.
-This report explores inequalities by ethnicity along the coronary heart disease (CHD) pathway.
-
-The objectives of the report are to:
-
-1.  Set out the methods by which the Strategy Unit have sought to represent CHD pathway inequalities by ethnicity.
-
-2.  Quantify and illustrate the inequalities by ethnicity over the disease progression and treatment pathway for CHD.
-
-The analysis has been conducted by the Strategy Unit on behalf of the British Heart Foundation.
-
-# Inequities in healthcare
-
-The term ‘inequities’ is used to describe unjustifiable differences in rates of access between subgroups.
-An equity analysis must control for levels of need within each population subgroup.
-Having done this, an equitable distribution of services is one where rates of access to a service or population follow the distribution of need, such that a patient with a given level of need in one subgroup has the same chance of accessing a service as their counterparts with a similar level of need in other subgroups.
-This is the standard that the NHS seeks to achieve.
-Assessing equity is challenging.
-Further detail about inequalities and inequities in healthcare can be found in a previous Strategy Unit report[^3]
-
-[^3]: <https://www.strategyunitwm.nhs.uk/publications/socio-economic-inequalities-access-planned-hospital-care-causes-and-consequences>
-
-In our previous work, visualising socio-economic inequalities, our units of analysis were GP practices.
-The metrics of interest, risk factors, primary prevention interventions, secondary care procedures, outcomes etc, are readily available at GP practice level and there are established methods of assigning GP practices to deciles of deprivation.
-Having calculated a metric value for each decile of GP practices, we used the relative index of inequality (RII) to estimate the scale and direction of inequality.
-The RII can only be used when the dimension of inequality can be expressed as a set of order groups.
-
-Assessing inequalities across ethnic groups is more challenging, since unlike deprivation, ethnicity cannot be expresseed as a set of ordered groups.
-The distribution of patients across ethnic groups is unique to each practice.
-There is no meaningful way to numerically aggregate these distributions into a single variable (e.g.
-% BME) without significant loss of information.
-An alternative method of grouping practices according to the distribution of its patients over ethnic group was required.
-We have used k-medoids clustering to assign practices to one of a small number of groups such that practices within a group have similar distributions of patients over ethnic groups.
-K-medoids is an established, and commonly used unsupervised machine learning technique.
-This clustering approach required data on the ethnicity of a practices registered population.
-We imputed it based on the Lower Super Output Area (LSOA) of residence of GP registrants and the ethnicity distribution of patients living in these LSOAs using the Census 2021 data.
-
-Ethnicity, unlike deprivation, is not an ordered variable.
-The approach in our previous work used the relative index of inequality to measure the degree of inequality.
-This measure relies on the ordered quality of the socio-economic deprivation variable.
-To handle the categorical nature of the ethnicity variable we have used the relative index of disparity to indicate the extent to which the rate of an activity or event varies across groups.
-The index estimates the proportion of events (e.g., admissions) that would need to be redistributed between clusters in order that event rates follow levels of need.
-Further detailed explanations of the methods used in the analysis can be found in the appendix.
-
-# CHD Metrics
-
-This report quantifies and illustrates levels of inequity across 35 metrics at various points along the continuum of coronary heart disease progression and over a typical treatment pathway.
-They are shown in the table below grouped by domain (risk factors, risk factor identification, primary prevention, disease identification, secondary prevention, tertiary prevention, intermediate and full outcomes), which represent the various stages along the pathway.
-Full definitions and data sources for each metric are included in the appendix.
-
-*Table 1 - coronary heart disease pathway metrics*
+## Clustered GP Practices
 
 ```{r}
-library(gt)
-metrics_table <- tibble(metric_domain = c("Need","Risk factors","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Risk factor identification","Primary prevention","Primary prevention","Primary prevention","Primary prevention","Disease identification","Disease identification","Disease identification","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Secondary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Tertiary prevention","Intermediate outcome","Intermediate outcome","Intermediate outcome","Full outcomes","Full outcomes","Full outcomes","Full outcomes"),
-  name_metric=c("CHD synthetic prevalence estimates",
-                                      "Smoking synthetic prevalence estimates",
-                                      "Smoking register",
-                                      "Obesity register",
-                                      "Diabetes register",
-                                      "Depression register",
-                                      "The percentage of patients aged 45 or over who have a record of blood pressure in the preceding 5 years",
-                                      "CVD risk register",
-                                      "Percentage of patients aged 18 and over with GP recorded CVD (narrow definition), who are currently treated with lipid lowering therapy",
-                                      "Percentage of patients aged 18 and over, with no GP recorded CVD and a GP recorded QRISK score of 10% or more, CKD (G3a to G5), T1 diabetes (aged 40 and over) or T2 diabetes aged 60 and over, who are currently treated with lipid lowering therapy.",
-                                      "Smoking cessation support offered",
-                                      "Exception reporting for Smoking cessation support offered",
-
-                                      "CHD register",
-                                      "CT angiography",
-                                      "Electrocardiography",
-                                      "Aspirin, anti-platelet or anti-coagulent",
-                                      "Exception reporting for Aspirin, anti-platelet or anti-coagulent",
-                                      "Flu vaccination",
-                                      "Exception reporting for Flu vaccination",
-                                      "Percentage of patients aged 65 or over
-who received a seasonal influenza vaccination between 1 September 2022 and 31 March 2023",
-                                      " Percentage of patients aged 18 to 64
-years and in a clinical at-risk group who received
-a seasonal influenza vaccination between 1
-September 2022 and 31 March 2023",
-"Referral to cardiology (First outpatient)",
-                                      "Cardiology outpatient DNAs",
-"Elective PCI","Elective CABG","Waiting time for elective PCI / CABG","Elective PCI / CABG patients discharged before trimpoint",
-"Cardiac rehabilitation - Started", "Cardiac rehabilitation - Completed","BP reading < 140/90","Readmission within 30 days of elective PCI / CABG","Emergency admissions for CHD","Deaths in hospital from CHD","Deaths in hospital from CHD <75","Deaths from CHD","Deaths from CHD <75"
-                                      )
-                        )
-
-metrics_table |>
-    gt()|>
-  cols_hide(metric_domain)|>
- # tab_stubhead(label = md("**Domain**"))|>
-    tab_header(
-    title = md("**Table 1** - *Coronary heart disease pathway metrics*")
-  )|>
-   tab_row_group(
-     id="Need",
-    label = md("**Need**"),
-    rows = 1
-  ) |>
-  tab_row_group(
-    id="Risk factors",
-    label = md("**Risk factors**"),
-    rows = 2
-  ) |>
-  tab_row_group(
-    id="Risk factor identification",
-    label = md("**Risk factor identification**"),
-    rows = 3:7
-  )|>
-  tab_row_group(
-    id="Primary prevention",
-    label = md("**Primary prevention**"),
-    rows = 8:12
-  )|>
-  tab_row_group(
-    id="Disease identification",
-    label = md("**Disease identification**"),
-    rows = 13:15
-  )|>
-  tab_row_group(
-    id="Secondary prevention",
-    label = md("**Secondary prevention**"),
-    rows = 16:23
-  )|>
-    tab_row_group(
-    id="Tertiary prevention",
-    label = md("**Tertiary prevention**"),
-    rows = 24:29
-  )|>
-    tab_row_group(
-    id="Intermediate outcome",
-    label = md("**Intermediate outcome**"),
-    rows = 30:32
-  )|>
-    tab_row_group(
-    id="Full outcomes",
-    label = md("**Full outcomes**"),
-    rows = 33:36
-  )|>
-  row_group_order(groups=c("Need","Risk factors","Risk factor identification","Primary prevention","Disease identification","Secondary prevention","Tertiary prevention","Intermediate outcome","Full outcomes"))|>
-  cols_label(name_metric="")
-
+tar_read(cluster2_map)
 ```
 
-::: {.callout-note icon="false"}
-## Cardiac rehabilitation metrics
+### Cluster descriptions
 
-Cardiac rehabilitation metrics will be added once they become available from NACR (expected before 31/3/24)
-:::
-
-# Clustering
-
-In this analysis we used the k-medoids clustering method to group practices.
-Each practice was assigned to one of a small number of groups such that practices within the group have similar distributions of patients over ethnic groups.
-In order to obtain data on the ethnicity of a practice's registered population we imputed it based on the Lower Super Output Area (LSOA) of residence of GP registrants and the ethnicity distribution of patients living in those LSOAs using the Census 2021 data.
-Further detailed explanations of the method can be found in the appendix.
-
-The clustering analysis resulted in 5 groups (clusters) of practices which are illustrated in the charts and map below.
-
-## Cluster descriptions
-
-The median percentage of each ethnic group for each cluster was calculated and is presented in these charts to describe their respective diversity.
-
-### Cluster 1 - Least diverse
+#### Cluster 1 - Least diverse
 
@@
 ::: panel-tabset
-## Five ethnic groups
+## Five categories
 
@@
 
-## Fourteen ethnic groups
+## Fourteen categories
 
@@
 
-### Cluster 2
+#### Cluster 2
 
@@
 
-### Cluster 3
+#### Cluster 3
 
@@
 
-### Cluster 4
+#### Cluster 4
 
@@
 
-### Cluster 5 - Most diverse
+#### Cluster 5 - Most diverse
 
@@
 
-## Clustered GP Practices
-
-The interactive map below presents each GP practice coloured according to their assigned cluster. Cluster 1 (red) is the least diverse and cluster 5 (orange) is the most diverse.
+## Relative Index of Disparity
 
 ```{r}
-tar_read(cluster2_map)
+tar_read(rel_iod_chart)
 ```
 
-# Rate of activity
+```{r}
+tar_read(ci_iod_chart)
+```
 
-The charts below show the activity rates by cluster for each metric along the CHD pathway.
-The horizontal line is the overall rate.
-The 95% confidence intervals are shown at the top of each bar.
-These activity rates take account of need by using CHD prevalence (or list size for risk factor) as the denominator of the calculation.
-Further detail on the methodology can be found in the appendix.
-
-In some cases these upper and lower limits are so small as to not be visible (e.g.
-Obesity register), whereas for other metrics (e.g.
-Readmission within 30 days of a PCI or CABG) they are much wider.
-Metrics with larger volumes of activity data give rise to a greater confidence in the calculated rate (e.g.
-limits are closer together) when compared to metrics with smaller volumes of activity data where there is less confidence in the calculated rate (e.g.
-limits are further apart).
-
-In calculating these rates for each cluster within each metric and then comparing the rates by cluster to the global rate for the metric it is possible to identify how activity rates vary across the 5 clusters.
-For there to be equity between the clusters the rates would need to be the same.
-The charts below present the metrics grouped according to pathway domain.
+## Rate
 
 ### Risk factors
-
-::: callout-tip
-## Key Findings
-
--   Activity rates vary very little for the risk factor metric % patients 45+ with a record of BP \< 5 years
-
--   Rates of patients on the diabetes register are highest in the most diverse cluster
-
--   Rates of patients on the depression register are highest in the least diverse cluster and lowest in the most diverse cluster
-:::
-
-#### Risk factors
 
@@
 
-#### Risk factor identification
+### Risk factor identification
 
@@
 ### Primary prevention
-
-::: callout-tip
-## Key Findings
-
--   The activity rate for CVD patients treated with LLT is highest for those patients whose practice is in the least diverse cluster and lowest for patients whose practice is in cluster 4.
-
--   A similar pattern applies for at risk patients treated with LLT, activity rates being highest for those in the least diverse cluster.
-:::
 
@@
 
-::: callout-tip
-## Key Findings
-
--   Activity rates vary considerably across the clusters in the disease identification metrics, with differing patterns for each metric
-
--   Patients from clusters 4 and 5 (most diverse) are less likely to be recorded on the CHD register
-
--   Clusters 5 and 4 have lower rates of CT angiography and electrocardiography
-:::
-
 ```{r}
@@
 ### Secondary prevention
-
-::: callout-tip
-## Key Findings
-
--   There is little variation between clusters 1 to 4 in referral rates to outpatient cardiology, although the referrals are lower for the most diverse cluster (cluster 5)
-
--   Flu vaccination rates for 65+ patients are highest among patients whose practice is in the least diverse cluster. the two most diverse clusters having the lowest rates
-
--   Amongst patients under 65 who are at risk, flu vaccination rates are less varied between the clusters than for 65+ patients. Cluster 4 has the lowest rate of vaccination
-
--   Cardiology outpatient DNA rates are higher the more diverse the cluster
-:::
 
@@
 
-::: callout-tip
-## Key Findings
-
--   All 4 tertiary prevention metrics follow the same pattern of activity rates with the highest rates being in the least diverse cluster (cluster 1), followed by clusters 2 and then cluster 5 (the most diverse) and cluster 3.
-Cluster 4 has the lowest activity rates.
-:::
-
 ```{r}
@@
 ### Intermediate outcome
-
-::: callout-tip
-## Key Findings
-
--   The least diverse cluster has the highest rates of BP readings \< 140/90.
-
--   Emergency admissions for CHD and readmissions within 30 days of a PCI or CABG are highest for the least diverse cluster and lowest for cluster 4.
-:::
 
@@
 
-::: callout-tip
-## Key Findings
-
-...to be determined - awaiting death data check
-:::
-
 ```{r}
@@
 ```
-
-# Relative Index of Disparity
-
-The relative index of deprivation for the metric is expressed here as a %, such that the % indicates the amount of activity that would need to be redistributed between clusters in order that event rates follow levels of need.
-The detail of the methodology used for its calculation can be found in the appendix.
-
-The chart below presents the metrics along the CHD pathway from risk factor and identification through to outcome measures of death from CHD.
-For each metric the index of disparity is shown as a point estimate (yellow dot) with the upper and lower confidence intervals in grey.
-
-::: callout-tip
-## Key Findings
-
--   The greatest disparity is flu vaccinations for patients aged 65+ where the relative index of disparity is 16.70% (95% ci - 16.64%, 16.70%), indicating that 16.7% of activity needs to be redistributed so that rates reflect need.
-:::
-
-```{r}
-tar_read(ci_iod_chart)
-```
-
-::: {.callout-tip icon="false"}
-## Confidence Intervals
-
-Confidence intervals are shown in grey. There is 95% confidence that the index of disparity is within this range. The confidence interval estimation methods are detailed in the appendix.
-:::
-
-## Routes to equity
-
-A previous Strategy Unit report "Strategies to reduce inequalities in access to planned hospital procedures"[^4] highlights in chapter 1, three of the many routes from inequity to equity: levelling-up, levelling-down and zero-sum redistribution.
-Using the zero-sum redistribution route as an example, along with the metric flu vaccinations for patients aged 65+, this illustration shows how equity could be achieved by increasing and decreasing the number of vaccinations delivered in each cluster to give the same activity rate per cluster.
-
-[^4]: <https://www.midlandsdecisionsupport.nhs.uk/knowledge-library/strategies-to-reduce-inequalities-in-access-to-planned-hospital-procedures/>
-
-|  Cluster  | Number of vaccinations | CHD synthetic prevalence | Rate (No. of vaccinations / CHD synthetic prevalence) | Global rate |  Change in number of vaccinations to match global rate  |
-|:----------:|-----------:|-----------:|:----------:|:----------:|:----------:|
-|           |                    *n* |                      *d* |                        *r=n/d*                        | *gr=𝚺n/𝚺d*  |                      *(r-gr) x d*                       |
-|     1     |              3,356,002 |                1,026,826 |                         3.27                          |    2.33     |                        -961,239                         |
-|     2     |              2,638,698 |                  924,088 |                         2.86                          |    2.33     |                        -483,540                         |
-|     3     |              1,862,430 |                  958,825 |                         1.94                          |    2.33     |                        +373,741                         |
-|     4     |                476,616 |                  478,728 |                         1.00                          |    2.33     |                        +639,873                         |
-|     5     |                332,728 |                  327,542 |                         1.02                          |    2.33     |                        +431,166                         |
-| **Total** |          **8,666,474** |            **3,716,009** |                       **2.33**                        |             |                          **0**                          |
-
-: In reality, in attempting to achieve equity for this metric, it is unlikely that the offer of vaccination would be removed from patients whose practice is in clusters 1 and 2.
-More likely would be an approach to focus public health and vaccination roll-out campaigns on practices in clusters 3, 4 and 5.
-
-Each of the metrics along the CHD pathway with the higher relative indicies of disparity would almost certainly require a variety of different routes from inequity to equity given their differing nature.
-The more effective interventions on the pathway might be levelled-up, whilst those of more limited value might be levelled-down.
-
-# Regional Analysis
-
-## Clusters
-
-The charts below show the distribution of the patients in each NHS region according to the cluster of their GP practice.
-
-::: panel-tabset
-## London
-
-```{r}
-#1
-region_cluster_charts <- tar_read(region_cluster_charts)
-region_cluster_charts[2]
-```
-
-## South West
-
-```{r}
-#2
-region_cluster_charts[7]
-```
-
-## South East
-
-```{r}
-#3
-region_cluster_charts[6]
-```
-
-## Midlands
-
-```{r}
-#4
-region_cluster_charts[3]
-```
-
-## East of England
-
-```{r}
-#5
-region_cluster_charts[1]
-```
-
-## North West
-
-```{r}
-#6
-region_cluster_charts[5]
-```
-
-## North East and Yorkshire
-
-```{r}
-#7
-region_cluster_charts[4]
-```
-:::
-
-## Relative Index of Disparity
-
-::: callout-tip
-## Key Findings
-
--   The greatest disparity
-:::
-
-::: panel-tabset
-## London
-
-```{r}
-regional_charts <- tar_read(regional_charts)
-regional_charts[1]
-```
-
-## South West
-
-```{r}
-regional_charts[2]
-```
-
-## South East
-
-```{r}
-regional_charts[3]
-```
-
-## Midlands
-
-```{r}
-regional_charts[4]
-```
-
-## East of England
-
-```{r}
-regional_charts[5]
-```
-
-## North West
-
-```{r}
-regional_charts[6]
-```
-
-## North East and Yorkshire
-
-```{r}
-regional_charts[7]
-```
-:::
-
-The appendix contains equivalent charts to those above, showing the Index of Disparity for each CHD pathway metric for each of the 42 ICBs.
-
-# Conclusions
-
-.......
-
-# Appendix
-
-## Definitions and data sources of pathway metrics
-
-The table below sets out the sources of the various metrics as well as the time-period to which they relate, and details of the selection criteria used.
-
-| Domain                     | Metric                                 | Data Source | Year    | Definition and selection criteria/codes                                                 |
-|---------------|---------------|---------------|---------------|---------------|
-| Need                       | CHD synthetic prevalence estimates     | PHE         | 2015    | Provided directly by OHID (formerly indicator 92847 in FingertipsR)                     |
-| Risk factors               | Smoking synthetic prevalence estimates | GP survey   | 2023    |                                                                                         |
-| Risk factor identification | Smoking register                       | NHSD QOF    | 2022/23 | [Indicator 91280 extracted fromFingertipsR](https://fingertips.phe.org.uk/search/91280) |
-|                            | Obesity register                       | NHSD QOF    |         | [Indicator 93088 extracted fromFingertipsR](https://fingertips.phe.org.uk/search/93088) |
-|                            | Diabetes register                      | NHSD QOF    |         |                                                                                         |
-|                            | Depression register                    | NHSD QOF    |         |                                                                                         |
-| Primary prevention         |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-|                            |                                        |             |         |                                                                                         |
-| Full Outcomes              | Deaths in hospital from CHD            |             |         |                                                                                         |
-|                            | Deaths in hospital from CHD \< 75      |             |         |                                                                                         |
-|                            | Deaths from CHD                        |             |         |                                                                                         |
-|                            | Deaths from CHD \<75                   |             |         |                                                                                         |
-
-## Methods explained
-
-### QOF data
-
-Metrics taken from the QOF data via Fingertips have been extracted by GP practice and the counts (numerator of the measure) from the calculations have been used.
-Exception reporting metrics have used the Personalised Care Adjustment (PCA) data to obtain the number of patients by GP practice for whom a PCA has been recorded for the relevant metric.
-Possible reasons for a PCA include; Newly diagnosed/registered, Intervention is clinically unsuitable, patient choice, did not respond to offers of care, service not available.
-
-### SUS data
-
-SUS data was extracted using Transact-SQL and using relevant OPCS, ICD10, HRG, and Treatment function codes as detailed in table 2 above.
-
-### Cardiac rehabilitation
-
-This data was supplied by NACR.
-The data wasn’t available by GP Practice, so an alternative method was followed.
-Practices were allocated to a cluster and data was then supplied aggregated into clusters.
-In some cases, the data had to be redacted due to small numbers in the category.
-For the purposes of this analysis the redacted data item was reassigned an average value.
-The true definition of eligible for rehabilitation is the HES admission and successful discharge of all patients with CVD e.g., MI, revascularization, or HF.
-However, to link with the NACR only those records entered into the database can be used so NACR use pseudo eligible as this is as close as the audit can get, at a granular level, to eligible.
-It is though affected by areas without entry to the database, The data used relates to the 2023 calendar year, which is the most recent complete year, and following changes in the way rehabilitation has been delivered due to the Covid pandemic, most closely reflects the rehabilitation model now in place.
-
-### Readmission within 30 days
-
-The methodology used for this metric follows, as far as possible, that detailed by NHS Digital[^5].
-The readmissions counted relate to those who had an elective PCI / CABG in the period 1/3/2022 to 28/2/2023.
-
-[^5]: <https://digital.nhs.uk/data-and-information/publications/statistical/ccg-outcomes-indicator-set/specifications/3.2-emergency-readmissions-within-30-days-of-discharge-from-hospital_1_4>
-
-### ONS Death records
-
-ONS death records contain an encrypted HES ID which was used to link to outpatient, inpatient and A&E HES records, from which the latest GP practice recorded (on a spine traced record) for the patient was then assigned to the death record.
-
-### CHD synthetic prevalence estimates
-
-Public Health England (PHE) CHD synthetic prevalence estimates data was previously available from OHID via Fingertips.
-This data has now been removed from Fingertips and archived, however OHID supplied it via email for this project following a request.
-This data is for 7564 practices from 2015, of which 6297 are current practices.
-In order to assign a prevalence for practices with missing data a nearest neighbours methodology was used to impute the prevalence by taking an average of the prevalence of each practice within a 1.5km radius of the practice without data.
-Further details, including code with sample data, for this methodology are available in a blog post on the Strategy Unit Data Science website[^6].
-
-[^6]: <https://the-strategy-unit.github.io/data_science/blogs/posts/2024-01-17_nearest_neighbour.html>
-
-### Ethnicity by GP practice
-
-Ethnicity by GP practice was imputed based on the Lower Super Output Area (LSOA) of residence of GP registrants and the ethnicity distribution of patients living in these LSOAs using the Census 2021 data.
-
-#### GP list size by LSOA
-
-Since metric data is mainly from the 2022/23 year, the October 2022 list size data was taken from the NHS Digital Patients Registered at a GP practice dataset[^7] as this is the data closest to the mid point in the 22/23 year.
-The GP to ICB mapping file was also downloaded and joined to the list size to easily create analysis by ICB and region.
-
-[^7]: <https://digital.nhs.uk/data-and-information/publications/statistical/patients-registered-at-a-gp-practice/october-2022>
-
-#### Census 2021 Ethnicity by LSOA
-
-The 2021 Census Ethnicity data was downloaded from the .gov.uk website.[^8] Summary details of the census ethnicity data is also available and this includes descriptions of the standardised list of ethnic groups and the differences from the 2011 ethnic groups.[^9]
-
-[^8]: <https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/#download-the-data>
-
-[^9]: <https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/>
-
-The 2021 Census Ethnicity data by LSOA uses the 2021 LSOAs.
-The GP list size data uses the 2011 LSOAs.
-This results in gaps when the datasets are combined.
-To overcome this the following method was used:
-
--   An LSOA lookup showing 2011 and 2021 LSOA parent and child relationships was joined to the 2021 Census Ethnicity data
-
--   Where a 2021 LSOAs has a parent LSOA in 2011 then we applied the average ethnicity % for the child LSOAs to the parent (this is when one 2011 LSOA has been split into multiple 2021 LSOAs).
-
--   Where the LSOAs in 2021 have a child LSOA in 2011 then we applied the ethnicity % for the 2021 parent to the 2011 children (this is when several LSOAs have been combined into one new LSOA).
-
-Further details about item editing and imputation processes for Census 2021, England and Wales are available on the ONS website.
-This details techniques used by ONS with the aim of arriving at a fully populated clean database.
-These techniques include manual imputation and nearest neighbour donor imputation.
-For the England and Wales Census 2021 the imputation rate for ethnic group was 1.3%.[^10]
-
-[^10]: <https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/methodologies/itemeditingandimputationprocessforcensus2021englandandwales>
-
-### K-Medoids Clustering
-
-K-medoids is an established, and commonly used unsupervised machine learning technique.
-This clustering approach required data on the ethnicity of a practices registered population which was obtained as per the methodology explained above (Ethnicity by GP practice).
-The data used in the clustering was the % of patients in each ethnic group per practice, such that for each practice all the ethnic group %'s added to 100%.
-
-The K-medoids clustering was performed using the {cluster} package[^11] in R using partitioning (clustering) of the data into k clusters "around medoids" (PAM), which is a more robust version of K-means.
-The algorithm is based on the search for k representative objects or medoids among the observations of the dataset.
-After finding a set of k medoids, k clusters are constructed by assigning each observation to the nearest medoid.
-The goal is to find k representative objects which minimise the sum of the dissimilarities of the observations to their closest representative object.[^12]
-
-[^11]: Maechler M, Rousseeuw P, Struyf A, Hubert M, Hornik K (2023). *cluster: Cluster Analysis Basics and Extensions*.
-R package version 2.1.6 — For new features, see the 'NEWS' and the 'Changelog' file in the package source), [https://CRAN.R-project.org/package=cluster](https://cran.r-project.org/package=cluster).
-
-[^12]: <https://cran.r-project.org/web/packages/cluster/cluster.pdf>
-
-Clustering experiments were conducted using a variety of different combinations of ethnic breakdowns, such as 14 ethnic groups and 5 ethnic groups, as well as variations that excluded the % of White British patients or factored in age group.
-Techniques were also used to attempt to determine the optimal number of clusters such as the elbow method, gap plots, and scree plots.
-Consideration was given to combining some ethnic groups and in doing so a correlation matrix was produced to identify ethnic groups that commonly featured together.
-It was also important to seek a number of clusters that could be defined easily, where characteristics did not overlap each other so that resulting clusters were too similar to each other.
-It was also necessary to ensure that there was a sufficient number of practices in each cluster, too few practices in a cluster could result in issues with small amounts of metric data when calculating activity rates and consequently lead to wide confidence intervals in the index of disparity calculations.
-
-In conclusion, the final version selected, and presented here, used the percentage of the GP list size in each of the 14 census 2021 ethnic groups and five clusters was felt to give a sufficient number of describable clusters for which the index of disparity calculations could be performed.
-The chart below is an elbow plot generated for the 14 ethnic groups showing an elbow at around 4 or 5 clusters.
-
-```{r}
-tar_read(elbow_plot)
-```
-
-Table 3 shows the number of GP practices assigned to each of the 5 clusters and the characteristics of the clusters are described in the main body of the report above.
-
-```{r}
-final_data_full_cats_percent_5_clusters <- tar_read(final_data_full_cats_percent_5_clusters) |> as_tibble()
-
-final_data_full_cats_percent_5_clusters |>
-  group_by(cluster)|>
-  count("practice_code")|>
-  ungroup() |>
-  select(cluster,n)|>
-  gt()|>
-    tab_header(title = md("**Table 3** - *Number of practices per cluster*"))|>
-    cols_label(cluster=md("**Cluster**"),
-             n=md("**Number of practices**"))
-```
-
-### Index of Disparity calculations
-
-#### Rate of activity
-
-Calculations of the index of disparity begin by calculating an activity rate using a numerator of the activity ( eg number of elective procedures) divided by a denominator, which for metrics other than risk factors has been taken to be CHD synthetic prevalence (as a determinant of need).
-
-Where the metric is in the risk factor domain the denominator used is GP practice list size (with the exception of the percentage of patients aged 45 or over who have a record of blood pressure in the preceding 5 years, where the 45+ list size has been used).
-
-In calculating these rates for each cluster within each metric and then comparing these rates by cluster to the global rate for the metric it is possible to identify how activity rates vary across the 5 clusters
-
-#### Index of disparity
-
-The relative index of disparity indicates the extent to which the rate of an activity or event varies across groups.
-Having calculated the rates for each cluster within each metric the disparity is calculated by taking the differences for each cluster from the global rate and then determining the amount of activity by which this cluster varies from the global.
-This absolute difference is then summed for the 5 clusters and divided by twice the sum of the numerators.
-This produces a relative index of deprivation for the metric, expressed as a %, such that the % indicates the amount of activity that would need to be redistributed between clusters in order that event rates follow levels of need.
-
-Index of disparity = (𝚺 \| r~(1–n)~ - R\| / n) / R\*100
-
-where r = group rate and R = total population rate
-
-A report by Pearcy and Keppel (2002) presents the Index of Disparity as a summary measure of disparity across population groups and advocates its use for groups defined in terms of categories such as ethnicity.[^13]
-
-[^13]: <https://journals.sagepub.com/doi/epdf/10.1093/phr/117.3.273>
-
-### Confidence Intervals
-
-Confidence intervals of 95% are presented in the national charts presented in this report.
-Confidence intervals have been used, rather than calculations of statistical significance as this presents a clearer picture of the likelihood of the disparity and due to concerns with the suitability in this case of statistical significance.
-
-> "Practitioners need to be aware that statistical significance differs from practical importance in that statistical significance is highly dependent upon sample size.
-For a large sample, a statistically significant result is likely to be obtained even when the actual magnitude of an effect is small and of little or no practical importance.
-On the other hand, for a small sample, it is quite likely that insufficient evidence of a statistically significant result will be obtained – even when there is, indeed, an effect of practical importance." (Hahn, G.
-J., Doganaksoy, N., Meeker, W.
-Q., 2019)[^14]
-
-[^14]: <https://academic.oup.com/jrssig/article/16/4/20/7038025>
-
-In order to estimate the confidence intervals for the index of disparity, a bootstrap method was adopted, assuming independence, generating 1000 random samples for each metric for which an index of disparity was calculated.
-The 2.5th and 97.5th percentiles were then taken for each metric to give the upper and lower limits of the 95% confidence interval for each.
-
-An explanation of bootstrap theory and confidence intervals is available in Practical Statistics for Data Scientists (Bruce, P. and Bruce, A., 2017)[^15]
-
-[^15]: <https://www.oreilly.com/library/view/practical-statistics-for/9781491952955/>
-
-## Integrated Care Board (ICB) Analysis
-
-#### Relative Index of Disparity
-
-```{r}
-icb_charts <- tar_read(icb_charts)
-icb_charts[1]
-icb_charts[2]
-icb_charts[3]
-icb_charts[4]
-icb_charts[5]
-icb_charts[6]
-icb_charts[7]
-icb_charts[8]
-icb_charts[9]
-icb_charts[10]
-icb_charts[11]
-icb_charts[12]
-icb_charts[13]
-icb_charts[14]
-icb_charts[15]
-icb_charts[16]
-icb_charts[17]
-icb_charts[18]
-icb_charts[19]
-icb_charts[20]
-icb_charts[21]
-icb_charts[22]
-icb_charts[23]
-icb_charts[24]
-icb_charts[25]
-icb_charts[26]
-icb_charts[27]
-icb_charts[28]
-icb_charts[23]
-icb_charts[29]
-icb_charts[30]
-icb_charts[31]
-icb_charts[32]
-icb_charts[33]
-icb_charts[34]
-icb_charts[35]
-icb_charts[36]
-icb_charts[37]
-icb_charts[38]
-icb_charts[39]
-icb_charts[40]
-icb_charts[41]
-icb_charts[42]
-```
-
-## R code on Github
-
-All the code used in this project is available on the Strategy Unit GitHub[^16].
-
-[^16]: <https://github.com/The-Strategy-Unit/bhf_chd-ethnicity>

### 2024-04-24 deaths_explanation.qmd
Start write up of deaths query

---
title: "deaths_explanation"
format: html
editor: visual
---

## Question

Why are the rates of CHD deaths and CHD hospital deaths lower in the 2 most diverse clusters, yet, "South Asians living in the UK have a higher rate of [coronary heart disease](https://www.bhf.org.uk/informationsupport/conditions/coronary-heart-disease "https://www.bhf.org.uk/informationsupport/conditions/coronary-heart-disease") and [diabetes](https://www.bhf.org.uk/informationsupport/risk-factors/diabetes "https://www.bhf.org.uk/informationsupport/risk-factors/diabetes") compared to White Europeans".

This study refers to the incidence rate.

The method of rate calculation is simply to take the amount of activity in the metric and divide this by the number of patients on the gp list expected to have CHD (calculated by multiplying the PHE synthetic prevalence estimate % by the GP list size).

The PHE synthetic prevalence calculations take ethnicity and age and sex into account in their methodology...

We have calculated Age and Sex Standardised Mortality Rates (direct) for each of the clusters for CHD Deaths and All Cause deaths. These calculations used the same death data as was used in the numerator of the rates from the project.

SC - Prevalent population of the most diverse clusters is younger then would expect fewer deaths.

SC - Prevalence is a function of incidence and mortality and survival is a curve

![](secret/asmrchdprojrates.png){fig-align="center"}

JG - This little table on the right shows the CHD deaths (same data that forms the numerator of my project calculation) as a crude rate per 10,000 population, as well as all cause deaths as  crude rate. The table on the left is mortality rates age and sex standardised using census population. So, if the age and sex standardised mortality rates show the most diverse cluster to have the highest rate then why do the project rates using need not follow the same pattern?

SC - The standardised rate is saying if the populations have no difference in age and sex then cluster 5 is higher than you would expect, too many deaths for its age/sex profile. It is comparing to the total population whether people have the condition or not.

SC -The project chd rate is a case fatality rate, deaths in those that have the condition, this is saying that the highest observed deaths per person with the condition is in cluster 1. It's not comparing like with like and it's not age standardised so could be argued that cluster 1 could be higher because its prevalent population are older.

### 2024-04-24 deaths_qa.R

@@ -175,7 +175,7 @@ all_cause_deaths_2223_gp_prac <- all_cause_deaths_2223|>
   summarise(deaths=sum(deaths))
 
 all_cause_deaths_2223|>filter(gp_practice_code!="NULL")|>
-  group_by(gp_practice_code)|>summarise(deaths=sum(deaths))
+  group_by(gp_practice_code)|>summarise(deaths=sum(num_deaths))
 
 #Cluster population data from gp lists
 
@@ -275,6 +275,8 @@ males_gp_reg_pat <- gp_reg_pat_prac_sing_age_male |>
   select(-age)|>
   mutate(sex=1)
 
+library(gt)
+
 all_gp_reg_pat <- males_gp_reg_pat|>
   rbind(females_gp_reg_pat)|>
   rename(gp_practice_code=org_code)|>
@@ -302,7 +304,7 @@ all_gp_reg_pat <- males_gp_reg_pat|>
     sex_name = md("**Sex**"))|>
   fmt_number(columns = "perc", decimals = 2)
 
-library(gt)
+
   
   all_gp_reg_pat |>
     pivot_wider(names_from=c(cluster,sex_name),
@@ -318,4 +320,5 @@ library(gt)
                 column_labels.font.weight = "bold")|>
     fmt_percent(decimals = 2)|>
     tab_header(title = "Age / Sex Breakdown by Cluster")
+  
   

### 2024-04-24 deaths_explanation.qmd
Added more notes

@@
 
+SW - Are they dying of other causes? = No
+
 ![](secret/asmrchdprojrates.png){fig-align="center"}
 
-JG - This little table on the right shows the CHD deaths (same data that forms the numerator of my project calculation) as a crude rate per 10,000 population, as well as all cause deaths as  crude rate.
+JG - The table on the right shows the CHD deaths (same data that forms the numerator of the project calculation) as a crude rate per 10,000 population, as well as all cause deaths as crude rate.
 The table on the left is mortality rates age and sex standardised using census population.
@@
 It's not comparing like with like and it's not age standardised so could be argued that cluster 1 could be higher because its prevalent population are older.
+
+SC - higher incidence and better survival?
+
+On register:
+
+![](secret/cvdreg.png){fig-align="center"}
+
+Deaths:
+
+![](secret/deaths.png){fig-align="center"}

### 2024-04-25 deaths_explanation.qmd
Updated deaths notes

@@
 
+## Project
+
 The method of rate calculation is simply to take the amount of activity in the metric and divide this by the number of patients on the gp list expected to have CHD (calculated by multiplying the PHE synthetic prevalence estimate % by the GP list size).
 
-The PHE synthetic prevalence calculations take ethnicity and age and sex into account in their methodology...
+The PHE synthetic prevalence calculations take ethnicity and age and sex into account in their methodology.
+
+The factors controlled for in the prevalence model are:
+
+-   age
+
+-   sex
+
+-   ethnicity
+
+-   hypertension
+
+-   diabetes
+
+-   dyslipidemia
+
+-   obesity
+
+-   smoking
+
+-   CKD
+
+-   physical activity
+
+-   family history
+
+-   area deprivation (townsend score)
+
+SW - Ethnicity is a covariate in the prevalence model although the variable only has two levels, white / non-white.
+ The effect is quite large.
+ They estimate that a non-white individual is 33% more likely to develop CHD than a white person with similar risk factors (age, sex, family history, etc etc).
+
+Since their prevalence estimates account for the fact that a non-white patient would have 33% higher likelihood of developing CHD then this feeds into the denominator of our calculations and the larger denominator leads to a lower rate.
+This increased risk of developing CHD offsets the lower age profile of ethnic minority populatons (which is why the distribution of list size over the clusters is similar to the distribution of prevalent cases).
+ But we are still saying that having controlled for prevalence, the more diverse clusters have lower death rates.
+ We might have expected that more diverse clusters would have similar or higher death rates having controlled for prevalence.
 
@@
 
-SC - higher incidence and better survival?
+SC - higher incidence but better survival is possible
+
+SC - \*\*the project rates are different to asmr because the asmr are population age standardised but the case fatality in the project that we are comparing to are not \*\* .
+It's complicated by not comparing like with like as the first is a measure of the rate death from CHD occurs in the population and the second is a measure of how many people already with the condition die.
 
@@
 ![](secret/deaths.png){fig-align="center"}
+
+Ecological Fallacy:
+
+link to Robinson's paradox
+
+ASMR only accounts for age and sex, it is partial and biased, doesnt account for need.
+CHD prevalence as a need denominator does control for other influencing factors such as deprivation, ethnicity etc.
+
+Next steps
+
+Take 2 practices and examine
+
+Under 75 mortality and another indicator split it by deprivation and by clusters

### 2024-04-25 deaths_qa.R
Expand death explanation

@@ -1,5 +1,6 @@
 library(targets)
 library(janitor)
+library(tidyverse)
 
 clustered_gp_and_metrics <- tar_read(clustered_gp_and_metrics)|>as_tibble()
 activity_by_type_clusters_stg6 <- tar_read(activity_by_type_clusters_stg6)|>as_tibble()
@@ -322,3 +323,37 @@ all_gp_reg_pat <- males_gp_reg_pat|>
     tab_header(title = "Age / Sex Breakdown by Cluster")
   
   
+  
+#  cluster 1 practive B83033 gp_imd_decile=5 middle
+# cluster 5 practice B83052 gp_imd_decile=1 most deprived
+  
+test_practices <-  clustered_gp_and_metrics |>
+  filter(metric1 != "NA") |>
+  filter(cluster2==1|cluster2==5)|>
+ # filter(gp_practice_code=="B83033"|gp_practice_code=="B83052")|>
+  select(gp_practice_code,cluster2,metric20,metric29b,list_size,metric1)|>
+  mutate(list_size = replace_na(list_size, 0)) |>
+  mutate(metric1 = replace_na(metric1, 0)) |>
+  mutate(metric20 = replace_na(metric20, 0)) |>
+  mutate(metric29b = replace_na(metric29b, 0)) |>
+  mutate(need=metric1/list_size,
+         metric20_needrate=metric20/metric1,
+         metric29b_needrate=metric29b/metric1,
+         metric20_listrate=metric20/list_size,
+         metric29b_listrate=metric29b/list_size
+         )|>
+  left_join(patientweighted_practice_imd)|>
+  select(-comm_region_code,-comm_region_name,-icb_name,-icb_code,-practice_name,-cluster)|>
+  group_by(cluster2)|>
+  summarise(metric20_total=sum(metric20),
+            metric29b_total=sum(metric29b),
+            listsize_total=sum(list_size),
+            metric1_total=sum(metric1))|>
+  mutate(need=metric1_total/listsize_total,
+         metric20_needrate=metric20_total/metric1_total,
+         metric29b_needrate=metric29b_total/metric1_total,
+         metric20_listrate=metric20_total/listsize_total,
+         metric29b_listrate=metric29b_total/listsize_total
+  )
+
+

### 2024-04-26 draft_report.qmd
Further report additions re deaths

@@
 
-Table 3 shows that the most diverse cluster has the greatest proportion of patients in practices in the most deprived quantile.
+Table 3 shows that the most diverse cluster (cluster 5) has the greatest proportion of patients in practices in the most deprived quantile.
 Cluster 2 has the greatest proportion of patients in practices in the least deprived quantile.
@@
 patientweighted_practice_imd |>
-  mutate(gp_imd_quantile = case_when(gp_imd_decile<=2 ~ 1,
-                                     gp_imd_decile==3 ~2,
-                                     gp_imd_decile==4 ~2,
-                                     gp_imd_decile==5 ~3,
-                                     gp_imd_decile==6 ~3,
-                                     gp_imd_decile==7 ~4,
-                                     gp_imd_decile==8 ~4,
-                                     gp_imd_decile>=9 ~5,
-                                     .default = 0
+  mutate(gp_imd_quantile = case_when(gp_imd_decile<=2 ~ "1 - most deprived",
+                                     gp_imd_decile==3 ~"2",
+                                     gp_imd_decile==4 ~"2",
+                                     gp_imd_decile==5 ~"3",
+                                     gp_imd_decile==6 ~"3",
+                                     gp_imd_decile==7 ~"4",
+                                     gp_imd_decile==8 ~"4",
+                                     gp_imd_decile>=9 ~"5 - least deprived",
+                                     .default = "0"
   ))|>
@@
   arrange(factor(gp_imd_quantile, 
-                 levels = c("1","2","3","4","5")))|>
+                 levels = c("1 - most deprived","2","3","4","5 - least deprived")))|>
 
@@
 
-...to be determined - awaiting death data checks
+-   The least diverse cluster has the highest rate of CHD deaths and premature CHD deaths. Cluster 4 has the lowest rate of CHD deaths.
+
+-   CHD hospital deaths and premature CHD hospital deaths also have the lowest rate in cluster 4.
+Premature CHD hospital deaths have a similar rate for the other clusters including the most diverse cluster.
 :::
@@
 ```
+
+::: {.callout-note appearance="minimal"}
+Work led by Professor Sir Michael Marmot and colleagues in the late 1980s revealed that first-generation South Asians living in the UK have a higher rate of coronary heart disease and diabetes compared to White Europeans[^4].
+This gives rise to the question, why are the death rates lower in the more diverse clusters?
+There are a number of factors to consider in understanding this including; the age/sex profile of the clusters, the methodology for calculating the CHD synthetic prevalence (the measure of need in this project), and incidence rates vs mortality rates.
+This is examined further in the appendix.
+:::
+
+[^4]: <https://www.bhf.org.uk/what-we-do/our-research/research-successes/ethnicity-and-heart-disease>
 
@@
 
-A previous Strategy Unit report "Strategies to reduce inequalities in access to planned hospital procedures"[^4] highlights in chapter 1, three of the many routes from inequity to equity: levelling-up, levelling-down and zero-sum redistribution.
+A previous Strategy Unit report "Strategies to reduce inequalities in access to planned hospital procedures"[^5] highlights in chapter 1, three of the many routes from inequity to equity: levelling-up, levelling-down and zero-sum redistribution.
 Using the zero-sum redistribution route as an example, along with the metric flu vaccinations for patients aged 65+, the illustration below shows how equity could be achieved by increasing and decreasing the number of vaccinations delivered in each cluster to give the same activity rate per cluster.
 
-[^4]: <https://www.midlandsdecisionsupport.nhs.uk/knowledge-library/strategies-to-reduce-inequalities-in-access-to-planned-hospital-procedures/>
+[^5]: <https://www.midlandsdecisionsupport.nhs.uk/knowledge-library/strategies-to-reduce-inequalities-in-access-to-planned-hospital-procedures/>
 
@@
 |  Cluster  | Number of vaccinations | CHD synthetic prevalence | Rate (No. of vaccinations / CHD synthetic prevalence) | Global rate | Change in number of vaccinations to match global rate |
-|:---------:|----------:|----------:|:------------:|:---------:|:------------:|
+|:----------:|-----------:|-----------:|:----------:|:----------:|:----------:|
 |           |                    *n* |                      *d* |                        *r=n/d*                        | *gr=𝚺n/𝚺d*  |                     *(gr-r) x d*                      |
@@
 
-The methodology used for this metric follows, as far as possible, that detailed by NHS Digital[^5].
+The methodology used for this metric follows, as far as possible, that detailed by NHS Digital[^6].
 The readmissions counted relate to those who had an elective PCI / CABG in the period 1/3/2022 to 28/2/2023.
 
-[^5]: <https://digital.nhs.uk/data-and-information/publications/statistical/ccg-outcomes-indicator-set/specifications/3.2-emergency-readmissions-within-30-days-of-discharge-from-hospital_1_4>
+[^6]: <https://digital.nhs.uk/data-and-information/publications/statistical/ccg-outcomes-indicator-set/specifications/3.2-emergency-readmissions-within-30-days-of-discharge-from-hospital_1_4>
 
@@
 In order to assign a prevalence for practices with missing data a nearest neighbours methodology was used to impute the prevalence by taking an average of the prevalence of each practice within a 1.5km radius of the practice without data.
-Further details, including code with sample data, for this methodology are available in a blog post on the Strategy Unit Data Science website[^6].
-
-[^6]: <https://the-strategy-unit.github.io/data_science/blogs/posts/2024-01-17_nearest_neighbour.html>
+Further details, including code with sample data, for this methodology are available in a blog post on the Strategy Unit Data Science website[^7].
+
+[^7]: <https://the-strategy-unit.github.io/data_science/blogs/posts/2024-01-17_nearest_neighbour.html>
 
@@
 
-Since metric data is mainly from the 2022/23 year, the October 2022 list size data was taken from the NHS Digital Patients Registered at a GP practice dataset[^7] as this is the data closest to the mid point in the 22/23 year.
+Since metric data is mainly from the 2022/23 year, the October 2022 list size data was taken from the NHS Digital Patients Registered at a GP practice dataset[^8] as this is the data closest to the mid point in the 22/23 year.
 The GP to ICB mapping file was also downloaded and joined to the list size to easily create analysis by ICB and region.
 
-[^7]: <https://digital.nhs.uk/data-and-information/publications/statistical/patients-registered-at-a-gp-practice/october-2022>
+[^8]: <https://digital.nhs.uk/data-and-information/publications/statistical/patients-registered-at-a-gp-practice/october-2022>
 
@@
 
-The 2021 Census Ethnicity data was downloaded from the .gov.uk website.[^8] Summary details of the census ethnicity data is also available and this includes descriptions of the standardised list of ethnic groups and the differences from the 2011 ethnic groups.[^9]
-
-[^8]: <https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/#download-the-data>
-
-[^9]: <https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/>
+The 2021 Census Ethnicity data was downloaded from the .gov.uk website.[^9] Summary details of the census ethnicity data is also available and this includes descriptions of the standardised list of ethnic groups and the differences from the 2011 ethnic groups.[^10]
+
+[^9]: <https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/#download-the-data>
+
+[^10]: <https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/>
 
@@
 These techniques include manual imputation and nearest neighbour donor imputation.
-For the England and Wales Census 2021 the imputation rate for ethnic group was 1.3%.[^10]
-
-[^10]: <https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/methodologies/itemeditingandimputationprocessforcensus2021englandandwales>
+For the England and Wales Census 2021 the imputation rate for ethnic group was 1.3%.[^11]
+
+[^11]: <https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/methodologies/itemeditingandimputationprocessforcensus2021englandandwales>
 
@@
 
-The K-medoids clustering was performed using the {cluster} package[^11] in R using partitioning (clustering) of the data into k clusters "around medoids" (PAM), which is a more robust version of K-means.
+The K-medoids clustering was performed using the {cluster} package[^12] in R using partitioning (clustering) of the data into k clusters "around medoids" (PAM), which is a more robust version of K-means.
 The algorithm is based on the search for k representative objects or medoids among the observations of the dataset.
 After finding a set of k medoids, k clusters are constructed by assigning each observation to the nearest medoid.
-The goal is to find k representative objects which minimise the sum of the dissimilarities of the observations to their closest representative object.[^12]
-
-[^11]: Maechler M, Rousseeuw P, Struyf A, Hubert M, Hornik K (2023). *cluster: Cluster Analysis Basics and Extensions*.
+The goal is to find k representative objects which minimise the sum of the dissimilarities of the observations to their closest representative object.[^13]
+
+[^12]: Maechler M, Rousseeuw P, Struyf A, Hubert M, Hornik K (2023). *cluster: Cluster Analysis Basics and Extensions*.
 R package version 2.1.6 — For new features, see the 'NEWS' and the 'Changelog' file in the package source), [https://CRAN.R-project.org/package=cluster](https://cran.r-project.org/package=cluster).
 
-[^12]: <https://cran.r-project.org/web/packages/cluster/cluster.pdf>
+[^13]: <https://cran.r-project.org/web/packages/cluster/cluster.pdf>
 
@@
 
-A report by Pearcy and Keppel (2002) presents the Index of Disparity as a summary measure of disparity across population groups and advocates its use for groups defined in terms of categories such as ethnicity.[^13]
-
-[^13]: <https://journals.sagepub.com/doi/epdf/10.1093/phr/117.3.273>
+A report by Pearcy and Keppel (2002) presents the Index of Disparity as a summary measure of disparity across population groups and advocates its use for groups defined in terms of categories such as ethnicity.[^14]
+
+[^14]: <https://journals.sagepub.com/doi/epdf/10.1093/phr/117.3.273>
 
@@
 J., Doganaksoy, N., Meeker, W.
-Q., 2019)[^14]
-
-[^14]: <https://academic.oup.com/jrssig/article/16/4/20/7038025>
+Q., 2019)[^15]
+
+[^15]: <https://academic.oup.com/jrssig/article/16/4/20/7038025>
 
@@
 
-An explanation of bootstrap theory and confidence intervals is available in Practical Statistics for Data Scientists (Bruce, P. and Bruce, A., 2017)[^15]
-
-[^15]: <https://www.oreilly.com/library/view/practical-statistics-for/9781491952955/>
+An explanation of bootstrap theory and confidence intervals is available in Practical Statistics for Data Scientists (Bruce, P. and Bruce, A., 2017)[^16]
+
+[^16]: <https://www.oreilly.com/library/view/practical-statistics-for/9781491952955/>
+
+## CHD death rates
+
+In examining the findings of this project the question arises, why are the death rates lower in the more diverse clusters? There are a number of factors to consider in understanding this:
+
+::: {.callout-note appearance="minimal" icon="false"}
+##### Age/Sex profile of the clusters
+
+The age/sex profile of the clusters can be seen in table 2.
+This shows that the most diverse clusters have a much younger profile, 68% are under 45 in the most diverse cluster, compared to 48% of the least diverse cluster.
+Those aged 75+ represent only 4% of the most diverse cluster, compared to 12% of the least diverse cluster.
+:::
+
+::: {.callout-note appearance="minimal" icon="false"}
+## Methodology for calculating CHD synthetic prevalence
+
+The measure of need used in this project to calculate the activity rates is the PHE synthetic prevalence estimates of CHD.
+The prevalence calculations take ethnicity and age and sex into account in their methodology.
+Other factors include: hypertension, diabetes, dyslipidemia, obesity, smoking, CKD, physical activity, family history, and area deprivation (townsend score).
+Ethnicity is a covariate in the prevalence model (although the variable only has two levels, white / non-white).
+ It is estimated that a non-white individual is 33% more likely to develop CHD than a white person with similar risk factors (age, sex, family history, etc etc).
+The prevalence estimates account for the fact that a non-white patient would have a higher likelihood of developing CHD which feeds into the denominator of the project calculations and the larger denominator leads to a lower rate.
+
+Examining clusters 1 and 5 in more detail, the % CHD prevalence (need) is much higher in cluster 5 (most diverse) and this gives rise to activity rates which are lower even when the % activity volume is the same as cluster 1.
+This is illustrated with sample data in table 7 below.
+:::
+
+| Cluster       | CHD Prevalence (%) |  List size | CHD Prevalence | CHD Deaths (% of list size) | Activity rate (CHD Deaths/Prevalence) |
+|------------|:----------:|--------:|----------:|:----------:|:-------------:|
+| Least diverse |        8.7%        | 14,000,000 |      1,218,000 |        6000 (0.043%)        |                0.00492                |
+| Most diverse  |       10.6%        |  4,000,000 |        424,000 |        1714 (0.043%)        |                0.00404                |
+
+: **Table 7**- Example data to illustrate the effect of higher prevalence on activity rates
+
+::: {.callout-note appearance="minimal"}
+## Incidence rates vs mortality rates
+
+The Age and Sex Standardised Mortality Rates (direct) for each of the clusters for CHD Deaths and All Cause deaths have been calculated.
+These calculations used the same death data as was used in the numerator of the activity rate calculations.
+Table 8 .
+Are they dying of other causes?
+= No
+:::
+
+![](images/clipboard-3667502623.png)
+
+| Cluster | Population | ASMR - All cause | ASMR - CHD | Rate - All Cause | Rate CHD |
+|:-------:|------------|------------------|------------|------------------|----------|
+|    1    |            |                  |            |                  |          |
+|    2    |            |                  |            |                  |          |
+|    3    |            |                  |            |                  |          |
+|    4    |            |                  |            |                  |          |
+|    5    |            |                  |            |                  |          |
+
+: **Table 8** - ASMR
+
+JG - The table on the right shows the CHD deaths (same data that forms the numerator of the project calculation) as a crude rate per 10,000 population, as well as all cause deaths as crude rate.
+The table on the left is mortality rates age and sex standardised using census population.
+So, if the age and sex standardised mortality rates show the most diverse cluster to have the highest rate then why do the project rates using need not follow the same pattern?
+
+Show the charts with the deaths by age group and gender for each cluster - Stacey
+
+Having controlled for prevalence, the more diverse clusters have lower death rates.
+ We might have expected that more diverse clusters would have similar or higher death rates having controlled for prevalence.
+
+SC - Prevalent population of the most diverse clusters is younger then would expect fewer deaths.
+
+SC - Prevalence is a function of incidence and mortality and survival is a curve
+
+SC - The standardised rate is saying if the populations have no difference in age and sex then cluster 5 is higher than you would expect, too many deaths for its age/sex profile.
+It is comparing to the total population whether people have the condition or not.
+
+SC -The project chd rate is a case fatality rate, deaths in those that have the condition, this is saying that the highest observed deaths per person with the condition is in cluster 1.
+It's not comparing like with like and it's not age standardised so could be argued that cluster 1 could be higher because its prevalent population are older.
+
+SC - \*\*the project rates are different to asmr because the asmr are population age standardised but the case fatality in the project that we are comparing to are not \*\* .
+It's complicated by not comparing like with like as the first is a measure of the rate death from CHD occurs in the population and the second is a measure of how many people already with the condition die.
 
@@
 
-All the code used in this project is available on the Strategy Unit GitHub[^16].
-
-[^16]: <https://github.com/The-Strategy-Unit/bhf_chd-ethnicity>
+All the code used in this project is available on the Strategy Unit GitHub[^17].
+
+[^17]: <https://github.com/The-Strategy-Unit/bhf_chd-ethnicity>

### 2024-04-26 deaths_qa.R

@@ -327,7 +327,7 @@ all_gp_reg_pat <- males_gp_reg_pat|>
 #  cluster 1 practive B83033 gp_imd_decile=5 middle
 # cluster 5 practice B83052 gp_imd_decile=1 most deprived
   
-test_practices <-  clustered_gp_and_metrics |>
+examine_clusters <-  clustered_gp_and_metrics |>
   filter(metric1 != "NA") |>
   filter(cluster2==1|cluster2==5)|>
  # filter(gp_practice_code=="B83033"|gp_practice_code=="B83052")|>
@@ -356,4 +356,38 @@ test_practices <-  clustered_gp_and_metrics |>
          metric29b_listrate=metric29b_total/listsize_total
   )
 
+test_practices <-  clustered_gp_and_metrics |>
+  filter(metric1 != "NA") |>
+  filter(cluster2==1|cluster2==5)|>
+  # filter(gp_practice_code=="B83033"|gp_practice_code=="B83052")|>
+  select(gp_practice_code,cluster2,metric20,metric29b,list_size,metric1)|>
+  mutate(list_size = replace_na(list_size, 0)) |>
+  mutate(metric1 = replace_na(metric1, 0)) |>
+  mutate(metric20 = replace_na(metric20, 0)) |>
+  mutate(metric29b = replace_na(metric29b, 0)) |>
+  mutate(need=metric1/list_size,
+         metric20_needrate=metric20/metric1,
+         metric29b_needrate=metric29b/metric1,
+         metric20_listrate=metric20/list_size,
+         metric29b_listrate=metric29b/list_size
+  )|>
+  left_join(patientweighted_practice_imd)|>
+  select(-comm_region_code,-comm_region_name,-icb_name,-icb_code,-practice_name,-cluster)
+
+
+dist_20 <- test_practices|>
+  group_by(metric20)|>
+  count(gp_practice_code)|>
+  ungroup()|>
+  group_by(metric20)|>
+  summarise(n=sum(n))|>
+  print(n=24)
+
+dist_29b <- test_practices|>
+  group_by(metric29b,cluster2)|>
+  count(gp_practice_code)|>
+  ungroup()|>
+  group_by(metric29b,cluster2)|>
+  summarise(n=sum(n))|>
+  print(n=29)
 

### 2024-04-29 draft_report.qmd
Further explanation of death findings and chart tweak

@@
 This gives rise to the question, why are the death rates lower in the more diverse clusters?
-There are a number of factors to consider in understanding this including; the age/sex profile of the clusters, the methodology for calculating the CHD synthetic prevalence (the measure of need in this project), and incidence rates vs mortality rates.
-This is examined further in the appendix.
+There are a number of factors to consider in understanding this including; the age/sex profile of the clusters, the methodology for calculating the CHD synthetic prevalence (the measure of need in this project), and age and sex standardised mortality rates (ASMR) compared to the project CHD mortality rates.
+These are examined further in the appendix.
 :::
@@
 
-## CHD death rates
-
-In examining the findings of this project the question arises, why are the death rates lower in the more diverse clusters? There are a number of factors to consider in understanding this:
+## CHD mortality rates
+
+In examining the findings of this project the question arises, why are the mortality rates lower in the more diverse clusters? There are a number of factors to consider in understanding this:
 
@@
 ::: {.callout-note appearance="minimal" icon="false"}
-## Methodology for calculating CHD synthetic prevalence
-
-The measure of need used in this project to calculate the activity rates is the PHE synthetic prevalence estimates of CHD.
+##### Methodology for calculating CHD synthetic prevalence
+
+The measure of need used in this project to calculate the activity rates is the PHE synthetic prevalence estimate of CHD.
 The prevalence calculations take ethnicity and age and sex into account in their methodology.
@@
 Examining clusters 1 and 5 in more detail, the % CHD prevalence (need) is much higher in cluster 5 (most diverse) and this gives rise to activity rates which are lower even when the % activity volume is the same as cluster 1.
-This is illustrated with sample data in table 7 below.
+This is illustrated with example data in table 7 below.
 :::
@@
 | Cluster       | CHD Prevalence (%) |  List size | CHD Prevalence | CHD Deaths (% of list size) | Activity rate (CHD Deaths/Prevalence) |
-|------------|:----------:|--------:|----------:|:----------:|:-------------:|
+|---------------|:------------------:|-----------:|---------------:|:---------------------------:|:-------------------------------------:|
 | Least diverse |        8.7%        | 14,000,000 |      1,218,000 |        6000 (0.043%)        |                0.00492                |
@@
 
-: **Table 7**- Example data to illustrate the effect of higher prevalence on activity rates
+: **Table 7**- *Example data to illustrate the effect of higher prevalence on activity rates*
 
 ::: {.callout-note appearance="minimal"}
-## Incidence rates vs mortality rates
+##### Mortality rates
 
 The Age and Sex Standardised Mortality Rates (direct) for each of the clusters for CHD Deaths and All Cause deaths have been calculated.
-These calculations used the same death data as was used in the numerator of the activity rate calculations.
-Table 8 .
-Are they dying of other causes?
-= No
-:::
-
-![](images/clipboard-3667502623.png)
-
-| Cluster | Population | ASMR - All cause | ASMR - CHD | Rate - All Cause | Rate CHD |
-|:-------:|------------|------------------|------------|------------------|----------|
-|    1    |            |                  |            |                  |          |
-|    2    |            |                  |            |                  |          |
-|    3    |            |                  |            |                  |          |
-|    4    |            |                  |            |                  |          |
-|    5    |            |                  |            |                  |          |
-
-: **Table 8** - ASMR
-
-JG - The table on the right shows the CHD deaths (same data that forms the numerator of the project calculation) as a crude rate per 10,000 population, as well as all cause deaths as crude rate.
-The table on the left is mortality rates age and sex standardised using census population.
-So, if the age and sex standardised mortality rates show the most diverse cluster to have the highest rate then why do the project rates using need not follow the same pattern?
-
-Show the charts with the deaths by age group and gender for each cluster - Stacey
+These calculations use the same death data as was used in the numerator of the activity rate calculations.
+They are a measure of the rate death from CHD occurs in the population, if the populations have no difference in age and sex then the more diverse clusters are higher than would be expected.
+It is comparing to the total population whether those in that population have the condition or not.
+
+These rates can be seen in table 8 below, alongside the CHD deaths crude rate from this project and an all cause deaths crude rate also calculated by dividing the number of deaths by the CHD prevalence and showing per 10,000 population.
+These crude rates are a measure of how many people already with CHD die.
+:::
+
+| Cluster | Population | ASMR - All cause | ASMR - CHD | Rate - All Cause (All deaths / Prevalence) | Rate - CHD (CHD deaths / Prevalence) |
+|:-------:|-----------:|-----------------:|-----------:|-------------------------------------------:|-------------------------------------:|
+|    1    | 16,914,882 |            913.5 |       93.9 |                                     1188.4 |                                123.1 |
+|    2    | 15,788,339 |            894.3 |       87.6 |                                     1029.5 |                                100.5 |
+|    3    | 16,416,864 |            935.4 |       93.3 |                                      813.7 |                                 80.6 |
+|    4    |  7,883,280 |            773.7 |       82.2 |                                      431.4 |                                 45.9 |
+|    5    |  4,961,545 |            868.2 |      105.8 |                                      511.1 |                                 62.7 |
+
+: **Table 8** - *ASMR and Activity Rate*
 
 Having controlled for prevalence, the more diverse clusters have lower death rates.
- We might have expected that more diverse clusters would have similar or higher death rates having controlled for prevalence.
-
-SC - Prevalent population of the most diverse clusters is younger then would expect fewer deaths.
-
-SC - Prevalence is a function of incidence and mortality and survival is a curve
-
-SC - The standardised rate is saying if the populations have no difference in age and sex then cluster 5 is higher than you would expect, too many deaths for its age/sex profile.
-It is comparing to the total population whether people have the condition or not.
-
-SC -The project chd rate is a case fatality rate, deaths in those that have the condition, this is saying that the highest observed deaths per person with the condition is in cluster 1.
-It's not comparing like with like and it's not age standardised so could be argued that cluster 1 could be higher because its prevalent population are older.
-
-SC - \*\*the project rates are different to asmr because the asmr are population age standardised but the case fatality in the project that we are comparing to are not \*\* .
-It's complicated by not comparing like with like as the first is a measure of the rate death from CHD occurs in the population and the second is a measure of how many people already with the condition die.
+ We might have expected that more diverse clusters would have similar or higher death rates.
+There are a number of factors potentially contributing to this finding.
+The CHD deaths crude rate is a case fatality rate, deaths in those that have CHD.
+The highest observed deaths per person with CHD is in the least diverse cluster.
+It is not age standardised so it could be argued that the rate is higher in this cluster because its prevalent population are older and in the more diverse clusters it is younger so fewer deaths would be expected.
+
+In July 2021 ONS produced an experimental analysis of ethnic differences in life-expectancy and cause-specific mortality rates[^17].
+The analysis found that White and Mixed ethnic groups had lower life expectancy than all other ethnic groups, while the Black African ethnic group had statistically significant higher life expectancy than most groups.
+However, the analysis also found that ischaemic heart disease mortality was highest in the Bangladeshi, Pakistani and Indian ethnic groups and lowest among Black ethnic groups.
+The analysis revealed complex patterns in life expectancy and mortality by ethnic group.
+Potential reasons are offered for the higher life expectancy found in Black African and Asian Other ethnic groups including that they contain a greater proportion of more recent migrants than other ethnic groups and it also cites previous research showing that people who migrate tend to be healthier than others.
+The ONS report also suggests that Asian and Black ethnic groups engage less in harmful health-related behaviours.
+These findings show some similarities to research in the USA into what has been termed the Hispanic paradox, where most Hispanic groups are characterised by low economic status and a high prevalence of cardiovascular disease risk factors, but better than expected health and mortality outcomes[^18].
+Possible explanations for the Hispanic paradox include the healthy migrant hypothesis (those who migrate tend to be younger and healthier), the presence of nuclear families and high levels of social support, and dietary factors.
+
+[^17]: <https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/lifeexpectancies/articles/ethnicdifferencesinlifeexpectancyandmortalityfromselectedcausesinenglandandwales/2011to2014>
+
+[^18]: <https://www.sciencedirect.com/science/article/abs/pii/S0033062014001339?via%3Dihub>
+
+In this project the more diverse clusters comprise of a mixture of different ethnic groups, in particular clusters 4 and 5 (most diverse).
+It is possible that a combination of factors described above are giving rise to the mortality by cluster and that the Black ethnic group in cluster 4, with the lowest mortality, is offsetting the higher mortality of the Asian ethnic group, alongside the higher CHD prevalence in the denominator and the younger age profile, as well as factors such as the healthy migrant hypothesis, and higher levels of social support.
 
@@
 
-All the code used in this project is available on the Strategy Unit GitHub[^17].
-
-[^17]: <https://github.com/The-Strategy-Unit/bhf_chd-ethnicity>
+All the code used in this project is available on the Strategy Unit GitHub[^19].
+
+[^19]: <https://github.com/The-Strategy-Unit/bhf_chd-ethnicity>

### 2024-04-29 R/chart_functions.R

@@ -220,8 +220,8 @@ get_ci_iod_chart <- function(iod_with_ci){
     )) |>
     ggplot() +
     geom_segment( aes(x=metric_long_name, xend=metric_long_name, y=lower_ci*100, yend=upper_ci*100), color="#686f73") +
-    geom_point( aes(x=metric_long_name, y=lower_ci*100), color="#686f73", size=2.5 ) +
-    geom_point( aes(x=metric_long_name, y=upper_ci*100), color="#686f73", size=2.5 ) +
+    #geom_point( aes(x=metric_long_name, y=lower_ci*100), color="#686f73", size=2.5 ) +
+    #geom_point( aes(x=metric_long_name, y=upper_ci*100), color="#686f73", size=2.5 ) +
     geom_point( aes(x=metric_long_name, y=iod*100), color="#f9bf07", size=2.5 ) +
     coord_flip()+
     facet_grid(rows = factor(pathway_level,levels=c('Risk', 'Risk identif.', '1°c prevent', 'Dis. ident','2°c prevent',

### 2024-04-29 draft_report.qmd
Additional references and tweaked chart heights

@@
 This measure relies on the ordered quality of the socio-economic deprivation variable.
-To handle the categorical nature of the ethnicity variable we have used the relative index of disparity to indicate the extent to which the rate of an activity or event varies across groups.
+To handle the categorical nature of the ethnicity variable we have used the relative index of disparity[^4] to indicate the extent to which the rate of an activity or event varies across groups.
 The index estimates the proportion of events (e.g., admissions) that would need to be redistributed between clusters in order that event rates follow levels of need.
 Further detailed explanations of the methods used in the analysis can be found in the appendix.
+
+[^4]: *Pearcy J, Keppel K, A Summary Measure of Health Disparity, Public Health Reports, Vol 117, May-June 2002*
+
+    [*https://open.umich.edu/sites/default/files/downloads/PublicHealthRep-Pearcy.pdf*](https://gbr01.safelinks.protection.outlook.com/?url=https%3A%2F%2Fopen.umich.edu%2Fsites%2Fdefault%2Ffiles%2Fdownloads%2FPublicHealthRep-Pearcy.pdf&data=05%7C02%7Cjacqueline.grout1%40nhs.net%7Ca9c82e4d3cc94a72de8408dc68333fc0%7C37c354b285b047f5b22207b48d774ee3%7C0%7C0%7C638499816545595235%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C0%7C%7C%7C&sdata=ddk0rdD6HnQ1524JNCsTfr3AoiBePESeckGAAnQPSA4%3D&reserved=0)
 
@@
 ::: {.callout-note appearance="minimal"}
-Work led by Professor Sir Michael Marmot and colleagues in the late 1980s revealed that first-generation South Asians living in the UK have a higher rate of coronary heart disease and diabetes compared to White Europeans[^4].
+Work led by Professor Sir Michael Marmot and colleagues in the late 1980s revealed that first-generation South Asians living in the UK have a higher rate of coronary heart disease and diabetes compared to White Europeans[^5].
 This gives rise to the question, why are the death rates lower in the more diverse clusters?
@@
 
-[^4]: <https://www.bhf.org.uk/what-we-do/our-research/research-successes/ethnicity-and-heart-disease>
+[^5]: <https://www.bhf.org.uk/what-we-do/our-research/research-successes/ethnicity-and-heart-disease>
 
@@
 ```{r}
+#| out-height: 600px
+
 tar_read(ci_iod_chart)
@@
 
-A previous Strategy Unit report "Strategies to reduce inequalities in access to planned hospital procedures"[^5] highlights in chapter 1, three of the many routes from inequity to equity: levelling-up, levelling-down and zero-sum redistribution.
+A previous Strategy Unit report "Strategies to reduce inequalities in access to planned hospital procedures"[^6] highlights in chapter 1, three of the many routes from inequity to equity: levelling-up, levelling-down and zero-sum redistribution.
 Using the zero-sum redistribution route as an example, along with the metric flu vaccinations for patients aged 65+, the illustration below shows how equity could be achieved by increasing and decreasing the number of vaccinations delivered in each cluster to give the same activity rate per cluster.
 
-[^5]: <https://www.midlandsdecisionsupport.nhs.uk/knowledge-library/strategies-to-reduce-inequalities-in-access-to-planned-hospital-procedures/>
+[^6]: <https://www.midlandsdecisionsupport.nhs.uk/knowledge-library/strategies-to-reduce-inequalities-in-access-to-planned-hospital-procedures/>
 
@@
 |  Cluster  | Number of vaccinations | CHD synthetic prevalence | Rate (No. of vaccinations / CHD synthetic prevalence) | Global rate | Change in number of vaccinations to match global rate |
-|:---------:|-----------------------:|-------------------------:|:-----------------------------------------------------:|:-----------:|:-----------------------------------------------------:|
+|:---------:|----------:|----------:|:------------:|:---------:|:------------:|
 |           |                    *n* |                      *d* |                        *r=n/d*                        | *gr=𝚺n/𝚺d*  |                     *(gr-r) x d*                      |
@@
 ```{r}
+#| out-height: 600px
 regional_charts <- tar_read(regional_charts)
@@
 ```{r}
+#| out-height: 600px
 regional_charts[2]
@@
 ```{r}
+#| out-height: 600px
 regional_charts[3]
@@
 ```{r}
+#| out-height: 600px
 regional_charts[4]
@@
 ```{r}
+#| out-height: 600px
 regional_charts[5]
@@
 ```{r}
+#| out-height: 600px
 regional_charts[6]
@@
 ```{r}
+#| out-height: 600px
 regional_charts[7]
@@
 
-The methodology used for this metric follows, as far as possible, that detailed by NHS Digital[^6].
+The methodology used for this metric follows, as far as possible, that detailed by NHS Digital[^7].
 The readmissions counted relate to those who had an elective PCI / CABG in the period 1/3/2022 to 28/2/2023.
 
-[^6]: <https://digital.nhs.uk/data-and-information/publications/statistical/ccg-outcomes-indicator-set/specifications/3.2-emergency-readmissions-within-30-days-of-discharge-from-hospital_1_4>
+[^7]: <https://digital.nhs.uk/data-and-information/publications/statistical/ccg-outcomes-indicator-set/specifications/3.2-emergency-readmissions-within-30-days-of-discharge-from-hospital_1_4>
 
@@
 In order to assign a prevalence for practices with missing data a nearest neighbours methodology was used to impute the prevalence by taking an average of the prevalence of each practice within a 1.5km radius of the practice without data.
-Further details, including code with sample data, for this methodology are available in a blog post on the Strategy Unit Data Science website[^7].
-
-[^7]: <https://the-strategy-unit.github.io/data_science/blogs/posts/2024-01-17_nearest_neighbour.html>
+Further details, including code with sample data, for this methodology are available in a blog post on the Strategy Unit Data Science website[^8].
+
+[^8]: <https://the-strategy-unit.github.io/data_science/blogs/posts/2024-01-17_nearest_neighbour.html>
 
@@
 
-Since metric data is mainly from the 2022/23 year, the October 2022 list size data was taken from the NHS Digital Patients Registered at a GP practice dataset[^8] as this is the data closest to the mid point in the 22/23 year.
+Since metric data is mainly from the 2022/23 year, the October 2022 list size data was taken from the NHS Digital Patients Registered at a GP practice dataset[^9] as this is the data closest to the mid point in the 22/23 year.
 The GP to ICB mapping file was also downloaded and joined to the list size to easily create analysis by ICB and region.
 
-[^8]: <https://digital.nhs.uk/data-and-information/publications/statistical/patients-registered-at-a-gp-practice/october-2022>
+[^9]: <https://digital.nhs.uk/data-and-information/publications/statistical/patients-registered-at-a-gp-practice/october-2022>
 
@@
 
-The 2021 Census Ethnicity data was downloaded from the .gov.uk website.[^9] Summary details of the census ethnicity data is also available and this includes descriptions of the standardised list of ethnic groups and the differences from the 2011 ethnic groups.[^10]
-
-[^9]: <https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/#download-the-data>
-
-[^10]: <https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/>
+The 2021 Census Ethnicity data was downloaded from the .gov.uk website.[^10] Summary details of the census ethnicity data is also available and this includes descriptions of the standardised list of ethnic groups and the differences from the 2011 ethnic groups.[^11]
+
+[^10]: <https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/#download-the-data>
+
+[^11]: <https://www.ethnicity-facts-figures.service.gov.uk/uk-population-by-ethnicity/demographics/age-groups/latest/>
 
@@
 These techniques include manual imputation and nearest neighbour donor imputation.
-For the England and Wales Census 2021 the imputation rate for ethnic group was 1.3%.[^11]
-
-[^11]: <https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/methodologies/itemeditingandimputationprocessforcensus2021englandandwales>
+For the England and Wales Census 2021 the imputation rate for ethnic group was 1.3%.[^12]
+
+[^12]: <https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/methodologies/itemeditingandimputationprocessforcensus2021englandandwales>
 
@@
 
-The K-medoids clustering was performed using the {cluster} package[^12] in R using partitioning (clustering) of the data into k clusters "around medoids" (PAM), which is a more robust version of K-means.
+The K-medoids clustering was performed using the {cluster} package[^13] in R using partitioning (clustering) of the data into k clusters "around medoids" (PAM), which is a more robust version of K-means.
 The algorithm is based on the search for k representative objects or medoids among the observations of the dataset.
 After finding a set of k medoids, k clusters are constructed by assigning each observation to the nearest medoid.
-The goal is to find k representative objects which minimise the sum of the dissimilarities of the observations to their closest representative object.[^13]
-
-[^12]: Maechler M, Rousseeuw P, Struyf A, Hubert M, Hornik K (2023). *cluster: Cluster Analysis Basics and Extensions*.
+The goal is to find k representative objects which minimise the sum of the dissimilarities of the observations to their closest representative object.[^14]
+
+[^13]: Maechler M, Rousseeuw P, Struyf A, Hubert M, Hornik K (2023). *cluster: Cluster Analysis Basics and Extensions*.
 R package version 2.1.6 — For new features, see the 'NEWS' and the 'Changelog' file in the package source), [https://CRAN.R-project.org/package=cluster](https://cran.r-project.org/package=cluster).
 
-[^13]: <https://cran.r-project.org/web/packages/cluster/cluster.pdf>
+[^14]: <https://cran.r-project.org/web/packages/cluster/cluster.pdf>
 
@@
 
-A report by Pearcy and Keppel (2002) presents the Index of Disparity as a summary measure of disparity across population groups and advocates its use for groups defined in terms of categories such as ethnicity.[^14]
-
-[^14]: <https://journals.sagepub.com/doi/epdf/10.1093/phr/117.3.273>
+A report by Pearcy and Keppel (2002) presents the Index of Disparity as a summary measure of disparity across population groups and advocates its use for groups defined in terms of categories such as ethnicity.[^15]
+
+[^15]: <https://journals.sagepub.com/doi/epdf/10.1093/phr/117.3.273>
 
@@
 J., Doganaksoy, N., Meeker, W.
-Q., 2019)[^15]
-
-[^15]: <https://academic.oup.com/jrssig/article/16/4/20/7038025>
+Q., 2019)[^16]
+
+[^16]: <https://academic.oup.com/jrssig/article/16/4/20/7038025>
 
@@
 
-An explanation of bootstrap theory and confidence intervals is available in Practical Statistics for Data Scientists (Bruce, P. and Bruce, A., 2017)[^16]
-
-[^16]: <https://www.oreilly.com/library/view/practical-statistics-for/9781491952955/>
+An explanation of bootstrap theory and confidence intervals is available in Practical Statistics for Data Scientists (Bruce, P. and Bruce, A., 2017)[^17]
+
+[^17]: <https://www.oreilly.com/library/view/practical-statistics-for/9781491952955/>
 
@@
 | Cluster       | CHD Prevalence (%) |  List size | CHD Prevalence | CHD Deaths (% of list size) | Activity rate (CHD Deaths/Prevalence) |
-|---------------|:------------------:|-----------:|---------------:|:---------------------------:|:-------------------------------------:|
+|------------|:----------:|-----------:|-----------:|:----------:|:-------------:|
 | Least diverse |        8.7%        | 14,000,000 |      1,218,000 |        6000 (0.043%)        |                0.00492                |
@@
 | Cluster | Population | ASMR - All cause | ASMR - CHD | Rate - All Cause (All deaths / Prevalence) | Rate - CHD (CHD deaths / Prevalence) |
-|:-------:|-----------:|-----------------:|-----------:|-------------------------------------------:|-------------------------------------:|
+|:---------:|----------:|----------:|----------:|---------------:|------------:|
 |    1    | 16,914,882 |            913.5 |       93.9 |                                     1188.4 |                                123.1 |
@@
 
-In July 2021 ONS produced an experimental analysis of ethnic differences in life-expectancy and cause-specific mortality rates[^17].
+In July 2021 ONS produced an experimental analysis of ethnic differences in life-expectancy and cause-specific mortality rates[^18].
 The analysis found that White and Mixed ethnic groups had lower life expectancy than all other ethnic groups, while the Black African ethnic group had statistically significant higher life expectancy than most groups.
@@
 The ONS report also suggests that Asian and Black ethnic groups engage less in harmful health-related behaviours.
-These findings show some similarities to research in the USA into what has been termed the Hispanic paradox, where most Hispanic groups are characterised by low economic status and a high prevalence of cardiovascular disease risk factors, but better than expected health and mortality outcomes[^18].
+These findings show some similarities to research in the USA into what has been termed the Hispanic paradox, where most Hispanic groups are characterised by low economic status and a high prevalence of cardiovascular disease risk factors, but better than expected health and mortality outcomes[^19].
 Possible explanations for the Hispanic paradox include the healthy migrant hypothesis (those who migrate tend to be younger and healthier), the presence of nuclear families and high levels of social support, and dietary factors.
 
-[^17]: <https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/lifeexpectancies/articles/ethnicdifferencesinlifeexpectancyandmortalityfromselectedcausesinenglandandwales/2011to2014>
-
-[^18]: <https://www.sciencedirect.com/science/article/abs/pii/S0033062014001339?via%3Dihub>
+[^18]: <https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/lifeexpectancies/articles/ethnicdifferencesinlifeexpectancyandmortalityfromselectedcausesinenglandandwales/2011to2014>
+
+[^19]: <https://www.sciencedirect.com/science/article/abs/pii/S0033062014001339?via%3Dihub>
 
@@
 
-All the code used in this project is available on the Strategy Unit GitHub[^19].
-
-[^19]: <https://github.com/The-Strategy-Unit/bhf_chd-ethnicity>
+All the code used in this project is available on the Strategy Unit GitHub[^20].
+
+[^20]: <https://github.com/The-Strategy-Unit/bhf_chd-ethnicity>

### 2024-07-24 final_report.qmd
Add data tables, download buttons and conclusion

--- 2024-04-29 draft_report.qmd
@@
 ---
-title: "Visualising CHD Pathway Inequalities by ethnicity - DRAFT"
+title: "Visualising CHD Pathway Inequalities by ethnicity"
 lang: en-GB
@@
 library(gridExtra)
+library(downloadthis)
 
@@
 
+::: {.callout-tip collapse="true" appearance="minimal" icon="false"}
+## Activity rates data
+
+```{r}
+rate_chart_data <- tar_read(rate_chart_data) |> as_tibble()
+
+rate_chart_data |> 
+  dplyr::rename(cluster=cluster2)|>
+  dplyr::select(-name) |>
+  gt::gt() |>
+  gt::cols_move(columns = c(rate,lower_ci,upper_ci, global_rate),
+                after = metric_name) |>
+  gt::fmt_number(columns=c(rate,lower_ci,upper_ci, global_rate),
+                 decimals=3) |>
+  gt::cols_label(cluster =gt::md("**Cluster**"),
+                 pathway = gt::md("**Pathway**"),
+                 metric_name = gt::md("**Metric**"),
+                 rate= gt::md("**Rate**"),
+                 lower_ci = gt::md("**Lower CI**"),
+                 upper_ci = gt::md("**Upper CI**"),
+                 global_rate = gt::md("**Global Rate**")) |>
+  gt::tab_row_group(
+    label = gt::md("**Full outcomes**"),
+    rows = pathway == "Full outcomes"
+  ) |>
+  gt::tab_row_group(
+    label = gt::md("**Intermediate outcome**"),
+    rows = pathway == "Intermediate outcome"
+  ) |>
+  gt::tab_row_group(
+    label = gt::md("**Tertiary prevention**"),
+    rows = pathway == "Tertiary prevention"
+  ) |>
+  gt::tab_row_group(
+    label = gt::md("**Secondary prevention**"),
+    rows = pathway == "Secondary prevention"
+  ) |>
+  gt::tab_row_group(
+    label = gt::md("**Disease identification**"),
+    rows = pathway == "Disease identification"
+  ) |>
+  gt::tab_row_group(
+    label = gt::md("**Primary prevention**"),
+    rows = pathway == "Primary prevention"
+  ) |>
+  gt::tab_row_group(
+    label = gt::md("**Risk factor identification**"),
+    rows = pathway == "Risk factor identification"
+  ) |>
+  gt::tab_row_group(
+    label = gt::md("**Risk factors**"),
+    rows = pathway == "Risk factors"
+  ) 
+```
+:::
+
+```{r}
+rate_chart_data %>%
+  download_this(
+    output_name = "rate_data",
+    output_extension = ".xlsx",
+    button_label = "Download activity rate data",
+    button_type = "default",
+    has_icon = TRUE,
+    icon = "fa fa-save"
+  )
+```
+
 # Relative Index of Disparity
@@
 :::
+
+::: {.callout-tip collapse="true" appearance="minimal" icon="false"}
+## Index of Disparity data
+
+```{r}
+rate_chart_data <- tar_read(rate_chart_data) |> as_tibble()
+iod_with_ci <- tar_read(iod_with_ci) |> as_tibble()
+
+metric_names <- rate_chart_data |> 
+  select(pathway,metric_name,name) |> 
+  unique() |> 
+  rename(metric_name_full=metric_name, metric_name=name)
+
+disparity_data <- iod_with_ci |> left_join(metric_names)
+
+disparity_data |> 
+  select(-metric_name) |>
+  gt() |>
+  cols_move(columns = c(iod,lower_ci,upper_ci),
+            after = metric_name_full) |>
+  tab_row_group(
+    label = md("**Full outcomes**"),
+    rows = pathway == "Full outcomes"
+  ) |>
+  tab_row_group(
+    label = md("**Intermediate outcome**"),
+    rows = pathway == "Intermediate outcome"
+  ) |>
+  tab_row_group(
+    label = md("**Tertiary prevention**"),
+    rows = pathway == "Tertiary prevention"
+  ) |>
+  tab_row_group(
+    label = md("**Secondary prevention**"),
+    rows = pathway == "Secondary prevention"
+  ) |>
+  tab_row_group(
+    label = md("**Disease identification**"),
+    rows = pathway == "Disease identification"
+  ) |>
+  tab_row_group(
+    label = md("**Primary prevention**"),
+    rows = pathway == "Primary prevention"
+  ) |>
+  tab_row_group(
+    label = md("**Risk factor identification**"),
+    rows = pathway == "Risk factor identification"
+  ) |>
+  tab_row_group(
+    label = md("**Risk factors**"),
+    rows = pathway == "Risk factors"
+  ) |>
+
+  fmt_number(columns=c(iod,lower_ci,upper_ci),
+             decimals=3) |>
+  cols_label(pathway = md("**Pathway**"),
+             metric_name_full = md("**Metric**"),
+             iod= md("**IOD**"),
+             lower_ci = md("**Lower CI**"),
+             upper_ci = md("**Upper CI**"))
+```
+:::
+
+```{r}
+disparity_data %>%
+  download_this(
+    output_name = "iod_data",
+    output_extension = ".xlsx",
+    button_label = "Download iod data",
+    button_type = "default",
+    has_icon = TRUE,
+    icon = "fa fa-save"
+  )
+```
 
@@
 region_cluster_charts <- tar_read(region_cluster_charts)
-region_cluster_charts[2]
+region_cluster_charts[[2]]
 ```
@@
 #2
-region_cluster_charts[7]
+region_cluster_charts[[7]]
 ```
@@
 #3
-region_cluster_charts[6]
+region_cluster_charts[[6]]
 ```
@@
 #4
-region_cluster_charts[3]
+region_cluster_charts[[3]]
 ```
@@
 #5
-region_cluster_charts[1]
+region_cluster_charts[[1]]
 ```
@@
 #6
-region_cluster_charts[5]
+region_cluster_charts[[5]]
 ```
@@
 #7
-region_cluster_charts[4]
+region_cluster_charts[[4]]
 ```
@@
 
-To be added as part of finalisation of the report.
+In conclusion, there are three main aspects to this analysis on which to reflect: methods, findings and next steps.
+
+Assessing inequalities across ethnic groups was challenging on a number of levels, it was both experimental and complex.
+Ethnicity can’t be expressed as a set of ordered groups and there is no meaningful way to numerically aggregate into a single variable.
+Consequently, relative index of inequality can’t be used.
+
+This analysis attempted to overcome these challenges using novel methods, creating 5 clusters of GP practices using K-Medoids clustering on ethnicity %, followed by use of the relative index of disparity to indicate the extent to which the rate of an activity or event on the CHD pathway varied across the clusters.
+
+Further challenges arose during the analysis, which are detailed in the appendix, however this analysis has made progress in the methods used and they are useful in presenting a visualisation of the variation in disparity of equity across the CHD pathway.
+
+The results of the analysis should be viewed as tentative rather than actionable, adding to a fuller picture alongside other work and thinking on inequalities and CHD.
+
+Early findings show that variation clearly exists across the pathway with some metrics showing less than 1% disparity (blood pressure checks within 5 years for 45+), whilst others have as much as 17% disparity (flu vaccinations for 65+).
+
+The analysis found that risk factors varied little across clusters, although diabetes rates were identified more in the most diverse cluster and depression was identified more in the least diverse cluster.
+Moving through the pathway, CVD patients or at risk patients had higher rates of lipid lowering therapies in the least diverse clusters.
+Rates of diagnosis (on the CHD register, CT angiography, and electrocardiography) were all lower in the more diverse clusters.
+
+Moving across the pathway towards treatment and care, inequity showed up more between clusters in some of the secondary and tertiary prevention metrics in particular CHD patients receiving Asprin, APT or ACT (12.5%), starting cardiac rehabilitation(11%) and Elective CABG (10%).
+
+Mortality (CHD deaths and premature CHD deaths, including in hospital) was higher in the least diverse cluster, which may reflect age and sex profiles of the population (and which is examined and discussed in further detail in the appendix).
+
+A lot of risk factors had similar rates to those found when examining the pathway through a socio-economic lens but a lot of diagnostics such as being on the CHD register or getting a CT Angiography were lower in the more diverse clusters.
+There are also parallels with the previous work examining socio-economic inequalities along the pathway for secondary and tertiary prevention metrics with many of these metrics such as Elective PCI and CABG and receipt of Asprin etc also showing that inequity favoured the least deprived.
+
+Many of these findings give rise to further questions and therefore there are a number of further routes for analysis and possible next steps that could be taken to extend and follow up this project.
+
+Whilst this project includes some regional and ICB level presentations, in some cases the majority of GP practices in a region or ICB have the same cluster.
+The analysis could be recreated at a regional level, whereby new unique clusters could be assigned separately for a region, potentially giving rise to more nuanced clusters from which local disparities could be more easily identified.
+
+The methods used in this project could be transferred to other pathways and services, re-using the same clusters and then selecting different sets of metrics for example related to COPD, Cancer, Mental Health etc.
+
+The code for this project is available on GitHub and as such this enables the work to be reused and extended more easily.
+
+Many of the metrics used in this analysis were only available at a GP practice level, necessitating some of the approaches taken.
+Where record level data is available this would allow for more detailed exploration and understanding of some of these initial findings.
+Of particular interest would be further exploration of the findings surrounding mortality rates and the inter-relationships between different factors of inequality such as ethnicity and deprivation, as well as better understanding at an individual ethnicity level.
 
@@
 ```{r}
+#| output: "asis"
 icb_charts <- tar_read(icb_charts)
-icb_charts[1]
-icb_charts[2]
-icb_charts[3]
-icb_charts[4]
-icb_charts[5]
-icb_charts[6]
-icb_charts[7]
-icb_charts[8]
-icb_charts[9]
-icb_charts[10]
-icb_charts[11]
-icb_charts[12]
-icb_charts[13]
-icb_charts[14]
-icb_charts[15]
-icb_charts[16]
-icb_charts[17]
-icb_charts[18]
-icb_charts[19]
-icb_charts[20]
-icb_charts[21]
-icb_charts[22]
-icb_charts[23]
-icb_charts[24]
-icb_charts[25]
-icb_charts[26]
-icb_charts[27]
-icb_charts[28]
-icb_charts[23]
-icb_charts[29]
-icb_charts[30]
-icb_charts[31]
-icb_charts[32]
-icb_charts[33]
-icb_charts[34]
-icb_charts[35]
-icb_charts[36]
-icb_charts[37]
-icb_charts[38]
-icb_charts[39]
-icb_charts[40]
-icb_charts[41]
-icb_charts[42]
+
+purrr::iwalk(icb_charts, \(plot, name) {
+  name <- str_extract(name, "(?<=^NHS ).*(?= Integrated Care Board$)")
+  cat("\n\n#####", name, "\n\n")
+  print(plot)
+})
+```
+
+```{r}
+
+all_icb_data <- purrr::map(icb_charts, "data") |> dplyr::bind_rows(.id = "icb")
+
+all_icb_data %>%
+  download_this(
+    output_name = "iod_icb_data",
+    output_extension = ".xlsx",
+    button_label = "Download ICB IoD data",
+    button_type = "default",
+    has_icon = TRUE,
+    icon = "fa fa-save"
+  )
 ```

### 2024-07-24 data_exports.R

@@ -1,0 +1,15 @@
+
+  
+# This is it!  
+the_data <- purrr::map(icb_charts, "data") |> dplyr::bind_rows(.id = "icb") 
+write.csv(the_data, "icb_data.csv")
+
+
+rate_chart_data <- targets::tar_read(rate_chart_data) |> dplyr::as_tibble()
+rate_chart_data <- rate_chart_data |> 
+  dplyr::rename(cluster=cluster2)
+write.csv(rate_chart_data, "rate_chart_data.csv")
+
+iod_with_ci <- targets::tar_read(iod_with_ci) |> dplyr::as_tibble()
+disparity_data <- iod_with_ci |> dplyr::left_join(metric_names)
+write.csv(disparity_data, "iod_data.csv")

### 2024-07-24 deaths_qa.R

@@ -1,393 +1,0 @@
-library(targets)
-library(janitor)
-library(tidyverse)
-
-clustered_gp_and_metrics <- tar_read(clustered_gp_and_metrics)|>as_tibble()
-activity_by_type_clusters_stg6 <- tar_read(activity_by_type_clusters_stg6)|>as_tibble()
-gp_reg_pat_prac_sing_age_female<- tar_read(gp_reg_pat_prac_sing_age_female)|>as_tibble()
-gp_reg_pat_prac_sing_age_male<- tar_read(gp_reg_pat_prac_sing_age_male)|>as_tibble()
-gp_reg_pat_prac_lsoa <- tar_read(gp_reg_pat_prac_lsoa)|>as_tibble()
-clusters_for_nacr <- tar_read(clusters_for_nacr)|> as_tibble()
-
-need <- activity_by_type_clusters_stg6|>select(cluster2,metric1_total)|>rename(chdprev_need=metric1_total)
-
-gp_75over_perc <- gp_reg_pat_prac_sing_age_male |>
-  rbind(gp_reg_pat_prac_sing_age_female)|>
-  filter(age != "ALL") |>
-  group_by(org_code,age) |>
-  summarise(number_of_patients=sum(number_of_patients)) |>
-  dplyr::mutate(age=as.numeric(case_when(age=="95+" ~ 95,
-                                         .default = as.numeric(age))
-  )) |>
-  mutate(over_75=case_when(age>= 75 ~1,
-                           .default = 0)) |>
-  group_by(org_code,over_75)|>
-  summarise(number_of_patients=sum(number_of_patients))|>
-  group_by(org_code) |>
-  mutate(total_patients=sum(number_of_patients))|>
-  ungroup() |>
-  mutate(perc=(number_of_patients/total_patients)*100)|>
-  filter(over_75==1) |>
-  select(org_code,perc)|>
-  rename(gp_practice_code=org_code,perc_over75=perc)
-
-gp_16over_female_num <- gp_reg_pat_prac_sing_age_female |>
-  filter(age != "ALL") |>
-  group_by(org_code,age) |>
-  summarise(number_of_patients=sum(number_of_patients)) |>
-  dplyr::mutate(age=as.numeric(case_when(age=="95+" ~ 95,
-                                         .default = as.numeric(age))
-  )) |>
-  mutate(over_16=case_when(age>= 16 ~1,
-                           .default = 0)) |>
-  group_by(org_code,over_16)|>
[…]
